---
title: "dataprep_temporalspatial"
author: "Aleah Wong"
date: '2024-08-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

############################################################
1. Calculate annual production totals for each species (data saved/provided)
2. Clean species names and prepare dataset (data saved/provided)
3. Graphing production (don't need to run steps 1-2)
4. Calculating production by taxa--use production values to weight FCB scores by taxonomic category
############################################################

#####STEP 1: Calculate annual production totals for each species --> save 3 csv files, "inland_totals.csv"; "marine_totals.csv"; "nonfish_totals.csv" #####

```{r Aquaculture production data}
#load in FAO data to calculate total production in each year

# inland <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/inland_finfish_prod.csv")
# marine <-read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/marine_finfish_prod.csv")
# nonfish <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/nonfish_prod.csv")
```

```{r Annual production of inland species}

# str(inland)
# inland <- inland %>%mutate_at(c(6:77), as.numeric) #change yearly data to numeric
# 
# unique(inland$ASFIS.species.Name.En) #find unique species in data
# #create list of each species to loop through
# unique <- c("Bighead carp", "Black carp", "Catla", "Common carp", "Crucian carp", "Grass carp(=White amur)", "Largemouth black bass", "Nile tilapia", "Rainbow trout", "Roho labeo", "Silver carp", "Striped catfish", "Wuchang bream", "North African catfish")
# 
# new_frame <- filter(inland, ASFIS.species.Name.En == unique[1]) #create new frame
# new_frame <- new_frame[FALSE, ]
# 
# for (i in 1:length(unique)){ #for each of the species
#   u_species <- filter(inland, ASFIS.species.Name.En == unique[i])
# 
# #calculate the annual mean each year
# 
#   for (b in 6:77){
#     a <- sum(u_species[, b], na.rm = TRUE)
#     print(a)
#     new_frame[i, 1] <- unique[i]
#     new_frame[i, b] <-a #change row number
#   }
# }
# 
# new_frame <- new_frame[, c(1, 6:77)] #cut out unnecessary columns
# names(new_frame)[1] <- "Species" #rename first column
# 
# # #save this data
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/inland_totals.csv")
# write.csv(new_frame,file=NAME)
```

```{r Annual production of marine species}

# str(marine)
# marine <- marine %>%mutate_at(c(6:77), as.numeric) #change yearly data to numeric
# 
# unique(marine$ASFIS.species.Name.En) #find unique species in data
# 
# #create list of each species to loop through
# unique <- c("Atlantic salmon", "Barramundi(=Giant seaperch)", "Coho(=Silver) salmon", "European seabass", "Flathead grey mullet", "Gilthead seabream", "Japanese amberjack", "Japanese seabass", "Large yellow croaker", "Milkfish", "Nile tilapia", "Orange-spotted grouper", "Pompano", "Rainbow trout", "Red drum")
# 
# new_frame <- filter(marine, ASFIS.species.Name.En == unique[1]) #create new frame
# new_frame <- new_frame[FALSE, ]
# 
# for (i in 1:length(unique)){ #for each of the species
#   u_species <- filter(marine, ASFIS.species.Name.En == unique[i])
# 
# #calculate the annual mean each year
# 
#   for (b in 6:77){
#     a <- sum(u_species[, b], na.rm = TRUE)
#     print(a)
#     new_frame[i, 1] <- unique[i]
#     new_frame[i, b] <-a #change row number
#   }
# }
# 
# new_frame <- new_frame[, c(1, 6:77)] #cut out unnecessary columns
# names(new_frame)[1] <- "Species" #rename first column
# # 
# # #save this data
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/marine_totals.csv")
# write.csv(new_frame,file=NAME)
```

```{r Annual production of nonfish species}

# str(nonfish)
# nonfish <- nonfish %>%mutate_at(c(6:77), as.numeric) #change yearly data to numeric
# 
# unique(nonfish$ASFIS.species.Name.En) #find unique species in data
# 
# #create list of each species to loop through
# unique <- c("Blood cockle", "Blue mussel", "Chilean mussel", "Chinese mitten crab", "Constricted tagelus", "Elkhorn sea moss", "Eucheuma seaweeds nei", "Fusiform sargassum", "Giant river prawn", "Giant tiger prawn", "Gracilaria seaweeds", "Green mud crab", "Indo-Pacific swamp crab", "Japanese carpet shell", "Japanese kelp", "Laver (Nori)", "Mediterranean mussel", "New Zealand mussel", "Nori nei", "Oriental river prawn", "Pacific cupped oyster", "Red swamp crawfish", "Scallops nei", "Sea mussels nei", "Spiny eucheuma", "Wakame", "Whiteleg shrimp", "Yesso scallop", "American cupped oyster")
# 
# new_frame <- filter(nonfish, ASFIS.species.Name.En == unique[1]) #create new frame
# new_frame <- new_frame[FALSE, ]
# 
# for (i in 1:length(unique)){ #for each of the species
#   u_species <- filter(nonfish, ASFIS.species.Name.En == unique[i])
# 
# #calculate the annual mean each year
# 
#   for (b in 6:77){
#     a <- sum(u_species[, b], na.rm = TRUE)
#     print(a)
#     new_frame[i, 1] <- unique[i]
#     new_frame[i, b] <-a #change row number
#   }
# }
# 
# new_frame <- new_frame[, c(1, 6:77)] #cut out unnecessary columns
# names(new_frame)[1] <- "Species" #rename first column
# 
# #save this data
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/nonfish_totals.csv")
# write.csv(new_frame,file=NAME)
```

### STEP 2: Cleaning data to create a dataset with species, category and annual production; save 2 csv files, "prodbytaxa_alphab.csv" (has alphabetized scientific names, version 1); "prodbytaxa_clean.csv" (final dataset for graphing production) ############

```{r Cleaning species names in the data }

#load data
# setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Data")

# inland_totals <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/inland_totals.csv")
# marine_totals <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/marine_totals.csv")
# nonfish_totals <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/nonfish_totals.csv")
# nonfish_totals <- nonfish_totals[, c(-75)] #?
# 
# #rename (inland and marine) for nile tilapia and rainbow trout
# inland_totals$Species[inland_totals$Species == "Rainbow trout"] <- "Oncorhynchus mykiss (inland)"
# inland_totals$Species[inland_totals$Species == "Nile tilapia"] <- "Oreochromis niloticus (inland)"
# 
# #rename (inland and marine) for nile tilapia and rainbow trout
# marine_totals$Species[marine_totals$Species == "Rainbow trout"] <- "Oncorhynchus mykiss (marine)"
# marine_totals$Species[marine_totals$Species == "Nile tilapia"] <- "Oreochromis niloticus (marine)"
# 
# #remove random extra column in category_totals
# # category_totals <- category_totals[, c(-1)]
# 
# #combine the datasets with rowbind
# cat_totals <- rbind(inland_totals, marine_totals)
# cat_totals <- rbind(cat_totals, nonfish_totals)
# cat_totals <- cat_totals[, c(-1)]
# 
# #rename pompano, get rid of nori nei, scallops nei and sea mussels nei
# cat_totals <- cat_totals[!cat_totals$Species == "Nori nei",]
# cat_totals <- cat_totals[!cat_totals$Species == "Scallops nei",]
# cat_totals <- cat_totals[!cat_totals$Species == "Sea mussels nei",]
```

```{r Aligning species names in production dataset with FCB scores}
#put both datasets into the same order so the species line up
#rename the cat_totals species to be scientific names

# cat_totals$Species[cat_totals$Species == "American cupped oyster"] <- "Crassostrea virginica"
# cat_totals$Species[cat_totals$Species == "North African catfish"] <- "Clarias gariepinus"
# cat_totals$Species[cat_totals$Species == "Pompano"] <- "Trachinotus ovatus"
# 
# cat_totals$Species[cat_totals$Species == "Black carp"] <- "Mylopharyngodon piceus"
# cat_totals$Species[cat_totals$Species == "Bighead carp"] <- "Hypophthalmichthys nobilis"
# cat_totals$Species[cat_totals$Species == "Catla"] <- "Catla catla"
# cat_totals$Species[cat_totals$Species == "Common carp"] <- "Cyprinus carpio"
# cat_totals$Species[cat_totals$Species == "Crucian carp"] <- "Carassius carassius"
# cat_totals$Species[cat_totals$Species == "Grass carp(=White amur)"] <- "Ctenopharyngodon idella"
# cat_totals$Species[cat_totals$Species == "Largemouth black bass"] <- "Micropterus salmoides"
# cat_totals$Species[cat_totals$Species == "Nile tilapia"] <- "Oreochromis niloticus"
# cat_totals$Species[cat_totals$Species == "Rainbow trout"] <- "Oncorhynchus mykiss"
# cat_totals$Species[cat_totals$Species == "Roho labeo"] <- "Labeo rohita"
# cat_totals$Species[cat_totals$Species == "Silver carp"] <- "Hypophthalmichthys molitrix"
# cat_totals$Species[cat_totals$Species == "Striped catfish"] <- "Pangasianodon hypophthalmus"
# cat_totals$Species[cat_totals$Species == "Wuchang bream"] <- "Megalobrama amblycephala"
# cat_totals$Species[cat_totals$Species == "Atlantic salmon"] <- "Salmo salar"
# cat_totals$Species[cat_totals$Species == "Barramundi(=Giant seaperch)"] <- "Lates calcarifer"
# cat_totals$Species[cat_totals$Species == "Coho(=Silver) salmon"] <- "Oncorhynchus kisutch"
# cat_totals$Species[cat_totals$Species == "European seabass"] <- "Dicentrarchus labrax"
# cat_totals$Species[cat_totals$Species == "Flathead grey mullet"] <- "Mugil cephalus"
# cat_totals$Species[cat_totals$Species == "Gilthead seabream"] <- "Sparus aurata"
# cat_totals$Species[cat_totals$Species == "Japanese amberjack"] <- "Seriola quinqueradiata"
# cat_totals$Species[cat_totals$Species == "Japanese seabass"] <-  "Lateolabrax japonicus"
# cat_totals$Species[cat_totals$Species == "Large yellow croaker"] <- "Larimichthys croceus"
# cat_totals$Species[cat_totals$Species == "Milkfish"] <- "Chanos chanos"
# cat_totals$Species[cat_totals$Species == "Nile tilapia"] <- "Oreochromis niloticus"
# cat_totals$Species[cat_totals$Species == "Orange-spotted grouper"] <- "Epinephelus coioides"
# cat_totals$Species[cat_totals$Species == "Red drum"] <- "Sciaenops ocellatus"
# cat_totals$Species[cat_totals$Species == "Blood cockle"] <- "Tegillarca granosa"
# cat_totals$Species[cat_totals$Species == "Blue mussel" ] <- "Mytilus edulis"
# cat_totals$Species[cat_totals$Species == "Chilean mussel"] <- "Mytilus chilensis"
# cat_totals$Species[cat_totals$Species == "Chinese mitten crab" ] <- "Eriocheir sinensis"
# cat_totals$Species[cat_totals$Species == "Constricted tagelus"] <- "Sinonovacula constricta"
# cat_totals$Species[cat_totals$Species == "Elkhorn sea moss"] <- "Kappaphycus alvarezii"
# 
# cat_totals$Species[cat_totals$Species == "Eucheuma seaweeds nei"] <- "Eucheuma seaweeds"
# cat_totals$Species[cat_totals$Species == "Fusiform sargassum"] <- "Sargassum fusiforme"
# cat_totals$Species[cat_totals$Species == "Giant river prawn"] <- "Macrobrachium rosenbergii"
# cat_totals$Species[cat_totals$Species == "Giant tiger prawn"] <- "Penaeus monodon"
# 
# # cat_totals$Species[cat_totals$Species == "Gracilaria seaweeds"] <- "Gracilaria gracilis" 
# cat_totals$Species[cat_totals$Species == "Green mud crab"] <- "Scylla paramamosain" 
# cat_totals$Species[cat_totals$Species == "Indo-Pacific swamp crab"] <- "Scylla serrata"
# cat_totals$Species[cat_totals$Species == "Japanese carpet shell"] <- "Ruditapes philippinarum"
# cat_totals$Species[cat_totals$Species == "Japanese kelp"] <- "Laminaria japonica"
# 
# cat_totals$Species[cat_totals$Species == "Laver (Nori)"] <- "Porphyra"
# cat_totals$Species[cat_totals$Species == "Mediterranean mussel"] <- "Mytilus galloprovincialis"
# cat_totals$Species[cat_totals$Species == "New Zealand mussel"] <- "Perna canaliculus"
# 
# cat_totals$Species[cat_totals$Species == "Oriental river prawn" ] <- "Macrobrachium nipponense"
# cat_totals$Species[cat_totals$Species == "Pacific cupped oyster"] <- "Magallana gigas"
# cat_totals$Species[cat_totals$Species == "Red swamp crawfish"] <- "Procambarus clarkii"
# cat_totals$Species[cat_totals$Species == "Spiny eucheuma"] <- "Eucheuma denticulatum"
# cat_totals$Species[cat_totals$Species == "Wakame"] <- "Undaria pinnatifida"
# cat_totals$Species[cat_totals$Species == "Whiteleg shrimp"] <- "Penaeus vannamei"
# cat_totals$Species[cat_totals$Species == "Yesso scallop" ] <- "Mizuhopecten yessoensis"
# 
# #extra 2 porphyra in species data, create triplicate
# cat_totals <- rbind(cat_totals, cat_totals[rep(45, 2), ])
# 
# #put rows in alphabetical order
# cat_totals <- cat_totals[order(cat_totals$Species), ]
# 
# #renaming in cat_totals so cat_totals matches spp_list and they can be merged
# cat_totals$Species[6] <- "Ctenopharyngodon idellus"
# 
# cat_totals$Species[42] <- "Porphrya tenera"
# cat_totals$Species[43] <- "Porphyra haitanensis"
# cat_totals$Species[44] <- "Porphyra yezoensis"
# 
# cat_totals$Species[13] <- "Gracilaria gracilis"
# 
# #remove E. denticulatum, as it is represented in Eucheuma seaweeds
# cat_totals <- cat_totals[!cat_totals$Species == "Eucheuma denticulatum",]
# 
# # #save this data
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prodbytaxa_alphab.csv")
# write.csv(cat_totals,file=NAME)
```

```{r }
#check
# v <- unique(nonfish$ASFIS.species.Name.En) #there should be 29 species
# 
# v <- unique(cat_totals$Species) #there should be 56 species
```

```{r Merging the datasets}
# #load in the FCB scores data
# speciesdata <- read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/FCB_scores_2024.csv") #updated8_2024
# 
# #alphabetical order:
# speciesdata <- speciesdata[order(speciesdata$species), ]
# 
# #rename (inland and marine) for nile tilapia and rainbow trout
# speciesdata$species[34] <- "Oncorhynchus mykiss (inland)"
# speciesdata$species[35] <- "Oncorhynchus mykiss (marine)"
# speciesdata$species[36] <- "Oreochromis niloticus (inland)"
# speciesdata$species[37] <- "Oreochromis niloticus (marine)"
# 
# #rename Eucheuma denticulatum to Eucheuma seaweeds (this entry represents the category)
# speciesdata$species[12] <- "Eucheuma seaweeds"
# 
# #delete Eucheuma cottonii, as it is already represented by K. alvarezii
# speciesdata <- speciesdata[!speciesdata$species == "Eucheuma cottonii",]
# 
# #MERGE datasets
# #create species and type from speciesdata, bind with category totals by species
# 
# spplist <- speciesdata[, c("species", "category")]
# names(cat_totals)[1] <- "species" #check this bc u use "Species" later on...
# 
# production_data <- merge(spplist, cat_totals, by="species")
# 
# #Save this dataset
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prodbytaxa_clean.csv")
# write.csv(production_data,file=NAME)
```

###STEP 3: Graphing production ###

```{r Graph production of each species, split into graphs based on taxa}

# setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Figures")
# 
# #load in data (don't need to run steps 1-2)
# production_data <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prodbytaxa_clean.csv")
# 
# #first up, the finfish:
# finfish_prod <- production_data[(production_data$category == "finfish_marine" | production_data$category == "finfish_inland"),]
# 
# # top three species are inland species, next three are marine
# finfish_prod <- production_data[(production_data$species == "Ctenopharyngodon idellus" | production_data$species == "Hypophthalmichthys molitrix" | production_data$species == "Oreochromis niloticus (inland)" | production_data$species == "Salmo salar" | production_data$species == "Chanos chanos" | production_data$species == "Mugil cephalus"),]
# 
# #make long
# finfish_prod <- pivot_longer(finfish_prod, cols = -c("species", "category"), 
#                                 names_to = "Year", values_to = "Production")
# 
# finfish_prod <- finfish_prod[order(finfish_prod$Year),] #put earlier years first
# 
# vec <- c(1950 : 2021)
# # replicate each integer in vec 3 times
# year_labs <- rep(vec, each = 6) #6 species here
# 
# #rename year column in weighted_indexes
# for (i in 1:nrow(finfish_prod)) {
#   #replace each year with the actual year number from year_labs
#   
#   finfish_prod$Year[i] <- year_labs[i]
# }
# 
# finfish_prod$Year <- as.numeric(finfish_prod$Year) #make years numeric
# 
# #graph
# plot <- ggplot(finfish_prod, aes(x=Year, y=Production, color = species)) +geom_line(size =1)+
#   labs(y = "Tons (Live Weight)", title = "Production of Major Finfish Species Over Time")+
#     theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))
# 
# # #SAVE in the figures folder
# ggsave("prod_finfish.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
# --------------------------------------------------------------------------------
# #algae:
# alg_prod <- production_data[(production_data$category == "algae"),]
# 
# #just the top-produced species
# alg_prod <- production_data[(production_data$species == "Laminaria japonica" | production_data$species == "Eucheuma seaweeds" | production_data$species == "Gracilaria gracilis" | production_data$species == "Undaria pinnatifida"),]
# 
# #rename 
# 
# #make long
# alg_prod <- pivot_longer(alg_prod, cols = -c("species", "category"), 
#                                 names_to = "Year", values_to = "Production")
# 
# alg_prod <- alg_prod[order(alg_prod$Year),] #put earlier years first
# 
# vec <- c(1950 : 2021)
# # replicate each integer in vec 3 times
# year_labs <- rep(vec, each = 4) #4 species here
# 
# #rename year column in weighted_indexes
# for (i in 1:nrow(alg_prod)) {
#   #replace each year with the actual year number from year_labs
#   
#   alg_prod$Year[i] <- year_labs[i]
# }
# 
# alg_prod$Year <- as.numeric(alg_prod$Year) #make years numeric
# 
# 
# 
# #graph
# plot <- ggplot(alg_prod, aes(x=Year, y=Production, color = species)) +geom_line(size =1)+
#   labs(y = "Tons (Live Weight)", title = "Production of Major Algae Species Over Time")+
#     theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))
# 
# # #SAVE
# ggsave("prod_algae.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
# --------------------------------------------------------------------------------
# #molluscs:
# moll_prod <- production_data[(production_data$category == "mollusc"),]
# 
# #just the top-produced species
# moll_prod <- production_data[(production_data$species == "Crassostrea virginica" | production_data$species == "Ruditapes philippinarum" | production_data$species == "Magallana gigas" | production_data$species == "Mizuhopecten yessoensis"),]
# 
# #make long
# moll_prod <- pivot_longer(moll_prod, cols = -c("species", "category"), 
#                                 names_to = "Year", values_to = "Production")
# 
# moll_prod <- moll_prod[order(moll_prod$Year),] #put earlier years first
# 
# vec <- c(1950 : 2021)
# # replicate each integer in vec 3 times
# year_labs <- rep(vec, each = 4) #4 species here
# 
# #rename year column in weighted_indexes
# for (i in 1:nrow(moll_prod)) {
#   #replace each year with the actual year number from year_labs
#   
#   moll_prod$Year[i] <- year_labs[i]
# }
# 
# moll_prod$Year <- as.numeric(moll_prod$Year) #make years numeric
# 
# #graph
# plot <- ggplot(moll_prod, aes(x=Year, y=Production, color = species)) +geom_line(size =1)+
#   labs(y = "Tons (Live Weight)", title = "Production of Major Mollusc Species Over Time")+
#     theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))
# 
# # #SAVE
# ggsave("prod_mollusc.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
# -------------------------------------------------------------------------------
# 
# #crustaceans:
# crust_prod <- production_data[(production_data$category == "crustacean"),]
# 
# #just the top-produced species
# crust_prod <- production_data[(production_data$species == "Penaeus vannamei" | production_data$species == "Procambarus clarkii" | production_data$species == "Eriocheir sinensis"),]
# 
# #make long
# crust_prod <- pivot_longer(crust_prod, cols = -c("species", "category"), 
#                                 names_to = "Year", values_to = "Production")
# 
# crust_prod <- crust_prod[order(crust_prod$Year),] #put earlier years first
# 
# vec <- c(1950 : 2021)
# # replicate each integer in vec 3 times
# year_labs <- rep(vec, each = 3) #3 species here
# 
# #rename year column in weighted_indexes
# for (i in 1:nrow(crust_prod)) {
#   #replace each year with the actual year number from year_labs
#   
#   crust_prod$Year[i] <- year_labs[i]
# }
# 
# crust_prod$Year <- as.numeric(crust_prod$Year) #make years numeric
# 
# #graph
# plot <- ggplot(crust_prod, aes(x=Year, y=Production, color = species)) +geom_line(size =1)+
#   labs(y = "Tons (Live Weight)", title = "Production of Major Crustacean Species Over Time")+
#     theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))
# 
# # #SAVE
# ggsave("prod_crust.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

###STEP 4: CALCULATING PRODUCTION BY ALL, then by INVERTS, then PLANTS, then FISH. We will use these production values to weight FCB scores by taxonomic category. Save csv, "prodbytaxa_forweighting.csv"####
```{r Create a dataframe, 'weighting totals', which holds production values by taxonomic group}

# #load in data (don't need to run steps 1-3 to run this section)
# production_data <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prodbytaxa_clean.csv")
# cat_totals <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prodbytaxa_alphab.csv")
# 
# #remove extra columns
# production_data <- production_data[, c(2:75)]
# cat_totals <- cat_totals[, c(2:74)]
# 
# frame <- production_data[, c(2:74)]
# #c(1, 3:74)]
# 
# weighting_totals <- frame #create new frame
# weighting_totals <- weighting_totals[FALSE, ]
# 
# #use "cat_totals" for total of all spp
# for (i in 2:ncol(cat_totals)){ #for each year
#   # print(production_data[1, i])
#   #sum the columns into a new row
#   a <- sum(cat_totals[, i], na.rm = TRUE)
#   print(a)
#   weighting_totals[1, i] <- a
#   
# }
# #rename
# weighting_totals[1, 1] <- "all species" 
# 
# #Calculate finfish totals
# #filter out just the finfish from "production_data" dataframe
# production_data2 <- filter(production_data, category =="finfish_inland"| category =="finfish_marine")
# 
# #remove the category column so that this dataframe and the weighting_totals dataframe line up
# production_data2 <- production_data2[-c(1)]
# 
# for (i in 2:ncol(production_data2)){ #for each year
#   #sum the columns into a new row
#   a <- sum(production_data2[, i], na.rm = TRUE)
#   print(a)
#   weighting_totals[2, i] <- a #second row for this total
#   
# }
# 
# weighting_totals[2, 1] <- "finfish" #rename
# -------------------------------------------------------------------------------
# 
# #Calculate invert totals
# #filter out just the inverts
#   
# # production_data2 <- filter(production_data, category =="mollusc"| category =="crustacean")
# production_data2 <- production_data[(production_data$category == "crustacean" | production_data$category == "mollusc"),]
# 
# #remove the category column so that this dataframe and the weighting_totals dataframe line up
# production_data2 <- production_data2[-c(2)]
# 
# for (i in 2:ncol(production_data2)){ #for each year
#   #sum the columns into a new row
#   a <- sum(production_data2[, i], na.rm = TRUE)
#   print(a)
#   weighting_totals[3, i] <- a #third row for the inverts total
#   
# }
# 
# weighting_totals[3, 1] <- "inverts" #rename
# -------------------------------------------------------------------------------
#   
# #Calculate algae totals
# #filter out just the algae
# # production_data2 <- filter(production_data, category =="algae")
# production_data2 <- production_data[(production_data$category == "algae"),]
# 
# #remove the category column so that this dataframe and the weighting_totals dataframe line up
# production_data2 <- production_data2[-c(2)]
# 
# for (i in 2:ncol(production_data2)){ #for each year
#   #sum the columns into a new row
#   a <- sum(production_data2[, i], na.rm = TRUE)
#   print(a)
#   weighting_totals[4, i] <- a #fourth row for the algae total
#   
# }
# 
# weighting_totals[4, 1] <- "algae" #rename
# -------------------------------------------------------------------------------
#   
# #Calculate crustacean totals
# #filter out just the crustaceans
# # production_data2 <- filter(production_data, category =="crustacean")
# production_data2 <- production_data[(production_data$category == "crustacean"),]
# 
# #remove the category column so that this dataframe and the weighting_totals dataframe line up
# production_data2 <- production_data2[-c(2)]
# 
# for (i in 2:ncol(production_data2)){ #for each year
#   #sum the columns into a new row
#   a <- sum(production_data2[, i], na.rm = TRUE)
#   print(a)
#   weighting_totals[5, i] <- a #fifth row for the crustacean total
#   
# }
# 
# weighting_totals[5, 1] <- "crustacean" #rename
# -------------------------------------------------------------------------------
#   
# #Calculate mollusc totals
# #filter out just the molluscs
# # production_data2 <- filter(production_data, category =="mollusc")
# production_data2 <- production_data[(production_data$category == "mollusc"),]
# 
# #remove the category column so that this dataframe and the weighting_totals dataframe line up
# production_data2 <- production_data2[-c(2)]
# 
# for (i in 2:ncol(production_data2)){ #for each year
#   #sum the columns into a new row
#   a <- sum(production_data2[, i], na.rm = TRUE)
#   print(a)
#   weighting_totals[6, i] <- a #sixth row for the mollusc total
#   
# }
# 
# weighting_totals[6, 1] <- "mollusc" #rename
# 
# #SAVE
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prodbytaxa_forweighting.csv")
# write.csv(weighting_totals,file=NAME)
```

```{r Graph production over time by taxa}

# #load in data (don't need to run steps 1-4 to run this section)
# weighting_totals <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prodbytaxa_forweighting.csv")
# 
# #remove extra columns
# weighting_totals <- weighting_totals[, c(2:74)]
# 
# library(tidyr)
# #make long
# weighting_totals2 <- pivot_longer(weighting_totals, cols = -c(category), 
#                                 names_to = "Year", values_to = "Score")
# 
# year_labs <- rep(2021:1950, times=6) 
# 
# #rename year column in weighted_indexes
# for (i in 1:nrow(weighting_totals2)) {
#   print(weighting_totals2$Year[i])
#   #replace each year with the actual year number from year_labs
#   
#   weighting_totals2$Year[i] <- year_labs[i]
# }
# 
# weighting_totals2$Year <- as.numeric(weighting_totals2$Year) #make years numeric
# weighting_totals2 <- weighting_totals2[order(weighting_totals2$Year),] #put earlier years first
# 
# #remove inverts
# weighting_totals2 <- weighting_totals2[!(weighting_totals2$category == "inverts"),]
# 
# #graph
# plot <- ggplot(weighting_totals2, aes(x=Year, y=Score, color = category)) +geom_line(size =1)+
#   labs(y = "Tons (Live Weight)", title = "Aquaculture Production over Time")+
#     theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
#       scale_color_manual(name = "Taxa", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))

# #SAVE
# ggsave("prod_over_time_bytaxa.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set

# model <- lm(Score ~ Year, data = weighting_totals2)
# summary(model)

#calculate growth rates?
```

