---
title: "Spatialdata_prep"
author: "Aleah Wong"
date: '2024-09-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
.rs.restartR()
```

```{r}
library(dplyr)
library(tidyr)
```

############################################################################################
##TABLE OF CONTENTS
STEP 1: Rename species in the production dataset to match latin names in the FCB scores dataset (Line 30)

STEP 2: Calculate annual production totals by species and by continent (Line 135)

STEP 3: Weight FCB scores by production data

STEP 4:
############################################################################################

###STEP 1: Rename species ####
```{r Load in FAO data}

inland <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/inland_finfish_prod.csv")
marine <-read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/marine_finfish_prod.csv")
nonfish <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/nonfish_prod.csv")
nigeria <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Nigeria_prod.csv") #production specific to Nigeria, not already included

#delete 2022 from Nigeria, since the other datasets are missing this year
nigeria <- nigeria[, c(-6)]

####Rowbind
all_production <- bind_rows(inland, marine, nonfish, nigeria)

#Remove nori nei, scallops nei and sea mussels nei
all_production <- all_production[!all_production$ASFIS.species.Name.En == "Nori nei",]
all_production <- all_production[!all_production$ASFIS.species.Name.En == "Scallops nei",]
all_production <- all_production[!all_production$ASFIS.species.Name.En == "Sea mussels nei",]
all_production <- all_production[!all_production$ASFIS.species.Name.En == "Eucheuma seaweeds nei", ]

###Rename species names as latin names so they match data file containing FCB scores data
unique(all_production$ASFIS.species.Name.En) #get a list of unique species in all_production

#Specific to Nigeria:
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Cyprinids nei"] <- "Cyprinus carpio" 
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Penaeus shrimps nei"] <- "Penaeus vannamei"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Tilapias nei"] <- "Oreochromis niloticus"

all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Bighead carp"] <- "Hypophthalmichthys nobilis"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Black carp"] <- "Mylopharyngodon piceus"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En =="Catla"] <- "Catla catla"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En =="Common carp"] <- "Cyprinus carpio"                  
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En =="Crucian carp"] <- "Carassius carassius"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En =="Grass carp(=White amur)"] <- "Ctenopharyngodon idellus" #FIX ??
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Largemouth black bass"] <- "Micropterus salmoides"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En =="Nile tilapia"] <- "Oreochromis niloticus"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Rainbow trout"] <- "Oncorhynchus mykiss"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Roho labeo"] <- "Labeo rohita"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Silver carp"] <- "Hypophthalmichthys molitrix"                
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Striped catfish"] <- "Pangasianodon hypophthalmus"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Wuchang bream"] <- "Megalobrama amblycephala"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "North African catfish"] <- "Clarias gariepinus"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Atlantic salmon"] <- "Salmo salar"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Barramundi(=Giant seaperch)"] <- "Lates calcarifer"                                      
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Coho(=Silver) salmon"] <- "Oncorhynchus kisutch"                                
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "European seabass"] <- "Dicentrarchus labrax"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Flathead grey mullet"] <- "Mugil cephalus"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Gilthead seabream"] <- "Sparus aurata"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Japanese amberjack"] <- "Seriola quinqueradiata"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Japanese seabass"] <-  "Lateolabrax japonicus"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Large yellow croaker"] <- "Larimichthys croceus"                                       
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Milkfish"] <- "Chanos chanos"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Orange-spotted grouper"] <- "Epinephelus coioides"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Pompano"] <- "Trachinotus ovatus"                                       
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Red drum"] <- "Sciaenops ocellatus"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Blood cockle"] <- "Tegillarca granosa"                                       
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Blue mussel" ] <- "Mytilus edulis"                                       
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Chilean mussel"] <- "Mytilus chilensis"                                      
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Chinese mitten crab" ] <- "Eriocheir sinensis"                                      
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Constricted tagelus"] <- "Sinonovacula constricta"                                       
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Elkhorn sea moss"] <- "Kappaphycus alvarezii"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Fusiform sargassum"] <- "Sargassum fusiforme"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Giant river prawn"] <- "Macrobrachium rosenbergii"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Giant tiger prawn"] <- "Penaeus monodon"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Gracilaria seaweeds"] <- "Gracilaria gracilis"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Green mud crab"] <- "Scylla paramamosain"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Indo-Pacific swamp crab"] <- "Scylla serrata"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Japanese carpet shell"] <- "Ruditapes philippinarum"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Japanese kelp"] <- "Laminaria japonica"                                       
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Laver (Nori)"] <- "Porphyra yezoensis" #We will use yezoensis as a proxy for the other three species of poryphyra                                       
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Mediterranean mussel"] <- "Mytilus galloprovincialis"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "New Zealand mussel"] <- "Perna canaliculus"                                       
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Oriental river prawn" ] <- "Macrobrachium nipponense"                                        
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Pacific cupped oyster"] <- "Magallana gigas" 
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Red swamp crawfish"] <- "Procambarus clarkii"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Spiny eucheuma"] <- "Eucheuma denticulatum"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Wakame"] <- "Undaria pinnatifida"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Whiteleg shrimp"] <- "Penaeus vannamei"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "Yesso scallop" ] <- "Mizuhopecten yessoensis"
all_production$ASFIS.species.Name.En[all_production$ASFIS.species.Name.En == "American cupped oyster"] <- "Crassostrea virginica"

#REMEMBER: only weight Porphyra yezoensis (out of three species of Porphyra included in the FCB scores data), since the production data clumps them all together

#Put rows in alphabetical order
all_production <- all_production[order(all_production$ASFIS.species.Name.En), ]

#save
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prod_by_cat910.csv")
# write.csv(all_production,file=NAME)
```

```{r Create data subsets: continent data without China; just China data; datasets with select country data}

# all_production2 <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prod_by_cat910.csv") #can load in the new data if you don't run chunk above
# all_production2 <- all_production2[, c(-1)] #remove extra row

#Remove China from datasets so we can calculate Asian total production excluding China:
all_production_withoutCH <-  subset(all_production, all_production$Country.Name.En != "China")

#Just China
just_china_prod <- subset(all_production, all_production$Country.Name.En == "China")

#All country case studies: China, Norway, Chile and Nigeria #could probably just filter these from all_production during calculations 
country_case_study_prod <- all_production[all_production$Country.Name.En %in% c("China", "Norway", "Chile", "Nigeria"), ]
```

###STEP 2: Calculate annual production totals: just China, all continents excluding China, country-specific####
```{r Just China's production}

just_china_prod <- just_china_prod %>%mutate_at(c(6:77), as.numeric) #change yearly data to numeric

just_china_totals <- filter(just_china_prod) #create new frame
just_china_totals <- just_china_totals[FALSE, ]

  #find sum for each species, each year
  u_species <- unique(just_china_prod$ASFIS.species.Name.En)

  for (z in 1:length(u_species)) {
    # print(u_species[z])
    uu <- filter(just_china_prod, ASFIS.species.Name.En == u_species[z], )

      for (b in 6:77){ #for each year; calculate the annual mean each year
        a <- sum(uu[, b], na.rm = TRUE)

        just_china_totals[z, b] <-a

        just_china_totals[z, 1] <- u_species[z] #species name
        just_china_totals[z, 2] <- "China" #continent

    }
  }

just_china_totals <- just_china_totals[, c(1, 2, 6:77)] #cut out unnecessary columns
names(just_china_totals)[1] <- "Species" #rename first column

# #save this data
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/CH_prod_totals.csv")
# write.csv(just_china_totals,file=NAME)
```

```{r Species annual totals, without China}

all_production_withoutCH <- all_production_withoutCH %>%mutate_at(c(6:77), as.numeric) #change yearly data to numeric

#create list of each continent to loop through
uniquecontinent <- c("Africa", "Americas", "Asia", "Europe", "Oceania")

prod_without_CH_totals <- filter(all_production_withoutCH) #create new frame
prod_without_CH_totals <- prod_without_CH_totals[FALSE, ]

# Sum the annual production of each species for each continent
for (i in 1:5){ #for each of the 5 continents
  u_continent <- filter(all_production_withoutCH, Continent.Name.En == uniquecontinent[i], ) #subset one continent
  # print(u_continent$Continent.Name.En[i])
  y <- nrow(prod_without_CH_totals)

  #find sum for each species, each year
  u_species <- unique(u_continent$ASFIS.species.Name.En)

  for (z in 1:length(u_species)) {
    # print(u_species[z])
    uu <- filter(u_continent, ASFIS.species.Name.En == u_species[z], )

      for (b in 6:77){ #for each year; calculate the annual mean each year
        a <- sum(uu[, b], na.rm = TRUE)

        prod_without_CH_totals[y+z, b] <-a

        prod_without_CH_totals[y+z, 1] <- u_species[z] #species name
        prod_without_CH_totals[y+z, 2] <- uniquecontinent[i] #continent

    }
  }
}

prod_without_CH_totals <- prod_without_CH_totals[, c(1, 2, 6:77)] #cut out unnecessary columns
names(prod_without_CH_totals)[1] <- "Species" #rename first column

# #save this data
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prod_totals_without_CH.csv")
# write.csv(prod_without_CH_totals,file=NAME)

cat_totals <- rbind(prod_without_CH_totals, just_china_totals) #combine China data with the rest (China as own "continent" category)

#save cat_totals
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prod_by_cat.csv")
# write.csv(cat_totals,file=NAME)
```
 
```{r Production totals by continent}
# 
# cat_totals <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prod_by_cat.csv")
# cat_totals <- cat_totals[, c(-1)]

# str(cat_totals)
# cat_totals <- cat_totals %>%mutate_at(c(3:74), as.numeric) #change yearly data to numeric if not already

names(cat_totals)[2] <- "Continent" #rename continent column

new_frame <- cat_totals #create new frame
new_frame <- new_frame[FALSE, ]

#now add the totals for other categories
#create list of each continent to loop through
unique <- c("Africa", "Americas", "Asia", "Europe", "Oceania", "China") 


#WORKING HERE

for (i in 1:length(unique)){ #for each of the continents
  u_cont <- filter(cat_totals, Continent == unique[i])
  
#calculate the annual prod each year

  for (b in 3:74){
    a <- sum(u_cont[, b], na.rm = TRUE)
    print(a)
    new_frame[i, 2] <- unique[i]
    new_frame[i, b] <-a #change row number 
  }
}

new_frame <- new_frame[, c(2:74)] #cut out unnecessary columns

#save this data
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/category_totals_bycontinent.csv")
# write.csv(new_frame,file=NAME)
```

####STEP 3: Weight FCB scores by production data --> Can start here since all data are saved
```{r}
#load in the FCB scores data
speciesdata <- read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/FCB_scores_825.2.csv")
speciesdata <- speciesdata[, -c(1)]
speciesdata <- speciesdata[-c(4, 6, 8, 34, 36),] #delete duplicated nile tilapia and rainbow trout; delete Porphyra haitanensis and tenera because we will only use Porphyra yezoensis score

#load in the totals
totals <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/category_totals_bycontinent.csv")
totals <- totals[, c(-1)]

cat_totals <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prod_by_cat.csv")
cat_totals <- cat_totals[, c(-1)] #need to delete first column for this one only
```

###F scores weighted
```{r F weighted by all production}

# unique(cat_totals$Species) #52

new_frame <- cat_totals #create new frame
new_frame <- new_frame[FALSE, ]

for (i in 1:nrow(speciesdata)){ #for each species
  y <- nrow(new_frame) #reset nrow
  sp <- speciesdata$species[i] #get one species
  # print(sp)
  f_score <- speciesdata$f[i] #get the F score for the species, CHANGE FOR B AND C
  set <- filter(cat_totals, Species == sp, )
  # print(set$Species)
  # print(f_score)
     for (z in 1:nrow(set)) { #for each row in the species set (for each continent)
    for (b in 3:74) { #for each year
      index <- f_score*set[z, b] #multiply score by that species' total production in each year
      print(index)
      new_frame[y+z, b] <- index   #add these numbers to the frame
      new_frame[y+z, 1] <- speciesdata[i, 1] #add species
      new_frame[y+z, 2] <- set$Continent.Name.En[z] #add continent
      
            if(is.na(set$Continent.Name.En[z])) {x=FALSE
    } else if (set$Continent.Name.En[z] == "Africa") {
    index <- index/totals[1, b-1]
    } else if (set$Continent.Name.En[z] == "Americas") {
    index <- index/totals[2, b-1]
    } else if (set$Continent.Name.En[z] == "Asia") {
    index <- index/totals[3, b-1]
    } else if (set$Continent.Name.En[z] == "Europe") {
    index <- index/totals[4, b-1]
    } else if (set$Continent.Name.En[z] == "Oceania") {
    index <- index/totals[5, b-1]
    } else
    index <- index/totals[6, b-1]

      new_frame[y+z, b] <- index   #add these numbers to the frame
    }
     }
}

###WORKING HERE (see line 435 in production_analysis_without_china)
###Calculate the annual weighted mean for each continent

weighted_indexes <- new_frame #create new frame
weighted_indexes <- weighted_indexes[FALSE, ]

#select one continent at a time, then sum all values for each year
#create list of each continent to loop through
unique <- c("Africa", "Americas", "Asia", "Europe", "Oceania", "China") 

for (i in 1:length(unique)){ #for each of the continents
  u_cont <- filter(new_frame, Continent.Name.En == unique[i], )
  y <- nrow(new_frame) #reset nrow
  
#calculate the annual mean each year
  for (b in 3:74){
    a <- sum(u_cont[, b], na.rm = TRUE)
    print(a)
    
    #how to store?
    weighted_indexes[i, b] <- a   #add these numbers to the frame
    weighted_indexes[i, 2] <- unique[i] #add continent
  }
}

weighted_indexes <- weighted_indexes[, 2:74] #new frame with just the weighted scores each year

```

###WORKING HERE (see line 435 in production_analysis_without_china)
```{r}

```


```{r F weighted by all production}

weighted_indexes <- new_frame #create new frame
weighted_indexes <- weighted_indexes[FALSE, ]

#select one continent at a time, then sum all values for each year
#create list of each continent to loop through
unique <- c("Africa", "Americas", "Asia", "Europe", "Oceania", "China") 

for (i in 1:length(unique)){ #for each of the continents
  u_cont <- filter(new_frame, Continent.Name.En == unique[i], )
  y <- nrow(new_frame) #reset nrow
  
#calculate the annual mean each year
  for (b in 3:74){
    a <- sum(u_cont[, b], na.rm = TRUE)
    print(a)
    
    #how to store?
    weighted_indexes[i, b] <- a   #add these numbers to the frame
    weighted_indexes[i, 2] <- unique[i] #add continent
  }
}

weighted_indexes <- weighted_indexes[, 2:74] #new frame with just the weighted scores each year

################## Make long to prepare for graphing ###########################
weighted_indexes <- pivot_longer(weighted_indexes, cols = c(!Continent.Name.En),
                                names_to = "Year", values_to = "Score")


year_labs <- rep(2021:1950, times=6, each=1)
print(year_labs)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  #replace each year with the actual year number from year_labs
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#Unnecessary:
# weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first
#make alphabetical so all years for one continent are together
# weighted_indexes <- weighted_indexes[order(weighted_indexes$Continent.Name.En), ]
```











#REMEMBER: only weight by one species of Porphyra (out of three species of Porphyra included in the FCB scores data), since the production data clumps them all together


####COUNTRY DATA

CHINA and CANADA ###FIX this to include Nigeria, Chile and Norway
```{r}
#filter out China and Canada from the inland, marine and nonfish FAO data

inland_china <- inland[inland$Country.Name.En %in% c("China", "Canada"),]

marine_china <- marine[marine$Country.Name.En %in% c("China", "Canada"),]

nonfish_china <- nonfish[nonfish$Country.Name.En %in% c("China", "Canada"),]
```

```{r}
#calculate annual production totals for each inland species, for each country
#only need to do this for nonfish, since the yearly totals are already reflected in the datasets for marine, inland (no duplicates)?????

library(dplyr)

nonfish_china <- nonfish_china %>%mutate_at(c(6:77), as.numeric) #change yearly data to numeric

#create list of each country to loop through
uniquecontinent <- c("Canada", "China") 

new_frame <- filter(nonfish_china) #create new frame
new_frame <- new_frame[FALSE, ]

#TEST
for (i in 1:2){ #for each of the 2 countries
  u_continent <- filter(nonfish_china, Country.Name.En == uniquecontinent[i], ) #subset one country at a time
  # print(u_continent$Continent.Name.En[i])
  y <- nrow(new_frame)
  
  #find sum for each species, each year
  u_species <- unique(u_continent$ASFIS.species.Name.En)
  
  for (z in 1:length(u_species)) {
    # print(u_species[z])
    uu <- filter(u_continent, ASFIS.species.Name.En == u_species[z], )
  
      for (b in 6:77){ #for each year; calculate the annual mean each year
        a <- sum(uu[, b], na.rm = TRUE)
        
        new_frame[y+z, b] <-a
        
        new_frame[y+z, 1] <- u_species[z] #species name
        new_frame[y+z, 3] <- uniquecontinent[i] #continent
      
    }
  }
}

new_frame <- new_frame[, c(1, 3, 6:77)] #cut out unnecessary columns
names(new_frame)[1] <- "Species" #rename first column

# #save this data
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/nonfish_totals_CH_CA.csv")
# write.csv(new_frame,file=NAME)
```

Calculating weighted index scores (FIX)
```{r}
#load data
inland_totals_cont <- inland
inland_totals_cont <- inland_totals_cont[, c(1, 3, 6:77)] #cut out unnecessary columns
names(inland_totals_cont)[1] <- "Species" #rename first column

marine_totals_cont <- marine
marine_totals_cont <- marine_totals_cont[, c(1, 3, 6:77)] #cut out unnecessary columns
names(marine_totals_cont)[1] <- "Species" #rename first column

nonfish_totals_cont <- read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/nonfish_totals_CH_CA.csv")
nonfish_totals_cont <- nonfish_totals_cont[-1] #get rid of first column

#rename (inland and marine) for nile tilapia and rainbow trout
inland_totals_cont$Species[inland_totals_cont$Species == "Rainbow trout"] <- "Oncorhynchus mykiss (inland)"
inland_totals_cont$Species[inland_totals_cont$Species == "Nile tilapia"] <- "Oreochromis niloticus (inland)"

#rename (inland and marine) for nile tilapia and rainbow trout
marine_totals_cont$Species[marine_totals_cont$Species == "Rainbow trout"] <- "Oncorhynchus mykiss (marine)"
marine_totals_cont$Species[marine_totals_cont$Species == "Nile tilapia"] <- "Oreochromis niloticus (marine)"


#combine the datasets with rowbind
cat_totals <- rbind(inland_totals_cont, marine_totals_cont)
cat_totals <- rbind(cat_totals, nonfish_totals_cont)

#rename pompano, get rid of nori nei, scallops nei and sea mussels nei
cat_totals <- cat_totals[!cat_totals$Species == "Nori nei",]
cat_totals <- cat_totals[!cat_totals$Species == "Scallops nei",]
cat_totals <- cat_totals[!cat_totals$Species == "Sea mussels nei",]
```

```{r}
#put both datasets into the same order so the species line up
#rename the cat_totals species to be scientific names

cat_totals$Species[cat_totals$Species == "American cupped oyster"] <- "Crassostrea virginica"
cat_totals$Species[cat_totals$Species == "North African catfish"] <- "Clarias gariepinus"
cat_totals$Species[cat_totals$Species == "Pompano"] <- "Trachinotus ovatus"

cat_totals$Species[cat_totals$Species == "Black carp"] <- "Mylopharyngodon piceus"
cat_totals$Species[cat_totals$Species == "Bighead carp"] <- "Hypophthalmichthys nobilis"
cat_totals$Species[cat_totals$Species == "Catla"] <- "Catla catla"
cat_totals$Species[cat_totals$Species == "Common carp"] <- "Cyprinus carpio"
cat_totals$Species[cat_totals$Species == "Crucian carp"] <- "Carassius carassius"
cat_totals$Species[cat_totals$Species == "Grass carp(=White amur)"] <- "Ctenopharyngodon idellus"
cat_totals$Species[cat_totals$Species == "Largemouth black bass"] <- "Micropterus salmoides"
cat_totals$Species[cat_totals$Species == "Nile tilapia"] <- "Oreochromis niloticus"
cat_totals$Species[cat_totals$Species == "Rainbow trout"] <- "Oncorhynchus mykiss"
cat_totals$Species[cat_totals$Species == "Roho labeo"] <- "Labeo rohita"
cat_totals$Species[cat_totals$Species == "Silver carp"] <- "Hypophthalmichthys molitrix"
cat_totals$Species[cat_totals$Species == "Striped catfish"] <- "Pangasianodon hypophthalmus"
cat_totals$Species[cat_totals$Species == "Wuchang bream"] <- "Megalobrama amblycephala"
cat_totals$Species[cat_totals$Species == "Atlantic salmon"] <- "Salmo salar"
cat_totals$Species[cat_totals$Species == "Barramundi(=Giant seaperch)"] <- "Lates calcarifer"
cat_totals$Species[cat_totals$Species == "Coho(=Silver) salmon"] <- "Oncorhynchus kisutch"
cat_totals$Species[cat_totals$Species == "European seabass"] <- "Dicentrarchus labrax"
cat_totals$Species[cat_totals$Species == "Flathead grey mullet"] <- "Mugil cephalus"
cat_totals$Species[cat_totals$Species == "Gilthead seabream"] <- "Sparus aurata"
cat_totals$Species[cat_totals$Species == "Japanese amberjack"] <- "Seriola quinqueradiata"
cat_totals$Species[cat_totals$Species == "Japanese seabass"] <-  "Lateolabrax japonicus"
cat_totals$Species[cat_totals$Species == "Large yellow croaker"] <- "Larimichthys croceus"
cat_totals$Species[cat_totals$Species == "Milkfish"] <- "Chanos chanos"
cat_totals$Species[cat_totals$Species == "Nile tilapia"] <- "Oreochromis niloticus"
cat_totals$Species[cat_totals$Species == "Orange-spotted grouper"] <- "Epinephelus coioides"
cat_totals$Species[cat_totals$Species == "Red drum"] <- "Sciaenops ocellatus"
cat_totals$Species[cat_totals$Species == "Blood cockle"] <- "Tegillarca granosa"
cat_totals$Species[cat_totals$Species == "Blue mussel" ] <- "Mytilus edulis"
cat_totals$Species[cat_totals$Species == "Chilean mussel"] <- "Mytilus chilensis"
cat_totals$Species[cat_totals$Species == "Chinese mitten crab" ] <- "Eriocheir sinensis"
cat_totals$Species[cat_totals$Species == "Constricted tagelus"] <- "Sinonovacula constricta"
cat_totals$Species[cat_totals$Species == "Elkhorn sea moss"] <- "Kappaphycus alvarezii"

cat_totals$Species[cat_totals$Species == "Eucheuma seaweeds nei"] <- "Eucheuma cottonii"
cat_totals$Species[cat_totals$Species == "Fusiform sargassum"] <- "Sargassum fusiforme"
cat_totals$Species[cat_totals$Species == "Giant river prawn"] <- "Macrobrachium rosenbergii"
cat_totals$Species[cat_totals$Species == "Giant tiger prawn"] <- "Penaeus monodon"

cat_totals$Species[cat_totals$Species == "Gracilaria seaweeds"] <- "Gracilaria gracilis"
cat_totals$Species[cat_totals$Species == "Green mud crab"] <- "Scylla paramamosain" 
cat_totals$Species[cat_totals$Species == "Indo-Pacific swamp crab"] <- "Scylla serrata"
cat_totals$Species[cat_totals$Species == "Japanese carpet shell"] <- "Ruditapes philippinarum"
cat_totals$Species[cat_totals$Species == "Japanese kelp"] <- "Laminaria japonica"

cat_totals$Species[cat_totals$Species == "Laver (Nori)"] <- "Porphyra"
cat_totals$Species[cat_totals$Species == "Mediterranean mussel"] <- "Mytilus galloprovincialis"
cat_totals$Species[cat_totals$Species == "New Zealand mussel"] <- "Perna canaliculus"

cat_totals$Species[cat_totals$Species == "Oriental river prawn" ] <- "Macrobrachium nipponense"
cat_totals$Species[cat_totals$Species == "Pacific cupped oyster"] <- "Magallana gigas"
cat_totals$Species[cat_totals$Species == "Red swamp crawfish"] <- "Procambarus clarkii"
cat_totals$Species[cat_totals$Species == "Spiny eucheuma"] <- "Eucheuma denticulatum"
cat_totals$Species[cat_totals$Species == "Wakame"] <- "Undaria pinnatifida"
cat_totals$Species[cat_totals$Species == "Whiteleg shrimp"] <- "Penaeus vannamei"
cat_totals$Species[cat_totals$Species == "Yesso scallop" ] <- "Mizuhopecten yessoensis"
```

```{r}
#put rows in alphabetical order
cat_totals <- cat_totals[order(cat_totals$Species), ]
```

PRODUCTION TOTALS by COUNTRY
```{r}
str(cat_totals)

cat_totals <- cat_totals %>%mutate_at(c(3:74), as.numeric) #change yearly data to numeric
names(cat_totals)[2] <- "Country" #rename continent column

new_frame <- cat_totals #create new frame
new_frame <- new_frame[FALSE, ]

#now add the totals for other categories
#create list of each continent to loop through
unique <- c("China", "Canada") 

for (i in 1:length(unique)){ #for each of the countries
  u_cont <- filter(cat_totals, Country == unique[i])
  
#calculate the annual mean each year

  for (b in 3:74){
    a <- sum(u_cont[, b], na.rm = TRUE)
    print(a)
    new_frame[i, 2] <- unique[i]
    new_frame[i, b] <-a #change row number 
  }
}

new_frame <- new_frame[, c(2, 3:74)] #cut out unnecessary columns

#save this data
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/category_totals_CH_CA.csv")
# write.csv(new_frame,file=NAME)
```


