---
title: "Spatial_analysis_clean"
author: "Aleah Wong"
date: '2024-09-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

############################################################################################
##TABLE OF CONTENTS--FIX
STEPS 1-4: See "Spatialdata_prep.Rmd"

STEP 5: Calculating 5-year FCB means for every region 

STEP 6: Calculate change in 5-year means

STEP 7: Heatmap of regional scores

STEP 8: Summary stats

STEP 9: Visualize difference in 5-year means with boxplots

STEP 10: Find theoretical FCB max (MFCB) for each region and calculate the percentage of MFCB that is equal to the difference in means

############################################################################################

```{r}
library(dplyr)
library(ggpubr)
```

```{r Load in the saved data}

F_weighted_continent <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Fweighted_bycontinent.csv")
F_weighted_continent <- F_weighted_continent[, -c(1)]
C_weighted_continent <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Cweighted_bycontinent.csv")
C_weighted_continent <- C_weighted_continent[, -c(1)]
B_weighted_continent <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Bweighted_bycontinent.csv")
B_weighted_continent <- B_weighted_continent[, -c(1)]

F_weighted_country <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Fweighted_bycountry.csv")
F_weighted_country <- F_weighted_country[, -c(1)]
C_weighted_country <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Cweighted_bycountry.csv")
C_weighted_country <- C_weighted_country[, -c(1)]
B_weighted_country <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Bweighted_bycountry.csv")
B_weighted_country <- B_weighted_country[, -c(1)]

```

```{r}
####### Separate FCB data for each country, Combine F_weighted_country with C and B ################

FCB_countries <- cbind(F_weighted_country, C_weighted_country[, c(3)])
FCB_countries <- cbind(FCB_countries, B_weighted_country[, c(3)])
names(FCB_countries)[3] <- "Food" #rename columns as F, C and B
names(FCB_countries)[4] <- "Climate"
names(FCB_countries)[5] <- "Biodiversity"

#make long
FCB_countries <- pivot_longer(FCB_countries, cols = c(Food, Climate, Biodiversity), names_to = "Index", values_to = "Score")

#subset
China <- subset(FCB_countries, Country =="China")
Chile <- subset(FCB_countries, Country =="Chile")
Norway <- subset(FCB_countries, Country =="Norway")
Nigeria <- subset(FCB_countries, Country =="Nigeria")
```

```{r}
#Color palette with colorblind friendly colors

cbPalette1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#use orange, yellow, dark blue, red and pink
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#D55E00", "#CC79A7")


# # To use for fills, add
#   scale_fill_manual(values=cbPalette)
# 
# # To use for line and point colors, add
#   scale_colour_manual(values=cbPalette)
```

### Preliminary visualization of FCB trends in case study countries
```{r FCB trends over time; one graph per country}

#Just China
ChinaFCB <- ggplot(China, aes(x=Year, y=Score, color=Index)) + geom_line(size=1) + labs(title = "FCB Potential of Aquaculture in China") +
  theme(axis.text.x = element_text(size = 10), axis.text.y = element_text(size = 10))+
  ylim(33, 70)

#FACETED
case_study_FCB_facet <- ggplot(FCB_countries, aes(Year, y=Score, color=Index))+
  geom_line(size=1)+
  facet_wrap(~Country)+
  xlim(1980, 2021)+
  ylim(33, 65)

#save these?
# setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Figures")
# ggsave("China_FCB.png", ChinaFCB, height = 4, width = 8, dpi = 600)
# ggsave("Case_study_FCB.png", case_study_FCB_facet, height = 4, width = 8, dpi = 600)
```

###STEP 5: Calculating 5-year FCB means for every region #######
```{r Organizing the data--get just the years we want}

#### FCB means of continents #########
now_then_meansV2 <- cbind(F_weighted_continent, C_weighted_continent[, c(3)])
now_then_meansV2 <- cbind(now_then_meansV2, B_weighted_continent[, c(3)])
names(now_then_meansV2)[3] <- "Food"
names(now_then_meansV2)[4] <- "Climate"
names(now_then_meansV2)[5] <- "Biodiversity"

now_then_meansV2 <- filter(now_then_meansV2, Year %in% c(1980, 1981, 1982, 1983, 1984, 1985, 2016, 2017, 2018, 2019, 2020, 2021))

##### F ########
now_then_meansF <- now_then_meansV2[, c(1:3)]
now_then_meansF <- now_then_meansF %>%
  pivot_wider(names_from = "Region", values_from = "Food")

now_then_meansF$Year[now_then_meansF$Year == 1980 |now_then_meansF$Year == 1981 |now_then_meansF$Year == 1982 |now_then_meansF$Year == 1983 |now_then_meansF$Year == 1984 |now_then_meansF$Year == 1985] <- "1980-1985_mean"

now_then_meansF$Year[now_then_meansF$Year == 2016 |now_then_meansF$Year == 2017 |now_then_meansF$Year == 2018 |now_then_meansF$Year == 2019 |now_then_meansF$Year == 2020 |now_then_meansF$Year == 2021] <- "2016-2021_mean"

###### C ##########
now_then_meansC <- now_then_meansV2[, c(1, 2, 4)]
now_then_meansC <- now_then_meansC %>%
  pivot_wider(names_from = "Region", values_from = "Climate")

now_then_meansC$Year[now_then_meansC$Year == 1980 |now_then_meansC$Year == 1981 |now_then_meansC$Year == 1982 |now_then_meansC$Year == 1983 |now_then_meansC$Year == 1984 |now_then_meansC$Year == 1985] <- "1980-1985_mean"

now_then_meansC$Year[now_then_meansC$Year == 2016 |now_then_meansC$Year == 2017 |now_then_meansC$Year == 2018 |now_then_meansC$Year == 2019 |now_then_meansC$Year == 2020 |now_then_meansC$Year == 2021] <- "2016-2021_mean"

######## B ############
now_then_meansB <- now_then_meansV2[, c(1, 2, 5)]
now_then_meansB <- now_then_meansB %>%
  pivot_wider(names_from = "Region", values_from = "Biodiversity")

now_then_meansB$Year[now_then_meansB$Year == 1980 |now_then_meansB$Year == 1981 |now_then_meansB$Year == 1982 |now_then_meansB$Year == 1983 |now_then_meansB$Year == 1984 |now_then_meansB$Year == 1985] <- "1980-1985_mean"

now_then_meansB$Year[now_then_meansB$Year == 2016 |now_then_meansB$Year == 2017 |now_then_meansB$Year == 2018 |now_then_meansB$Year == 2019 |now_then_meansB$Year == 2020 |now_then_meansB$Year == 2021] <- "2016-2021_mean"

######################################################################################

#### FCB means of countries #########
FCB_countries <- cbind(F_weighted_country, C_weighted_country[, c(3)])
FCB_countries <- cbind(FCB_countries, B_weighted_country[, c(3)])
names(FCB_countries)[3] <- "Food" #rename columns as F, C and B
names(FCB_countries)[4] <- "Climate"
names(FCB_countries)[5] <- "Biodiversity"

FCB_countries <- filter(FCB_countries, Year %in% c(1980, 1981, 1982, 1983, 1984, 1985, 2016, 2017, 2018, 2019, 2020, 2021))

######## F ############
country_now_then_meansF <- FCB_countries[, c(1:3)]
country_now_then_meansF <- country_now_then_meansF %>%
  pivot_wider(names_from = "Country", values_from = "Food")

country_now_then_meansF$Year[country_now_then_meansF$Year == 1980 |country_now_then_meansF$Year == 1981 |country_now_then_meansF$Year == 1982 |country_now_then_meansF$Year == 1983 |country_now_then_meansF$Year == 1984 |country_now_then_meansF$Year == 1985] <- "1980-1985_mean"

country_now_then_meansF$Year[country_now_then_meansF$Year == 2016 |country_now_then_meansF$Year == 2017 |country_now_then_meansF$Year == 2018 |country_now_then_meansF$Year == 2019 |country_now_then_meansF$Year == 2020 |country_now_then_meansF$Year == 2021] <- "2016-2021_mean"

######## C ############
country_now_then_meansC <- FCB_countries[, c(1, 2, 4)]
country_now_then_meansC <- country_now_then_meansC %>%
  pivot_wider(names_from = "Country", values_from = "Climate")

country_now_then_meansC$Year[country_now_then_meansC$Year == 1980 |country_now_then_meansC$Year == 1981 |country_now_then_meansC$Year == 1982 |country_now_then_meansC$Year == 1983 |country_now_then_meansC$Year == 1984 |country_now_then_meansC$Year == 1985] <- "1980-1985_mean"

country_now_then_meansC$Year[country_now_then_meansC$Year == 2016 |country_now_then_meansC$Year == 2017 |country_now_then_meansC$Year == 2018 |country_now_then_meansC$Year == 2019 |country_now_then_meansC$Year == 2020 |country_now_then_meansC$Year == 2021] <- "2016-2021_mean"

######## B ############
country_now_then_meansB <- FCB_countries[, c(1, 2, 5)]
country_now_then_meansB <- country_now_then_meansB %>%
  pivot_wider(names_from = "Country", values_from = "Biodiversity")

country_now_then_meansB$Year[country_now_then_meansB$Year == 1980 |country_now_then_meansB$Year == 1981 |country_now_then_meansB$Year == 1982 |country_now_then_meansB$Year == 1983 |country_now_then_meansB$Year == 1984 |country_now_then_meansB$Year == 1985] <- "1980-1985_mean"

country_now_then_meansB$Year[country_now_then_meansB$Year == 2016 |country_now_then_meansB$Year == 2017 |country_now_then_meansB$Year == 2018 |country_now_then_meansB$Year == 2019 |country_now_then_meansB$Year == 2020 |country_now_then_meansB$Year == 2021] <- "2016-2021_mean"
```

### STEP 6: Calculate change in 5-year means ####
```{r Rearrange data, mutate column with change in F, C, B: means, means_stats_F, mean_stats_country}

### CONTINENTS #################

#### F ####
means <- pivot_longer(now_then_meansF, cols = c(!Year),
                                names_to = "Region", values_to = "Mean") #make long

means<- means %>% 
  group_by(Region, Year) %>%
  summarise(mean = mean(Mean))

#subtract the current mean from the old mean to find the change
means_stats_F <- means %>%
  group_by(Region) %>%
  mutate(changeinF = mean[Year == "2016-2021_mean"] - mean[Year =="1980-1985_mean"])

# ggplot(means, aes(x = Year, y=mean, color = Region))+ geom_boxplot()+
#   scale_x_discrete(limits = c("1980-1985_mean", "2016-2021_mean"), labels = c("1980 - 1985", "2016 - 2021"))+
#   labs(y = "Mean F Index")+
#   scale_color_manual(values=cbPalette)

#### C ####
means <- pivot_longer(now_then_meansC, cols = c(!Year),
                                names_to = "Region", values_to = "Mean") #make long

means<- means %>% 
  group_by(Region, Year) %>%
  summarise(mean = mean(Mean))

#subtract the current mean from the old mean to find the change
means_stats_C <- means %>%
  group_by(Region) %>%
  mutate(changeinC = mean[Year == "2016-2021_mean"] - mean[Year =="1980-1985_mean"])

#### B ####
means <- pivot_longer(now_then_meansB, cols = c(!Year),
                                names_to = "Region", values_to = "Mean") #make long

means<- means %>% 
  group_by(Region, Year) %>%
  summarise(mean = mean(Mean))

#subtract the current mean from the old mean to find the change
means_stats_B <- means %>%
  group_by(Region) %>%
  mutate(changeinB = mean[Year == "2016-2021_mean"] - mean[Year =="1980-1985_mean"])


#Combine stats (change in FCB indices) into one table
mean_stats <- cbind(means_stats_F, means_stats_C[, c(4)])
mean_stats <- cbind(mean_stats, means_stats_B[, c(4)])

mean_stats <- mean_stats %>%
  subset(Year == "1980-1985_mean") #remove duplicates
mean_stats <- mean_stats[, -c(2, 3)]#remove extra columns

############ COUNTRIES #######
#### F ####
means <- pivot_longer(country_now_then_meansF, cols = c(!Year),
                                names_to = "Region", values_to = "Mean") #make long

means<- means %>% 
  group_by(Region, Year) %>%
  summarise(mean = mean(Mean))

#subtract the current mean from the old mean to find the change
means_stats_F <- means %>%
  group_by(Region) %>%
  mutate(changeinF = mean[Year == "2016-2021_mean"] - mean[Year =="1980-1985_mean"])

#### C ####
means <- pivot_longer(country_now_then_meansC, cols = c(!Year),
                                names_to = "Region", values_to = "Mean") #make long

means<- means %>% 
  group_by(Region, Year) %>%
  summarise(mean = mean(Mean))

#subtract the current mean from the old mean to find the change
means_stats_C <- means %>%
  group_by(Region) %>%
  mutate(changeinC = mean[Year == "2016-2021_mean"] - mean[Year =="1980-1985_mean"])

#### B ####
means <- pivot_longer(country_now_then_meansB, cols = c(!Year),
                                names_to = "Region", values_to = "Mean") #make long

means<- means %>% 
  group_by(Region, Year) %>%
  summarise(mean = mean(Mean))

#subtract the current mean from the old mean to find the change
means_stats_B <- means %>%
  group_by(Region) %>%
  mutate(changeinB = mean[Year == "2016-2021_mean"] - mean[Year =="1980-1985_mean"])

#Combine stats (change in FCB indices) into one graph
mean_stats_country <- cbind(means_stats_F, means_stats_C[, c(4)])
mean_stats_country <- cbind(mean_stats_country, means_stats_B[, c(4)])

mean_stats_country <- mean_stats_country %>%
  subset(Year == "1980-1985_mean") #remove duplicates
mean_stats_country <- mean_stats_country[, -c(2, 3)]#remove extra column
```

###STEP 7: Heatmap of regional scores ########
```{r}

setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Figures")

##Continents 2021 stat:
FCB_2021_cont <- rbind(now_then_meansF[6, ], now_then_meansC[6, ], now_then_meansB[6,])

names(FCB_2021_cont)[1] <- "2021 Index"
FCB_2021_cont[1, 1] <- "Food"
FCB_2021_cont[2, 1] <- "Climate"
FCB_2021_cont[3, 1] <- "Biodiversity"

#Countries 2021 stat:
FCB_2021_country <- rbind(country_now_then_meansF[6, ], country_now_then_meansC[6, ], country_now_then_meansB[6,])

names(FCB_2021_country)[1] <- "2021 Index"
FCB_2021_country[1, 1] <- "Food"
FCB_2021_country[2, 1] <- "Climate"
FCB_2021_country[3, 1] <- "Biodiversity"

#bind 
FCB_2021 <- cbind(FCB_2021_cont, FCB_2021_country[, 3:5])

#make long
FCB_2021 <- pivot_longer(FCB_2021, cols = c(!"2021 Index"),
                                names_to = "Region", values_to = "2021 Score")
FCB_2021 <- pivot_wider(FCB_2021, names_from = "2021 Index", values_from = "2021 Score")

#export
# write.csv(FCB_2021, file = "Regions_FCB2021.csv")

###################################

#Continents 2016-2021 mean: 
F_2016_2021_cont <- F_weighted_continent %>% group_by(Region) %>%
  filter(Year %in% c( 2016, 2017, 2018, 2019, 2020, 2021)) %>%
  summarise(F_2016_2021_cont = mean(Score))

C_2016_2021_cont <- C_weighted_continent %>% group_by(Region) %>%
  filter(Year %in% c( 2016, 2017, 2018, 2019, 2020, 2021)) %>%
  summarise(C_2016_2021_cont = mean(Score))

B_2016_2021_cont <- B_weighted_continent %>% group_by(Region) %>%
  filter(Year %in% c( 2016, 2017, 2018, 2019, 2020, 2021)) %>%
  summarise(B_2016_2021_cont = mean(Score))

FCB_2016_2021_cont <- merge(F_2016_2021_cont, C_2016_2021_cont, by = "Region")
FCB_2016_2021_cont <- merge(FCB_2016_2021_cont, B_2016_2021_cont, by = "Region")
names(FCB_2016_2021_cont)[2] <- "Food"
names(FCB_2016_2021_cont)[3] <- "Climate"
names(FCB_2016_2021_cont)[4] <- "Biodiversity"

#Country 2016-2021 mean: 
F_2016_2021_country <- F_weighted_country %>% group_by(Country) %>%
  filter(Year %in% c( 2016, 2017, 2018, 2019, 2020, 2021)) %>%
  summarise(F_2016_2021_country = mean(Score))

C_2016_2021_country <- C_weighted_country %>% group_by(Country) %>%
  filter(Year %in% c( 2016, 2017, 2018, 2019, 2020, 2021)) %>%
  summarise(C_2016_2021_country = mean(Score))

B_2016_2021_country <- B_weighted_country %>% group_by(Country) %>%
  filter(Year %in% c( 2016, 2017, 2018, 2019, 2020, 2021)) %>%
  summarise(B_2016_2021_country = mean(Score))

FCB_2016_2021_country <- merge(F_2016_2021_country, C_2016_2021_country, by = "Country")
FCB_2016_2021_country <- merge(FCB_2016_2021_country, B_2016_2021_country, by = "Country")
names(FCB_2016_2021_country)[1] <- "Region"
names(FCB_2016_2021_country)[2] <- "Food"
names(FCB_2016_2021_country)[3] <- "Climate"
names(FCB_2016_2021_country)[4] <- "Biodiversity"


#bind
FCB_2016_2021 <- rbind(FCB_2016_2021_cont[c(1:3, 5:6), ], FCB_2016_2021_country)

#export
# write.csv(FCB_2016_2021, file = "Regions_FCB2016_2021.csv")


#Make a heatmap
heatmapdata <- read.csv("Regions_FCB2016_2021.csv")

heatmapdata <- heatmapdata[, c(2:5)]

heatmapdata$Food <- as.numeric(heatmapdata$Food)
heatmapdata$Climate <- as.numeric(heatmapdata$Climate)
heatmapdata$Biodiversity <- as.numeric(heatmapdata$Biodiversity)

heatmapdata <- pivot_longer(heatmapdata, cols = -c("Region"), 
                                names_to = "index", values_to = "score")

###
plot <- ggplot(heatmapdata, aes(y= index, x = Region, fill= score)) + 
  geom_raster()+
  scale_fill_viridis_c(option = "viridis")+
  labs(fill = "5-Year Mean")+
  theme(axis.title.x = element_blank(),
        axis.title.y=element_blank(),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 10, angle = 15))+
  scale_y_discrete(limits = c('Biodiversity', 'Climate', 'Food'), labels = c("B", "C", "F"))+
   scale_x_discrete(limits = c('Chile', 'China','Nigeria','Norway', 'Africa', 'Americas', 'Asia','Europe','Oceania'), labels = c('Chile', 'China','Nigeria','Norway', 'Africa', 'Americas', 'Asia','Europe','Oceania'))


#save
# ggsave("FCB_2016_2021_heatmap.png", plot, height = 4, width = 8, dpi = 600)
```

###STEP 8: SUMMARY STATS #############
```{r}

setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Figures")
library(gridExtra)

test <- rbind(mean_stats[c(1:3, 5:6), ], mean_stats_country)

#FOOD ###################################################################################
Foodsumstats <- group_by(FCB_countries, Country) %>% #for all data (to see interannual variability)
  summarise(
    count = n(),
    mean = mean(Food, na.rm = TRUE),
    IQR = IQR(Food, na.rm = TRUE),
    SD = sd(Food, na.rm = TRUE))

Foodsumstats2 <- group_by(now_then_meansV2, Region) %>% #for all data (to see interannual variability)
  summarise(
    count = n(),
    mean = mean(Food, na.rm = TRUE),
    IQR = IQR(Food, na.rm = TRUE),
    SD = sd(Food, na.rm = TRUE))

  names(Foodsumstats)[1] <- "Region"

  Foodsumstats <- rbind(Foodsumstats2[c(1:3, 5:6), ], Foodsumstats) #bind

  Foodsumstats <- cbind(Foodsumstats[, c(1, 3:5)], test[, 2])
  
png("Foodsummarystats.png", height = 200*nrow(Foodsumstats), width = 600*ncol(Foodsumstats), res = 600)
grid.table(Foodsumstats)
dev.off()

###CLIMATE CHANGE ###############################################################

Climsumstats <- group_by(FCB_countries, Country) %>% #for all data (to see interannual variability)
  summarise(
    count = n(),
    mean = mean(Climate, na.rm = TRUE),
    IQR = IQR(Climate, na.rm = TRUE),
    SD = sd(Climate, na.rm = TRUE))

Climsumstats2 <- group_by(now_then_meansV2, Region) %>% #for all data (to see interannual variability)
  summarise(
    count = n(),
    mean = mean(Climate, na.rm = TRUE),
    IQR = IQR(Climate, na.rm = TRUE),
    SD = sd(Climate, na.rm = TRUE))

  names(Climsumstats)[1] <- "Region"

  Climsumstats <- rbind(Climsumstats2[c(1:3, 5:6), ], Climsumstats) #bind

  Climsumstats <- cbind(Climsumstats[, c(1, 3:5)], test[, 3])
    
png("Climsumstats.png", height = 200*nrow(Climsumstats), width = 600*ncol(Climsumstats), res = 600)
grid.table(Climsumstats)
dev.off()

########BIODIVERSITY ########################################################################
Biodivsumstats <- group_by(FCB_countries, Country) %>% #for all data (to see interannual variability)
  summarise(
    count = n(),
    mean = mean(Biodiversity, na.rm = TRUE),
    IQR = IQR(Biodiversity, na.rm = TRUE),
    SD = sd(Biodiversity, na.rm = TRUE))

Biodivsumstats2 <- group_by(now_then_meansV2, Region) %>% #for all data (to see interannual variability)
  summarise(
    count = n(),
    mean = mean(Biodiversity, na.rm = TRUE),
    IQR = IQR(Biodiversity, na.rm = TRUE),
    SD = sd(Biodiversity, na.rm = TRUE))

  names(Biodivsumstats)[1] <- "Region"

  Biodivsumstats <- rbind(Biodivsumstats2[c(1:3, 5:6), ], Biodivsumstats) #bind

  Biodivsumstats <- cbind(Biodivsumstats[, c(1, 3:5)], test[, 4])
    
png("Biodivsumstats.png", height = 200*nrow(Biodivsumstats), width = 600*ncol(Biodivsumstats), res = 600)
grid.table(Biodivsumstats)
dev.off()
```

###STEP 9: Visualize difference in 5-year means with boxplots ################
```{r}

setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Figures")

SD_F_facet <- cbind(now_then_meansF, country_now_then_meansF[, c(3:5)]) %>%
  pivot_longer(cols = c(!Year), names_to = "Region", values_to = "Mean") %>%
    ggplot(aes(x= Year, y=Mean, color = Year))+
    geom_boxplot()+
    facet_wrap(~Region)+ #scales = "free"
    theme(axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank())+
    labs(y = "Food Security Index", color = "")+
    scale_color_manual(values = c("#E69F00", "#56B4E9"), labels = c('1980 - 1985', '2016 - 2021'))
  
ggsave("SD_F_facet.png", SD_F_facet, height = 4, width = 8, dpi = 600)

SD_C_facet <- cbind(now_then_meansC, country_now_then_meansC[, c(3:5)]) %>%
  pivot_longer(cols = c(!Year), names_to = "Region", values_to = "Mean") %>%
    ggplot(aes(x= Year, y=Mean, color = Year))+
    geom_boxplot()+
    facet_wrap(~Region)+
    theme(axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank())+
    labs(y = "Climate Change Index", color = "")+
    scale_color_manual(values = c("#E69F00", "#56B4E9"), labels = c('1980 - 1985', '2016 - 2021'))

ggsave("SD_C_facet.png", SD_C_facet, height = 4, width = 8, dpi = 600)

SD_B_facet <- cbind(now_then_meansB, country_now_then_meansB[, c(3:5)]) %>%
  pivot_longer(cols = c(!Year), names_to = "Region", values_to = "Mean") %>%
    ggplot(aes(x= Year, y=Mean, color = Year))+
    geom_boxplot()+
    facet_wrap(~Region)+
    theme(axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank())+
    labs(y = "Biodiversity Index", color = "")+
    scale_color_manual(values = c("#E69F00", "#56B4E9"), labels = c('1980 - 1985', '2016 - 2021'))

ggsave("SD_B_facet.png", SD_B_facet, height = 4, width = 8, dpi = 600)
```

#### STEP 10: Calculate theoretical max! ##############
```{r}
speciesdata <- read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/FCB_scores_2024.csv") #FCB scores for each species
speciesdata <- speciesdata[, c(2, 6:8)] #remove extra columns
speciesdata <- speciesdata[order(speciesdata$species), ] #put in alphabetical order
speciesdata <- speciesdata[!speciesdata$species == "Eucheuma cottonii",] #remove E. Cottonii, as it is a duplicate

production_data <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prod_by_cat910.csv") #annual production by species
production_data <- production_data[, c(2:4)] #remove unnecessary columns
names(production_data)[1] <- "species"

#bind with species scores?
allprod <- merge(production_data, speciesdata, by = "species")

#Remove China from datasets so we can calculate Asian total production excluding China:
all_production_withoutCH <-  subset(allprod, allprod$Country.Name.En != "China")

#Just China
just_china_prod <- subset(allprod, allprod$Country.Name.En == "China")

#All country case studies: China, Norway, Chile and Nigeria

country_case_study <- allprod[allprod$Country.Name.En %in% c("Norway", "Chile", "Nigeria"), ]

# "/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prod_totals_countries.csv"
```

```{r Find top-scoring species for each region}

#sort by region and then rank by highest-scoring species: Major continents not including China
highestF <- all_production_withoutCH %>%
  group_by(Continent.Name.En) %>%
  filter(f == max(f))

highestF <- highestF[c(1, 4, 5, 6, 10), c(1, 2, 4)] 
# highestF <- pivot_wider(highestF, names_from = "Continent.Name.En", values_from = "f")

highestC <- all_production_withoutCH %>%
  group_by(Continent.Name.En) %>%
  filter(c == max(c))

highestC <- highestC[c(1, 35, 37, 41, 42), c(1, 2, 5)]

highestB <- all_production_withoutCH %>%
  group_by(Continent.Name.En) %>%
  filter(b == max(b))

highestB <- highestB[c(1, 2, 7, 8, 23), c(1, 2, 6)]

#bind: just the highest scores
highestFCB <- merge(highestF[, c(2, 3)], highestC[, c(2,3)], by = "Continent.Name.En")
highestFCB <- merge(highestFCB, highestB[, c(2,3)], by = "Continent.Name.En")
names(highestFCB)[1] <- "Region"
names(highestFCB)[2] <- "Food"
names(highestFCB)[3] <- "Climate"
names(highestFCB)[4] <- "Biodiversity"

#bind: just the species names
highestFCBsp <- merge(highestF[, c(1, 2)], highestC[, c(1, 2)], by = "Continent.Name.En")
highestFCBsp <- merge(highestFCBsp, highestB[, c(1, 2)], by = "Continent.Name.En")
names(highestFCBsp)[1] <- "Region"
names(highestFCBsp)[2] <- "Food"
names(highestFCBsp)[3] <- "Climate"
names(highestFCBsp)[4] <- "Biodiversity"

####separate for China ################################################################################
CH_F <- just_china_prod %>%
  filter(f == max(f))

CH_C <- just_china_prod %>%
  filter(c == max(c))

CH_B <- just_china_prod %>%
  filter(b == max(b))

#bind: just the highest scores
CH_highestFCB <- cbind(CH_F[, c(3, 4)], CH_C[, 5])
CH_highestFCB <- cbind(CH_highestFCB, CH_B[, 6])
names(CH_highestFCB)[1] <- "Region"
names(CH_highestFCB)[2] <- "Food"
names(CH_highestFCB)[3] <- "Climate"
names(CH_highestFCB)[4] <- "Biodiversity"

#bind: just the species names
CH_highestFCB_sp <- merge(CH_F[, c(1, 3)], CH_C[, c(1, 3)], by = "Country.Name.En")
CH_highestFCB_sp <- merge(CH_highestFCB_sp, CH_B[, c(1, 3)], by = "Country.Name.En")
names(CH_highestFCB_sp)[1] <- "Region"
names(CH_highestFCB_sp)[2] <- "Food"
names(CH_highestFCB_sp)[3] <- "Climate"
names(CH_highestFCB_sp)[4] <- "Biodiversity"

####separate for Norway, Nigeria, Chile  ############################################################
highestF2 <- country_case_study %>%
  group_by(Country.Name.En) %>%
  filter(f == max(f))

highestF2 <- highestF2[, c(1, 3, 4)]

highestC2 <- country_case_study %>%
  group_by(Country.Name.En) %>%
  filter(c == max(c))

highestC2 <- highestC2[c(1:3), c(1, 3, 5)] 

highestB2 <- country_case_study %>%
  group_by(Country.Name.En) %>%
  filter(b == max(b))

highestB2 <- highestB2[c(1:3), c(1, 3, 6)] 

#bind: just the highest scores
highestFCB2 <- merge(highestF2[, c(2, 3)], highestC2[, c(2, 3)], by = "Country.Name.En")
highestFCB2 <- merge(highestFCB2, highestB2[, c(2, 3)], by = "Country.Name.En")
names(highestFCB2)[1] <- "Region"
names(highestFCB2)[2] <- "Food"
names(highestFCB2)[3] <- "Climate"
names(highestFCB2)[4] <- "Biodiversity"

#bind: just the species names
highestFCBsp2 <- merge(highestF2[, c(1, 2)], highestC2[, c(1, 2)], by = "Country.Name.En")
highestFCBsp2 <- merge(highestFCBsp2, highestB2[, c(1, 2)], by = "Country.Name.En")
names(highestFCBsp2)[1] <- "Region"
names(highestFCBsp2)[2] <- "Food"
names(highestFCBsp2)[3] <- "Climate"
names(highestFCBsp2)[4] <- "Biodiversity"

############### Merge data frames

#Theoretical max FCB scores
theo_max <- rbind(highestFCB, CH_highestFCB, highestFCB2)
theo_max <- theo_max[order(theo_max$Region),]

png("theo_max.png", height = 200*nrow(theo_max), width = 600*ncol(theo_max), res = 600)
grid.table(theo_max)
dev.off()

#Species
theo_max_sp <- rbind(highestFCBsp, CH_highestFCB_sp, highestFCBsp2)
theo_max_sp <- theo_max_sp[order(theo_max_sp$Region),]

png("theo_max_sp.png", height = 200*nrow(theo_max_sp), width = 1100*ncol(theo_max_sp), res = 600)
grid.table(theo_max_sp)
dev.off()

```

```{r Calculate % change}
setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Figures")

names(mean_stats_country)[1] <- "Region"

test <- rbind(mean_stats[c(1:3, 5:6),], mean_stats_country)

test <- test[order(test$Region),]

Perc_max <- mapply(`/`,test[, c(2:4)],theo_max[, c(2:4)])
Perc_max <- Perc_max*100
names(Perc_max)[2] <- "% of F Max"
names(Perc_max)[3] <- "% of C Max"
names(Perc_max)[4] <- "% of B Max"

Perc_max <-cbind(test[, c(1)], Perc_max)

write.csv(Perc_max, file = "FCB_change_perc.csv")
```

```{r Plot as heatmap}

Perc_max$changeinF <- as.numeric(Perc_max$changeinF)
Perc_max$changeinC <- as.numeric(Perc_max$changeinC)
Perc_max$changeinB <- as.numeric(Perc_max$changeinB)

heatmapdata <- Perc_max
heatmapdata <- pivot_longer(heatmapdata, cols = -c("Region"), 
                                names_to = "index", values_to = "perc_change")

###
plot <- ggplot(heatmapdata, aes(y= index, x = Region, fill= perc_change)) + 
  geom_raster()+
  scale_fill_viridis_c(option = "viridis")+ #viridis is colorblind friendly
  labs(fill = "perc_change")+
  theme(axis.title.x = element_blank(),
        axis.title.y=element_blank(),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 10, angle = 15))+
  scale_y_discrete(limits = c('changeinB', 'changeinC', 'changeinF'), labels = c("B", "C", "F"))+
   scale_x_discrete(limits = c('Chile', 'China','Nigeria','Norway', 'Africa', 'Americas', 'Asia','Europe','Oceania'), labels = c('Chile', 'China','Nigeria','Norway', 'Africa', 'Americas', 'Asia','Europe','Oceania'))


ggsave("mean_percchange_heatmap.png", plot, height = 4, width = 8, dpi = 600)
```
