---
title: "temporal_clean"
author: "Aleah Wong"
date: '2024-08-29'
output: html_document
---

```{r}
rm(list = ls())
.rs.restartR()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
```

############################################################
Steps 1-4 are stored in the dataprep file
1. Calculate annual production totals for each species (data saved/provided)
2. Clean species names and prepare dataset (data saved/provided)
3. Graphing production (don't need to run steps 1-2)
4. Calculating production by taxa--use production values to weight FCB scores by taxonomic category
-----------------------------------------------
5. Weight F scores by production
6. Weight C scores by production
7. Weight B scores by production
8. Graphing weighted FCB scores together 
9. FCB Correlations over time
10. Stats tests?
############################################################

###STEP 5: Weight food security scores by production ###
```{r}
#checking the difference between these datasets 

# prodbytaxa_alpha <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prodbytaxa_alphab.csv") 
# prodbytaxa_clean <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prodbytaxa_clean.csv") #has the category column
```

```{r Loading in data}

speciesdata <- read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/FCB_scores_2024.csv") #FCB scores for each species
speciesdata <- speciesdata[, c(2:12)] #remove extra first column
speciesdata <- speciesdata[order(speciesdata$species), ] #put in alphabetical order
speciesdata <- speciesdata[!speciesdata$species == "Eucheuma cottonii",] #remove E. Cottonii, as it is a duplicate

weighting_totals <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prodbytaxa_forweighting.csv") #annual production by taxonomic group
weighting_totals <- weighting_totals[, c(2:74)] #remove extra first column

production_data <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prodbytaxa_clean.csv") #annual production by species
production_data <- production_data[, c(2:75)] #remove extra first column

#load in this dataset to use as an empty frame, just species and year columns
frame <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/prodbytaxa_alphab.csv")
frame <- frame[, c(2:74)] #remove extra column
```

```{r FOOD SECURITY WEIGHTED BY ALL PRODUCTION}
#calculate index scores for each species weighted by production of ALL 55 species

new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

#loop through each species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata)){ #for each species
  print(speciesdata$species[i])
  f_score <- speciesdata$f[i] #get the score, CHANGE FOR B AND C
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[1, b-1]) #weight by production of ALL species
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[57, i] <- a
}

new_frame[57, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[57, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#SAVE this data to graph FCB together in step 8
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Fweightedby_all.csv")
# write.csv(weighted_indexes,file=NAME)
```

```{r FOOD SECURITY WEIGHTED BY FINFISH}
new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

#loop through each FINFISH species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine") #filter out just the finfish

production_data2 <- filter(production_data, category =="finfish_inland"| category =="finfish_marine")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the score #CHANGE THIS FOR B AND C
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data2[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[2, b-1]) #weight by production of finfish
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2] #include species
  } 
  
}

#calculate the annual mean each year
for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[57, i] <- a
}

new_frame[57, 1] <- "Weighted Index" #name the newest row
weighted_indexes1 <- new_frame[57, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes1 <- pivot_longer(weighted_indexes1, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes1 <- weighted_indexes1[order(weighted_indexes1$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes1)) {
  print(weighted_indexes1$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes1$Year[i] <- year_labs[i]
}

weighted_indexes1$Year <- as.numeric(weighted_indexes1$Year) #make years numeric
```

```{r FOOD SECURITY WEIGHTED BY ALGAE}
new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

#loop through each ALGAE species and multiply f score by (production of that species/production of all species)
#production_data (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "algae") #filter out just the algae

production_data2 <- filter(production_data, category =="algae")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$f[i] #get the score #CHANGE THIS FOR B AND C

  for (b in 3:74){ #for each year
  index <- f_score*production_data2[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[4, b-1]) #weight by production of algae, row 4

  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  }

}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)

    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes2 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes2 <- pivot_longer(weighted_indexes2, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes2 <- weighted_indexes2[order(weighted_indexes2$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes2)) {
  print(weighted_indexes2$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes2$Year[i] <- year_labs[i]
}

weighted_indexes2$Year <- as.numeric(weighted_indexes2$Year) #make years numeric
```

```{r FOOD SECURITY WEIGHTED BY CRUSTACEANS and MOLLUSCS}
new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "crustacean")

production_data2 <- filter(production_data, category == "crustacean")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score

  for (b in 3:74){ #for each year
  index <- f_score*production_data2[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[5, b-1]) #weight by production of crustaceans, row 5

  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  }

}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)

    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes3 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes3 <- pivot_longer(weighted_indexes3, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes3 <- weighted_indexes3[order(weighted_indexes3$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes3)) {
  print(weighted_indexes3$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes3$Year[i] <- year_labs[i]
}

weighted_indexes3$Year <- as.numeric(weighted_indexes3$Year) #make years numeric
----------------------------------------------------------------------------------------------------
#MOLLUSCS
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "mollusc")

production_data2 <- filter(production_data, category == "mollusc")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data2[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[6, b-1]) #weight by production of molluscs, row 6
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes4 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes4 <- pivot_longer(weighted_indexes4, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes4 <- weighted_indexes4[order(weighted_indexes4$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes4)) {
  print(weighted_indexes4$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes4$Year[i] <- year_labs[i]
}

weighted_indexes4$Year <- as.numeric(weighted_indexes4$Year) #make years numeric
```

```{r Graphing food security index over time with all taxa weightings}

setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Figures")

#weighted_indexes = weighted by all
#need to add identifier column
weighted_indexes$weighted_by <- "All"

#weighted_indexes1 = weighted by finfish
weighted_indexes1$weighted_by <- "Finfish"

#weighted_indexes2 = weighted by algae
weighted_indexes2$weighted_by <- "Algae"

#weighted_indexes3 = weighted by crustaceans
weighted_indexes3$weighted_by <- "Crustaceans"

#weighted_indexes4 = weighted by molluscs
weighted_indexes4$weighted_by <- "Molluscs"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexes1)
merge <- rbind(merge, weighted_indexes2)
merge <- rbind(merge, weighted_indexes3)
merge <- rbind(merge, weighted_indexes4)

#save this data
NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/weighted_Fscores.csv")
write.csv(merge,file=NAME)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = weighted_by)) +geom_line(size =1)+
  labs(y = "Mean Food Security Index", title = "Food Security Potential of Aquaculture")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))+
    ylim(30,65)

#SAVE
# ggsave("F_weighted_combined830.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

```{r}
#magnitude of F, C and B indices by taxonomic group in three time periods
# merge %>%
#   group_by(weighted_by) %>%
#   summarise(min = min(Score), max = max(Score))
```

```{r Graphing means}

setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Figures")

#modify dataset: create three time periods --> change this for C and B too
scores_3_periods <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/weighted_Fscores.csv")
scores_3_periods <- scores_3_periods[, c(2:4)] #remove extra column

#modify dataset: create three time periods
# scores_3_periods <- merge

scores_3_periods$Year[scores_3_periods$Year >= 1950 & scores_3_periods$Year <= 1980] <- "1950-1980"
scores_3_periods$Year[scores_3_periods$Year >= 1981 & scores_3_periods$Year <= 2000] <- "1981-2000"
scores_3_periods$Year[scores_3_periods$Year >= 2001 & scores_3_periods$Year <= 2021] <- "2001-2021"

plot <- ggplot(scores_3_periods, aes(x = Year, y=Score, color = weighted_by))+ geom_boxplot()+
  stat_summary(fun.y=mean,geom='point', position = position_dodge(width = .75))+
  stat_summary(fun=mean, geom="line", aes(group=weighted_by), position = position_dodge(width = .75))+
  scale_x_discrete(limits = c("1950-1980", "1981-2000", "2001-2021"))+
  scale_y_continuous(limits = c(30, 67))+
  labs(y = "Mean Food Security Index")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))


# ggsave("mean_F_boxplots830.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

###STEP 6: Weight climate scores by production ###
```{r CLIMATE CHANGE WEIGHTED BY ALL PRODUCTION}
#calculate index scores for each species weighted by production of ALL 54 species

new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

#loop through each species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata)){ #for each species
  print(speciesdata$species[i])
  f_score <- speciesdata$c[i] #get the score
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[1, b-1]) #weight by production of ALL species
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#save data to graph weighted FCB together in step 8
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Cweightedby_all.csv")
# write.csv(weighted_indexes,file=NAME)
--------------------------------------------------------
  
#CLIMATE CHANGE WEIGHTED BY FINFISH
new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

#loop through each FINFISH species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine") #filter out just the finfish

production_data2 <- filter(production_data, category =="finfish_inland"| category =="finfish_marine")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$c[i] #get the score
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data2[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[2, b-1]) #weight by production of finfish
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year
for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes1 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes1 <- pivot_longer(weighted_indexes1, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes1 <- weighted_indexes1[order(weighted_indexes1$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes1)) {
  print(weighted_indexes1$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes1$Year[i] <- year_labs[i]
}

weighted_indexes1$Year <- as.numeric(weighted_indexes1$Year) #make years numeric
-----------------------------------------------------------------------------------------
  
#CLIMATE CHANGE WEIGHTED BY ALGAE
new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

#loop through each ALGAE species and multiply f score by (production of that species/production of all species)
#production_data (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "algae") #filter out just the algae

production_data2 <- filter(production_data, category =="algae")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$c[i] #get the score
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data2[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[4, b-1]) #weight by production of algae, row 4
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes2 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes2 <- pivot_longer(weighted_indexes2, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes2 <- weighted_indexes2[order(weighted_indexes2$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes2)) {
  print(weighted_indexes2$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes2$Year[i] <- year_labs[i]
}

weighted_indexes2$Year <- as.numeric(weighted_indexes2$Year) #make years numeric
----------------------------------------------------------------------------------------------------
  
#CLIMATE CHANGE WEIGHTED BY CRUSTACEANS
new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "crustacean")

production_data2 <- filter(production_data, category == "crustacean")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$c[i] #get the c score
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data2[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[5, b-1]) #weight by production of crustaceans, row 5
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes3 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes3 <- pivot_longer(weighted_indexes3, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes3 <- weighted_indexes3[order(weighted_indexes3$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes3)) {
  print(weighted_indexes3$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes3$Year[i] <- year_labs[i]
}

weighted_indexes3$Year <- as.numeric(weighted_indexes3$Year) #make years numeric
----------------------------------------------------------------------------------------
#MOLLUSCS
new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "mollusc")

production_data2 <- filter(production_data, category == "mollusc")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$c[i] #get the c score
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data2[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[6, b-1]) #weight by production of molluscs, row 6
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes4 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes4 <- pivot_longer(weighted_indexes4, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes4 <- weighted_indexes4[order(weighted_indexes4$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes4)) {
  print(weighted_indexes4$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes4$Year[i] <- year_labs[i]
}

weighted_indexes4$Year <- as.numeric(weighted_indexes4$Year) #make years numeric
------------------------------------------------------------------------------------------------

#GRAPHING
#need to add identifier column
weighted_indexes$weighted_by <- "All"

#weighted_indexes1 = weighted by finfish
weighted_indexes1$weighted_by <- "Finfish"

#weighted_indexes2 = weighted by algae
weighted_indexes2$weighted_by <- "Algae"

#weighted_indexes3 = weighted by crustaceans
weighted_indexes3$weighted_by <- "Crustaceans"

#weighted_indexes4 = weighted by molluscs
weighted_indexes4$weighted_by <- "Molluscs"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexes1)
merge <- rbind(merge, weighted_indexes2)
merge <- rbind(merge, weighted_indexes3)
merge <- rbind(merge, weighted_indexes4)

#save this data
NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/weighted_Cscores.csv")
write.csv(merge,file=NAME)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = weighted_by)) +geom_line(size =1)+
  labs(y = "Mean Climate Change Index", title = "Climate Change Potential of Aquaculture")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))+
  ylim(30,65)

#SAVE
setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Figures")
# ggsave("C_weighted_combined.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set

```

```{r Graphing means in three time periods}

#modify dataset: create three time periods
scores_3_periods <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/weighted_Cscores.csv")
scores_3_periods <- scores_3_periods[, c(2:4)] #remove extra column

scores_3_periods$Year[scores_3_periods$Year >= 1950 & scores_3_periods$Year <= 1980] <- "1950-1980"
scores_3_periods$Year[scores_3_periods$Year >= 1981 & scores_3_periods$Year <= 2000] <- "1981-2000"
scores_3_periods$Year[scores_3_periods$Year >= 2001 & scores_3_periods$Year <= 2021] <- "2001-2021"

plot <- ggplot(scores_3_periods, aes(x = Year, y=Score, color = weighted_by))+ geom_boxplot()+
  stat_summary(fun.y=mean,geom='point', position = position_dodge(width = .75))+
  stat_summary(fun=mean, geom="line", aes(group=weighted_by), position = position_dodge(width = .75))+
  scale_x_discrete(limits = c("1950-1980", "1981-2000", "2001-2021"))+
  scale_y_continuous(limits = c(30, 67))+
  labs(y = "Mean Climate Index")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))

setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Figures")
# ggsave("mean_C_boxplots830.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

####STEP 7: Weight biodiversity  scores by production ################
```{r BIODIV WEIGHTED BY ALL TAXA}
#calculate index scores for each species weighted by production of ALL 54 species

new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

#loop through each species and multiply b score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata)){ #for each species
  print(speciesdata$species[i])
  f_score <- speciesdata$b[i] #get the score
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[1, b-1]) #weight by production of ALL species
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#Save data to graph weighted FCB together in Step 8
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Bweightedby_all.csv")
# write.csv(weighted_indexes,file=NAME)
```

```{r WEIGHTED BY FINFISH}
new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

#loop through each FINFISH species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine") #filter out just the finfish

production_data2 <- filter(production_data, category =="finfish_inland"| category =="finfish_marine")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$b[i] #get the score
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data2[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[2, b-1]) #weight by production of finfish
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year
for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes1 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes1 <- pivot_longer(weighted_indexes1, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes1 <- weighted_indexes1[order(weighted_indexes1$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes1)) {
  print(weighted_indexes1$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes1$Year[i] <- year_labs[i]
}

weighted_indexes1$Year <- as.numeric(weighted_indexes1$Year) #make years numeric
```

```{r WEIGHTED BY ALGAE}

new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

#loop through each ALGAE species and multiply f score by (production of that species/production of all species)
#production_data (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "algae") #filter out just the algae

production_data2 <- filter(production_data, category =="algae")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$b[i] #get the score
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data2[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[4, b-1]) #weight by production of algae, row 4
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes2 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes2 <- pivot_longer(weighted_indexes2, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes2 <- weighted_indexes2[order(weighted_indexes2$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes2)) {
  print(weighted_indexes2$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes2$Year[i] <- year_labs[i]
}

weighted_indexes2$Year <- as.numeric(weighted_indexes2$Year) #make years numeric
```

```{r WEIGHTED BY CRUSTACEANS}

new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "crustacean")

production_data2 <- filter(production_data, category == "crustacean")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$b[i] #get the b score
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data2[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[5, b-1]) #weight by production of crustaceans, row 5
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes3 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes3 <- pivot_longer(weighted_indexes3, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes3 <- weighted_indexes3[order(weighted_indexes3$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes3)) {
  print(weighted_indexes3$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes3$Year[i] <- year_labs[i]
}

weighted_indexes3$Year <- as.numeric(weighted_indexes3$Year) #make years numeric
```

```{r WEIGHTED BY MOLLUSCS}
new_frame <- frame #create new frame
new_frame <- frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "mollusc")

production_data2 <- filter(production_data, category == "mollusc")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$b[i] #get the b score
  
  for (b in 3:74){ #for each year
  index <- f_score*production_data2[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[6, b-1]) #weight by production of molluscs, row 6
  
  new_frame[i, b-1] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes4 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes4 <- pivot_longer(weighted_indexes4, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes4 <- weighted_indexes4[order(weighted_indexes4$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes4)) {
  print(weighted_indexes4$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes4$Year[i] <- year_labs[i]
}

weighted_indexes4$Year <- as.numeric(weighted_indexes4$Year) #make years numeric
```

```{r GRAPHING B}

#GRAPHING
#need to add identifier column
weighted_indexes$weighted_by <- "All"

#weighted_indexes1 = weighted by finfish
weighted_indexes1$weighted_by <- "Finfish"

#weighted_indexes2 = weighted by algae
weighted_indexes2$weighted_by <- "Algae"

#weighted_indexes3 = weighted by crustaceans
weighted_indexes3$weighted_by <- "Crustaceans"

#weighted_indexes4 = weighted by molluscs
weighted_indexes4$weighted_by <- "Molluscs"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexes1)
merge <- rbind(merge, weighted_indexes2)
merge <- rbind(merge, weighted_indexes3)
merge <- rbind(merge, weighted_indexes4)

#save this data
NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/weighted_Bscores.csv")
write.csv(merge,file=NAME)

#graph
ggplot(merge, aes(x=Year, y=Score, color = weighted_by)) +geom_line(size =1)+
  labs(y = "Mean Biodiversity Index", title = "Biodiversity Potential of Aquaculture")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))

#range 20-70?
#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = weighted_by)) +geom_line(size =1)+
  labs(y = "Mean Biodiversity Index", title = "Biodiversity Potential of Aquaculture")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))+
  ylim(30,70)

#SAVE
# ggsave("B_weighted_combined830.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set

```

```{r Graphing means in three time periods}
#magnitude of F, C and B indices by taxonomic group in three time periods

#modify dataset: create three time periods
# scores_3_periods <- merge
scores_3_periods <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/weighted_Bscores.csv")
scores_3_periods <- scores_3_periods[, c(2:4)] #remove extra column

scores_3_periods$Year[scores_3_periods$Year >= 1950 & scores_3_periods$Year <= 1980] <- "1950-1980"
scores_3_periods$Year[scores_3_periods$Year >= 1981 & scores_3_periods$Year <= 2000] <- "1981-2000"
scores_3_periods$Year[scores_3_periods$Year >= 2001 & scores_3_periods$Year <= 2021] <- "2001-2021"

plot <- ggplot(scores_3_periods, aes(x = Year, y=Score, color = weighted_by))+ geom_boxplot()+
  stat_summary(fun.y=mean,geom='point', position = position_dodge(width = .75))+
  stat_summary(fun=mean, geom="line", aes(group=weighted_by), position = position_dodge(width = .75))+
  scale_x_discrete(limits = c("1950-1980", "1981-2000", "2001-2021"))+
  scale_y_continuous(limits = c(30, 67))+
  labs(y = "Mean Biodiversity Index")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))

# setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Figures")
# ggsave("mean_B_boxplots830.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

#STEP 8:Combining weighted F, C, B into one graph ########
```{r COMBINED FCB GRAPH WEIGHTED BY PRODUCTION OF ALL SPECIES}

#NEW TEST 8.30
#load in the data saved in steps 5-7
F_weighted_all <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Fweightedby_all.csv")
F_weighted_all <- F_weighted_all[, c(2:3)]
C_weighted_all <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Cweightedby_all.csv")
C_weighted_all <- C_weighted_all[, c(2:3)]
B_weighted_all <- read.csv("/Users/aleahw/Documents/Master's!!/Chapter_2/Data/Bweightedby_all.csv")
B_weighted_all <- B_weighted_all[, c(2:3)]

#Add FCB labels to each dataset before merging
F_weighted_all$Index <- "Food Security"
#weighted_indexesc = C weighted by all 
C_weighted_all$Index <- "Climate Change"
#weighted_indexesb = B weighted by all 
B_weighted_all$Index <- "Biodiversity"

#put into one dataframe, all stacked up
merge <- rbind(F_weighted_all, C_weighted_all)
merge <- rbind(merge, B_weighted_all)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = Index)) +geom_line(size =1)+
  labs(y = "Mean Index", title = "FCB Potential of Aquaculture Weighted by 54 Major Species")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))+
  ylim(48, 60)


#SAVE
# setwd("/Users/aleahw/Documents/Master's!!/Chapter_2/Figures")
# ggsave("FCB_weightedby_all830.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

###STEP 9: Correlations over time ### NEEDS WORK!!!
```{r correlations of FCB over time}

#reconfigure f
weighted_indexes_graph <- weighted_indexes[, -3]
colnames(weighted_indexes_graph)[2] = "Food_Security"

#reconfigure c
weighted_indexes_graphc <- weighted_indexesc[, -3]
colnames(weighted_indexes_graphc)[2] = "Climate_Change"

#reconfigure b
weighted_indexes_graphb <- weighted_indexesb[, -3]
colnames(weighted_indexes_graphb)[2] = "Biodiversity"

#merge by year
a <- merge(weighted_indexes_graph, weighted_indexes_graphc, by="Year")
a <- merge(a, weighted_indexes_graphb, by="Year")

#find correlations

# Do the data from each of the 2 variables (x, y) follow a normal distribution?
# Use Shapiro-Wilk normality production_data –> R function: shapiro.production_data()
# and look at the normality plot —> R function: ggpubr::ggqqplot()

# Shapiro-Wilk production_data can be performed as follow:
# Null hypothesis: the data are normally distributed
# Alternative hypothesis: the data are not normally distributed

# Shapiro-Wilk normality production_data for f
shapiro.production_data(a$Food_Security) # => p NOT NORMAL p-value = 0.0005529
# Shapiro-Wilk normality production_data for c
shapiro.production_data(a$Climate_Change) # => p NOT NORMAL p-value = 0.006915
# Shapiro-Wilk normality production_data for b
shapiro.production_data(a$Biodiversity) # => NORMAL: p = 0.01736

# Visual inspection of the data normality using Q-Q plots (quantile-quantile plots). Q-Q plot draws the correlation between a given sample and the normal distribution.
library(ggpubr)
ggqqplot(a$Food_Security, ylab = "Food Security")
ggqqplot(a$Climate_Change, ylab = "Climate Change")
ggqqplot(a$Biodiversity, ylab = "Biodiversity")

#histogram to see dist
ggdensity(a$Food_Security, 
          main = "Density plot of F index",
          xlab = "Food Security Index")

ggdensity(a$Climate_Change, 
          main = "Density plot of C index",
          xlab = "Climate Change Index")

ggdensity(a$Biodiversity, 
          main = "Density plot of B index",
          xlab = "Biodiversity Index")


#PEARSON, assuming normal distribution...
#FOOD SECURITY AND CLIMATE CHANGE
ggscatter(a, x = "Food_Security", y = "Climate_Change", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Food Security Index", ylab = "Climate Change Index") #SIGNIFICANT, R = 0.92; p = < 2.2 e-16

res <- cor.production_data(a$Food_Security, a$Climate_Change, 
                    method = "pearson")
res

#FOOD SECURITY AND BIODIVERSITY
ggscatter(a, x = "Food_Security", y = "Biodiversity", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Food Security Index", ylab = "Biodiversity Index") #SIGNIFICANT, R = 0.95; p = < 2.2 e-16

res <- cor.production_data(a$Food_Security, a$Biodiversity, 
                    method = "pearson")
res

#BIODIVERSITY AND CLIMATE CHANGE
ggscatter(a, x = "Climate_Change", y = "Biodiversity", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Climate Change Index", ylab = "Biodiversity Index") #SIGNIFICANT, R = 0.79; p = 3.086e-16

res <- cor.production_data(a$Biodiversity, a$Climate_Change, 
                    method = "pearson")
res

#Spearman rank correlation, assuming non-normal dist
# The Kendall rank correlation coefficient or Kendall’s tau statistic is used to estimate a rank-based measure of association. This production_data may be used if the data do not necessarily come from a bivariate normal distribution.

res2 <- cor.production_data(a$Food_Security, a$Climate_Change,  method="kendall") #SIG; tau = 0.6697966, p-value < 2.2e-16
res2

res2 <- cor.production_data(a$Food_Security, a$Biodiversity,  method="kendall") #SIG; tau = 0.5547731; p-value = 5.453e-12
res2

res2 <- cor.production_data(a$Biodiversity, a$Climate_Change,  method="kendall") #SIG; tau = 0.7519562; p-value < 2.2e-16
res2
```


CALCULATING DIVERGENCE or CONVERGENCE
```{r}

#make five time slices, 13 to 14 years each

F1 <- filter(merge, Year < 1965, Index == "Food Security")
F2 <- filter(merge, Year > 1964 & Year < 1979, Index == "Food Security")
F3 <- filter(merge, Year > 1978 & Year < 1993, Index == "Food Security")
F4 <- filter(merge, Year > 1992 & Year < 2007, Index == "Food Security")
F5 <- filter(merge, Year > 2006, Index == "Food Security")

C1 <- filter(merge, Year < 1965, Index == "Climate Change")
C2 <- filter(merge, Year > 1964 & Year < 1979, Index == "Climate Change")
C3 <- filter(merge, Year > 1978 & Year < 1993, Index == "Climate Change")
C4 <- filter(merge, Year > 1992 & Year < 2007, Index == "Climate Change")
C5 <- filter(merge, Year > 2006, Index == "Climate Change")

B1 <- filter(merge, Year < 1965, Index == "Biodiversity")
B2 <- filter(merge, Year > 1964 & Year < 1979, Index == "Biodiversity")
B3 <- filter(merge, Year > 1978 & Year < 1993, Index == "Biodiversity")
B4 <- filter(merge, Year > 1992 & Year < 2007, Index == "Biodiversity")
B5 <- filter(merge, Year > 2006, Index == "Biodiversity")

#make new frame to store values
divergence = data.frame(matrix(
  vector(), 5, 3, dimnames=list(c(), c("B-F","B-C","F-C"))),
                stringsAsFactors=F)
rownames(divergence) <- c("1950-1964","1965-1978","1979-1992","1993-2006", "2007-2021")

#subtract and add values to dataframe
# divergence$B.F[1] <- B1-F1

BF1 <- data.frame(B1$Score - F1$Score)
BF2 <- data.frame(B2$Score - F2$Score)
BF3 <- data.frame(B3$Score - F3$Score)
BF4 <- data.frame(B4$Score - F4$Score)
BF5 <- data.frame(B5$Score - F5$Score)

BC1 <- data.frame(B1$Score - C1$Score)
BC2 <- data.frame(B2$Score - C2$Score)
BC3 <- data.frame(B3$Score - C3$Score)
BC4 <- data.frame(B4$Score - C4$Score)
BC5 <- data.frame(B5$Score - C5$Score)

FC1 <- data.frame(F1$Score - C1$Score)
FC2 <- data.frame(F2$Score - C2$Score)
FC3 <- data.frame(F3$Score - C3$Score)
FC4 <- data.frame(F4$Score - C4$Score)
FC5 <- data.frame(F5$Score - C5$Score)

slice1 <- cbind(BF1, BC1, FC1) #1950-1964
divergence$B.F[1] <- mean(slice1$B1.Score...F1.Score)
divergence$B.C[1] <- mean(slice1$B1.Score...C1.Score)
divergence$F.C[1] <- mean(slice1$F1.Score...C1.Score)

slice2 <- cbind(BF2, BC2, FC2)
divergence$B.F[2] <- mean(slice2$B2.Score...F2.Score)
divergence$B.C[2] <-mean(slice2$B2.Score...C2.Score)
divergence$F.C[2] <- mean(slice2$F2.Score...C2.Score)

slice3 <- cbind(BF3, BC3, FC3)
divergence$B.F[3] <- mean(slice3$B3.Score...F3.Score)
divergence$B.C[3] <- mean(slice3$B3.Score...C3.Score)
divergence$F.C[3] <- mean(slice3$F3.Score...C3.Score)

slice4 <- cbind(BF4, BC4, FC4)
divergence$B.F[4] <- mean(slice4$B4.Score...F4.Score)
divergence$B.C[4] <- mean(slice4$B4.Score...C4.Score)
divergence$F.C[4] <- mean(slice4$F4.Score...C4.Score)

slice5 <- cbind(BF5, BC5, FC5)
divergence$B.F[5] <- mean(slice5$B5.Score...F5.Score)
divergence$B.C[5] <- mean(slice5$B5.Score...C5.Score)
divergence$F.C[5] <- mean(slice5$F5.Score...C5.Score)

#SAVE

NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Fuzzy/Figures/divergence2.csv")
write.csv(divergence,file=NAME)
```

```{r COMBINED FCB GRAPH WEIGHTED BY FINFISH}

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each FINFISH species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine") #filter out just the finfish

production_data2 <- filter(production_data, category =="finfish_inland"| category =="finfish_marine")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category
  index <- index/(weighting_totals[2, b]) #weight by production of finfish
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year
for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric


#CLIMATE CHANGE
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine") #filter out just the finfish

production_data2 <- filter(production_data, category =="finfish_inland"| category =="finfish_marine")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$c[i] #get the c score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category
  index <- index/(weighting_totals[2, b]) #weight by production of finfish
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year
for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesc <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesc <- pivot_longer(weighted_indexesc, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesc <- weighted_indexesc[order(weighted_indexesc$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesc)) {
  print(weighted_indexesc$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesc$Year[i] <- year_labs[i]
}

weighted_indexesc$Year <- as.numeric(weighted_indexesc$Year) #make years numeric


#BIODIVERSITY
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine") #filter out just the finfish

production_data2 <- filter(production_data, category =="finfish_inland"| category =="finfish_marine")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$b[i] #get the b score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category
  index <- index/(weighting_totals[2, b]) #weight by production of finfish
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year
for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesb <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesb <- pivot_longer(weighted_indexesb, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesb <- weighted_indexesb[order(weighted_indexesb$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesb)) {
  print(weighted_indexesb$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesb$Year[i] <- year_labs[i]
}

weighted_indexesb$Year <- as.numeric(weighted_indexesb$Year) #make years numeric

#GRAPHING
#weighted_indexes = F weighted by all
weighted_indexes$Index <- "Food Security"
#weighted_indexesc = C weighted by all 
weighted_indexesc$Index <- "Climate Change"
#weighted_indexesb = B weighted by all 
weighted_indexesb$Index <- "Biodiversity"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexesc)
merge <- rbind(merge, weighted_indexesb)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = Index)) +geom_line(size =1)+
  labs(y = "Mean Index", title = "FCB Potential of Aquaculture Weighted by Finfish")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))+
        ylim(45, 54)


#SAVE
# ggsave("FCB_weightedby_finfish.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

```{r COMBINED FCB GRAPH WEIGHTED BY ALGAE}

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "algae") #filter out just the algae

production_data2 <- filter(production_data, category =="algae")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[4, b]) #weight by production of algae, row 4
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#CLIMATE CHANGE
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "algae") #filter out just the algae

production_data2 <- filter(production_data, category =="algae")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$c[i] #get the c score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[4, b]) #weight by production of algae, row 4
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesc <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesc <- pivot_longer(weighted_indexesc, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesc <- weighted_indexesc[order(weighted_indexesc$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesc)) {
  print(weighted_indexesc$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesc$Year[i] <- year_labs[i]
}

weighted_indexesc$Year <- as.numeric(weighted_indexesc$Year) #make years numeric

#BIODIVERSITY
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "algae") #filter out just the algae

production_data2 <- filter(production_data, category =="algae")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$b[i] #get the b score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[4, b]) #weight by production of algae, row 4
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesb <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesb <- pivot_longer(weighted_indexesb, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesb <- weighted_indexesb[order(weighted_indexesb$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesb)) {
  print(weighted_indexesb$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesb$Year[i] <- year_labs[i]
}

weighted_indexesb$Year <- as.numeric(weighted_indexesb$Year) #make years numeric

#GRAPHING
#weighted_indexes = F weighted by all
weighted_indexes$Index <- "Food Security"
#weighted_indexesc = C weighted by all 
weighted_indexesc$Index <- "Climate Change"
#weighted_indexesb = B weighted by all 
weighted_indexesb$Index <- "Biodiversity"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexesc)
merge <- rbind(merge, weighted_indexesb)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = Index)) +geom_line(size =1)+
  labs(y = "Mean Index", title = "FCB Potential of Aquaculture Weighted by Algae")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))+
  ylim(53, 67)


#SAVE
# ggsave("FCB_weightedby_algae.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set

```

```{r COMBINED FCB GRAPH WEIGHTED BY CRUSTACEANS}

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "crustacean")

production_data2 <- filter(production_data, category == "crustacean")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[5, b]) #weight by production of crustaceans
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#CLIMATE CHANGE
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "crustacean")

production_data2 <- filter(production_data, category == "crustacean")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$c[i] #get the c score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[5, b]) #weight by production of crustaceans
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesc <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesc <- pivot_longer(weighted_indexesc, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesc <- weighted_indexesc[order(weighted_indexesc$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesc)) {
  print(weighted_indexesc$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesc$Year[i] <- year_labs[i]
}

weighted_indexesc$Year <- as.numeric(weighted_indexesc$Year) #make years numeric

#BIODIVERSITY
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "crustacean")

production_data2 <- filter(production_data, category == "crustacean")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$b[i] #get the b score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[5, b]) #weight by crustaceans
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesb <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesb <- pivot_longer(weighted_indexesb, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesb <- weighted_indexesb[order(weighted_indexesb$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesb)) {
  print(weighted_indexesb$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesb$Year[i] <- year_labs[i]
}

weighted_indexesb$Year <- as.numeric(weighted_indexesb$Year) #make years numeric

#GRAPHING
#weighted_indexes = F weighted by all
weighted_indexes$Index <- "Food Security"
#weighted_indexesc = C weighted by all 
weighted_indexesc$Index <- "Climate Change"
#weighted_indexesb = B weighted by all 
weighted_indexesb$Index <- "Biodiversity"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexesc)
merge <- rbind(merge, weighted_indexesb)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = Index)) +geom_line(size =1)+
  labs(y = "Mean Index", title = "FCB Potential of Aquaculture Weighted by Crustaceans")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))+
  ylim(30, 55)


#SAVE
# ggsave("FCB_weightedby_crustaceans.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

```{r COMBINED FCB GRAPH WEIGHTED BY MOLLUSCS}

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "mollusc")

production_data2 <- filter(production_data, category == "mollusc")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[6, b]) #weight by production of molluscs
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#CLIMATE CHANGE
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "mollusc")

production_data2 <- filter(production_data, category == "mollusc")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$c[i] #get the c score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[6, b]) #weight by production of molluscs
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesc <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesc <- pivot_longer(weighted_indexesc, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesc <- weighted_indexesc[order(weighted_indexesc$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesc)) {
  print(weighted_indexesc$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesc$Year[i] <- year_labs[i]
}

weighted_indexesc$Year <- as.numeric(weighted_indexesc$Year) #make years numeric

#BIODIVERSITY
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "mollusc")

production_data2 <- filter(production_data, category == "mollusc")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$b[i] #get the b score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[6, b]) #weight by molluscs
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesb <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesb <- pivot_longer(weighted_indexesb, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesb <- weighted_indexesb[order(weighted_indexesb$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesb)) {
  print(weighted_indexesb$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesb$Year[i] <- year_labs[i]
}

weighted_indexesb$Year <- as.numeric(weighted_indexesb$Year) #make years numeric

#GRAPHING
#weighted_indexes = F weighted by all
weighted_indexes$Index <- "Food Security"
#weighted_indexesc = C weighted by all 
weighted_indexesc$Index <- "Climate Change"
#weighted_indexesb = B weighted by all 
weighted_indexesb$Index <- "Biodiversity"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexesc)
merge <- rbind(merge, weighted_indexesb)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = Index)) +geom_line(size =1)+
  labs(y = "Mean Index", title = "FCB Potential of Aquaculture Weighted by Molluscs")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))+
  ylim(44, 66)



#SAVE
# ggsave("FCB_weightedby_molluscs.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```


CHECKING FOR CORRELATION
```{r}
library("ggpubr")
library(ggplot2)

# Do the data from each of the 2 variables (x, y) follow a normal distribution?
# Use Shapiro-Wilk normality production_data –> R function: shapiro.production_data()
# and look at the normality plot —> R function: ggpubr::ggqqplot()

# Shapiro-Wilk production_data can be performed as follow:
# Null hypothesis: the data are normally distributed
# Alternative hypothesis: the data are not normally distributed

# Shapiro-Wilk normality production_data for f
shapiro.production_data(speciesdata$f) # => p-value = 0.00916 NOT NORMAL
# Shapiro-Wilk normality production_data for c
shapiro.production_data(speciesdata$c) # => p-value = 0.109 NORMAL
# Shapiro-Wilk normality production_data for b
shapiro.production_data(speciesdata$b) # => p-value = 0.01076 NORMAL 

# Visual inspection of the data normality using Q-Q plots (quantile-quantile plots). Q-Q plot draws the correlation between a given sample and the normal distribution.

ggqqplot(speciesdata$f, ylab = "Food Security")
ggqqplot(speciesdata$c, ylab = "Climate Change")
ggqqplot(speciesdata$b, ylab = "Biodiversity")

#histogram to see dist
ggdensity(speciesdata$f, 
          main = "Density plot of F index",
          xlab = "Food Security Index")

ggdensity(speciesdata$c, 
          main = "Density plot of C index",
          xlab = "Climate Change Index")

ggdensity(speciesdata$b, 
          main = "Density plot of B index",
          xlab = "Biodiversity Index")


#PEARSON, assuming normal distribution...
#FOOD SECURITY AND CLIMATE CHANGE
ggscatter(speciesdata, x = "f", y = "c", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Food Security Index", ylab = "Climate Change Index") #SIGNIFICANT, R = 0.48; p = 0.0001504

res <- cor.production_data(speciesdata$f, speciesdata$c, 
                    method = "pearson")
res

#FOOD SECURITY AND BIODIVERSITY
ggscatter(speciesdata, x = "f", y = "b", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Food Security Index", ylab = "Biodiversity Index") #SIGNIFICANT, R = 0.44; p = 0.0005818

res <- cor.production_data(speciesdata$f, speciesdata$b, 
                    method = "pearson")
res

#BIODIVERSITY AND CLIMATE CHANGE
ggscatter(speciesdata, x = "c", y = "b", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Climate Change Index", ylab = "Biodiversity Index") #SIGNIFICANT, R = 0.48; p = 0.000145

res <- cor.production_data(speciesdata$c, speciesdata$b, 
                    method = "pearson")
res

#Spearman rank correlation, assuming non-normal dist
# The Kendall rank correlation coefficient or Kendall’s tau statistic is used to estimate a rank-based measure of association. This production_data may be used if the data do not necessarily come from a bivariate normal distribution.

res2 <- cor.production_data(speciesdata$f, speciesdata$c,  method="kendall") #SIG; tau = 0.4261061, p-value = 2.944e-06
res2

res2 <- cor.production_data(speciesdata$f, speciesdata$b,  method="kendall") #SIG; tau = 0.3795163; p-value = 3.199e-05; z = 4.9297
res2

res2 <- cor.production_data(speciesdata$c, speciesdata$b,  method="kendall") #SIG; tau = 0.359071; p-value = 8.212e-05
res2

```

#################################################
Percent contribution to index
#make sure to run from line 565 (for food security) so you can use "weighted_indexes_1 etc" or from 878 for climate change, or 1183 for biodiv

#new approach: % contribution of taxonomic groups relative to F, C, B relative to the proportion of total production
#Make three graphs, one for each FCB. For each taxonomic group, divide mean F/total production etc, plot all taxonomic groups on the same graph. Y-axis will be 0 to 1; don't plot production line, just the ratio.

```{r}
#first, calculate the % contribution of each taxa to annual production. We will eventually divide FCB indices by this proportion to find a ratio.

weighting_totals4 <- weighting_totals[-c(3), ]

new_frame <- weighting_totals4 #create new frame
new_frame <- new_frame[FALSE, ]

#divide taxa production in each year by the total prod in that year

for (i in 2:nrow(weighting_totals4)){ #for each taxonomic group
    for (b in 2:73){ #for each year
      total_prod <- weighting_totals4[1, b] #get the total prod in that year
      index <- weighting_totals4[i, b]/total_prod*100 #divide score by that species' total prod
      new_frame[i, b] <- index   #add these numbers to the frame
      
  } 
}

#delete first row
new_frame <- new_frame[-c(1), ]
#rename
new_frame[1, 1] <-"Finfish"
new_frame[2, 1] <-"Algae"
new_frame[3, 1] <-"Crustacean"
new_frame[4, 1] <-"Mollusc"

#make long?
new_frame <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Perc_prod")



#fix years
year_labs <- rep(2021:1950, times=6) 


#FIX from here
#rename year column in weighted_indexes
for (i in 1:nrow(new_frame)) {
  print(new_frame$Year[i])
  #replace each year with the actual year number from year_labs
  
  new_frame$Year[i] <- year_labs[i]
}


#save this
totalproduction <- new_frame
#goes one taxa at a time, from 2021 to 1950

#NEED TO PUT IN year order??
totalproduction <- totalproduction %>% arrange(Year)
```

```{r}
#now, calculate the percentage contribution of each taxa to F, C and B in each year.

#FOOD00 from line 570
#CLIMATE--run from 920
#CLIMATE--run from 1258

#weighting totals has an extra row: get rid of it here
weighting_totals3 <- weighting_totals[-3, ]
#also put in reverse order
weighting_totals3 <- rev(weighting_totals3)

#merge data
all_taxa <- merge(weighted_indexes1, weighted_indexes2, by="Year")
all_taxa <- merge(all_taxa, weighted_indexes3, by="Year")
all_taxa <- merge(all_taxa, weighted_indexes4, by="Year")
all_taxa <- merge(all_taxa, weighted_indexes, by="Year")

all_taxa <- all_taxa[ -c(3, 5, 7, 9, 11) ]

#rename
colnames(all_taxa)[2] <- "Finfish"
colnames(all_taxa)[3] <- "Algae"
colnames(all_taxa)[4] <- "Crustaceans"
colnames(all_taxa)[5] <- "Molluscs"
colnames(all_taxa)[6] <- "All"

new_frame <- all_taxa[, -c(6)] #create new frame
# new_frame <- new_frame[FALSE, ]

#for each taxa
for (i in 2:5) {
  # print(weighting_totals3$species[i])
 for (y in 1:72){ #for each year
   taxa_perc <- weighting_totals3[i, y]/weighting_totals3[1,y]
   # print(weighting_totals3[i, y])
   # print(taxa_perc)
   product <- taxa_perc*all_taxa[y, i] #multiply the index of the taxa by taxa_perc
   # print(all_taxa[y, i])
   # print(product)
   product <- product/all_taxa[y, 6] #fix!!!!!!!?
   print(product) #put this number in a dataframe?
   
   new_frame[y, i] <- product*100
   
 }
}

#make long?
new_frame <- pivot_longer(new_frame, cols = -c("Year"),
                                names_to = "species", values_to = "index_perc")

#save this
totalindex <- new_frame
```

```{r}
#combine totalindex and totalprod
prod_index <- cbind(totalindex, totalproduction)
prod_index <- prod_index[, -c(4, 5)]

#new column divides perc contribution to index, by perc contribution to production
prod_index[, 5] <- "new"
colnames(prod_index)[5] <- "ratio"

#calculate ratio for every row
for (i in 1:nrow(prod_index)) {
  ratio <- prod_index$index_perc[i]/prod_index$Perc_prod[i]
  prod_index$ratio[i] <- ratio
}

#graph this:
plot <- ggplot(prod_index, aes(x=Year, y=as.numeric(ratio) , color=species)) + geom_point()+
      geom_hline(yintercept = 1)+
      labs(y = "Contribution to B index:contribution to production")
      # ylim(0, 1.5)

#SAVE
# setwd("/Users/aleahw/Documents/Master's!!/Chapter_2")
# ggsave("Ratio_B_prod.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set

```


```{r}
#graph
#make long
new_frame <- pivot_longer(new_frame, cols = -c("Year"),
                                names_to = "Taxa", values_to = "Score")

# weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

#specify order of bars (from top to bottom)
new_frame$Taxa <- factor(new_frame$Taxa, levels=c('Algae','Crustaceans', 'Molluscs','Finfish'))

#create stacked bar chart
ggplot(new_frame, aes(x=Year, y=Score, fill=Taxa)) + 
    geom_bar(position='stack', stat='identity')+
     labs(y = "Percent contribution to F index")

# weighting_totals4 <- pivot_wider(merge, names_from = "Year", values_from = "Score")
# 
# new_frame <- weighting_totals4 #create new frame
# new_frame <- new_frame[FALSE, ]
```


#OLD
```{r OLD}

#weighting totals has an extra row: get rid of it here
weighting_totals3 <- weighting_totals[-3, ]
#also put in reverse order
weighting_totals3 <- rev(weighting_totals3)

#merge data
all_taxa <- merge(weighted_indexes1, weighted_indexes2, by="Year")
all_taxa <- merge(all_taxa, weighted_indexes3, by="Year")
all_taxa <- merge(all_taxa, weighted_indexes4, by="Year")
all_taxa <- merge(all_taxa, weighted_indexes, by="Year")

all_taxa <- all_taxa[ -c(3, 5, 7, 9, 11) ]

#rename
colnames(all_taxa)[2] <- "Finfish"
colnames(all_taxa)[3] <- "Algae"
colnames(all_taxa)[4] <- "Crustaceans"
colnames(all_taxa)[5] <- "Molluscs"
colnames(all_taxa)[6] <- "All"

new_frame <- all_taxa[, -c(6)] #create new frame
# new_frame <- new_frame[FALSE, ]

#for each taxa
for (i in 2:5) {
  # print(weighting_totals3$species[i])
 for (y in 1:72){ #for each year
   taxa_perc <- weighting_totals3[i, y]/weighting_totals3[1,y]
   # print(weighting_totals3[i, y])
   # print(taxa_perc)
   product <- taxa_perc*all_taxa[y, i] #multiply the index of the taxa by taxa_perc
   # print(all_taxa[y, i])
   # print(product)
   product <- product/all_taxa[y, 6] #fix!!!!!!!?
   print(product) #put this number in a dataframe?
   
   new_frame[y, i] <- product
 }
}

#graph
#make long
new_frame <- pivot_longer(new_frame, cols = -c("Year"),
                                names_to = "Taxa", values_to = "Score")

# weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

#specify order of bars (from top to bottom)
new_frame$Taxa <- factor(new_frame$Taxa, levels=c('Algae','Crustaceans', 'Molluscs','Finfish'))

#create stacked bar chart
ggplot(new_frame, aes(x=Year, y=Score, fill=Taxa)) + 
    geom_bar(position='stack', stat='identity')+
     labs(y = "Percent contribution to F index")


# 
# ggplot(new_frame, aes(fill=Taxa, y=Score, x=Year)) + 
#     geom_bar(position="fill", stat="identity")+
#     labs(y = "Percent contribution to B index")+
#       scale_y_continuous(labels = scales::percent) +
#         geom_text(aes(label = paste0(Score*100,"%")),
#                   position = position_stack(vjust = 0.5), size = 0.5)
              
#change colors?

#SAVE
# setwd("/Users/aleahw/Documents/Master's!!/Fuzzy/production_totals")
# ggsave("percent_contribution_B2.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

```{r}
#compared to a stacked % barchart of just production

#delete "all" from weighting totals 2
weighting_totals2  <- weighting_totals2 %>%
  filter(species != "all species")

#specify order of bars (from top to bottom)
weighting_totals2$species <- factor(weighting_totals2$species, levels=c('algae','crustacean', 'mollusc','finfish'))

ggplot(weighting_totals2, aes(fill=species, y=Score, x=Year)) + 
    geom_bar(position="fill", stat="identity")+
    labs(y = "Percent contribution to total production")

# ggsave("percent_contribution_prod2.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

#########################################
#Minimum and Maximum FCB index of each taxa weighted by total production of taxa
```{r}
#find minimum and maximum FCB index of each species

#finfish:
speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine")

production_data2 <- filter(production_data, category =="finfish_inland"| category =="finfish_marine")

fmax <- subset(speciesdata2, f == max(f))
fmin <- subset(speciesdata2, f == min(f))

cmax <- subset(speciesdata2, c == max(c))
cmin <- subset(speciesdata2, c == min(c))

bmax <- subset(speciesdata2, b == max(b))
bmin <- subset(speciesdata2, b == min(b))

#combine in a dataframe
speciesdata2 <- rbind(fmax, fmin) #change for c and b

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each FINFISH species and multiply f score by (production of finfish /production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*weighting_totals[2, b] #multiply score by the taxa's total production in each year, b is 2 which works because the year starts in the second column
  index <- index/(weighting_totals[1, b]) #weight by production of all species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

library(tidyr)
#make long
weighted_indexes <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 2 times
year_labs <- rep(vec, each = 2) #2 species here

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#graph
# ggplot(weighted_indexes, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
#   labs(y = "Index", title = "Minimum and Maximum FCB Potential of Finfish")+
#     theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
#       scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))

#SAVE
# ggsave("FCB_weightedby_molluscs.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set


#ALGAE
speciesdata2 <- filter(speciesdata, category == "algae")

production_data2 <- filter(production_data, category == "algae")

fmax <- subset(speciesdata2, f == max(f))
fmin <- subset(speciesdata2, f == min(f))

cmax <- subset(speciesdata2, c == max(c))
cmin <- subset(speciesdata2, c == min(c))

bmax <- subset(speciesdata2, b == max(b))
bmin <- subset(speciesdata2, b == min(b))

#combine in a dataframe
speciesdata2 <- rbind(fmax, fmin)

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each ALGAE species and multiply f score by (production of algae /production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*weighting_totals[4, b] #multiply score by the taxa's total production in each year, b is 2 which works because the year starts in the second column
  index <- index/(weighting_totals[1, b]) #weight by production of all species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

library(tidyr)
#make long
weighted_indexes_algae <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Score")

weighted_indexes_algae <- weighted_indexes_algae[order(weighted_indexes_algae$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 2 times
year_labs <- rep(vec, each = 2) #2 species here

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes_algae)) {
  #replace each year with the actual year number from year_labs
  
  weighted_indexes_algae$Year[i] <- year_labs[i]
}

weighted_indexes_algae$Year <- as.numeric(weighted_indexes_algae$Year) #make years numeric


#MOLLUSCS
speciesdata2 <- filter(speciesdata, category == "mollusc")

production_data2 <- filter(production_data, category == "mollusc")

fmax <- subset(speciesdata2, f == max(f))
fmin <- subset(speciesdata2, f == min(f))

cmax <- subset(speciesdata2, c == max(c))
cmin <- subset(speciesdata2, c == min(c))

bmax <- subset(speciesdata2, b == max(b))
bmin <- subset(speciesdata2, b == min(b))

#combine in a dataframe
speciesdata2 <- rbind(fmax, fmin)

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each mollusc species and multiply f score by (production of mollusc /production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*weighting_totals[6, b] #multiply score by the taxa's total production in each year, b is 2 which works because the year starts in the second column
  index <- index/(weighting_totals[1, b]) #weight by production of all species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

library(tidyr)
#make long
weighted_indexes_moll <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Score")

weighted_indexes_moll <- weighted_indexes_moll[order(weighted_indexes_moll$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 2 times
year_labs <- rep(vec, each = 2) #2 species here

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes_moll)) {
  #replace each year with the actual year number from year_labs
  
  weighted_indexes_moll$Year[i] <- year_labs[i]
}

weighted_indexes_moll$Year <- as.numeric(weighted_indexes_moll$Year) #make years numeric

#CRUSTACEANS
speciesdata2 <- filter(speciesdata, category == "crustacean")

production_data2 <- filter(production_data, category == "crustacean")

fmax <- subset(speciesdata2, f == max(f))
fmin <- subset(speciesdata2, f == min(f))

cmax <- subset(speciesdata2, c == max(c))
cmin <- subset(speciesdata2, c == min(c))

bmax <- subset(speciesdata2, b == max(b))
bmin <- subset(speciesdata2, b == min(b))

#combine in a dataframe
speciesdata2 <- rbind(fmax, fmin)

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each mollusc species and multiply f score by (production of mollusc /production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*weighting_totals[5, b] #multiply score by the taxa's total production in each year, b is 2 which works because the year starts in the second column
  index <- index/(weighting_totals[1, b]) #weight by production of all species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

library(tidyr)
#make long
weighted_indexes_crust <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Score")

weighted_indexes_crust <- weighted_indexes_crust[order(weighted_indexes_crust$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 2 times
year_labs <- rep(vec, each = 2) #2 species here

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes_crust)) {
  #replace each year with the actual year number from year_labs
  
  weighted_indexes_crust$Year[i] <- year_labs[i]
}

weighted_indexes_crust$Year <- as.numeric(weighted_indexes_crust$Year) #make years numeric
```

#combine taxa on the same graph?
```{r}
#MERGE, put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexes_algae, weighted_indexes_moll, weighted_indexes_crust)

#graph
ggplot(merge, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
  labs(y = "Index", title = "Minimum and Maximum Food Potential by Taxa")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+ scale_color_manual(values=c('#7D627F',"orange",'#7D627F',"orange", "pink", "skyblue", "skyblue", "pink"), name = "Taxa")

# 
#   scale_color_manual(name = "Taxa", values=c('pink', 'skyblue', '#7D627F', "orange"), labels = c("Algae", "Crustaceans", "Finfish", "Molluscs"))


```

#OLD
#use species instead of taxa
```{r}
#ALGAE
speciesdata2 <- filter(speciesdata, category == "algae")
production_data2 <- filter(production_data, category == "algae")

fmax <- subset(speciesdata2, f == max(f))
fmin <- subset(speciesdata2, f == min(f))

cmax <- subset(speciesdata2, c == max(c))
cmin <- subset(speciesdata2, c == min(c))

bmax <- subset(speciesdata2, b == max(b))
bmin <- subset(speciesdata2, b == min(b))

#combine in a dataframe
speciesdata2 <- rbind(fmax, fmin)

new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[1, b]) #weight by production of all species, row 1
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

library(tidyr)
#make long
weighted_indexes_algae <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Score")

weighted_indexes_algae <- weighted_indexes_algae[order(weighted_indexes_algae$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 2 times
year_labs <- rep(vec, each = 2) #2 species here

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes_algae)) {
  #replace each year with the actual year number from year_labs
  
  weighted_indexes_algae$Year[i] <- year_labs[i]
}

weighted_indexes_algae$Year <- as.numeric(weighted_indexes_algae$Year) #make years numeric

ggplot(weighted_indexes_algae, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
  labs(y = "Index", title = "Minimum and Maximum FCB Potential of algae")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))
# +
#       scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))

###########################################################
#FINFISH
speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine")

production_data2 <- filter(production_data, category =="finfish_inland"| category =="finfish_marine")

fmax <- subset(speciesdata2, f == max(f))
fmin <- subset(speciesdata2, f == min(f))

cmax <- subset(speciesdata2, c == max(c))
cmin <- subset(speciesdata2, c == min(c))

bmax <- subset(speciesdata2, b == max(b))
bmin <- subset(speciesdata2, b == min(b))

#combine in a dataframe (F)
speciesdata2 <- rbind(fmax, fmin)

new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*production_data2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[1, b]) #weight by production of all species, row 1
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

library(tidyr)
#make long
weighted_indexes <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 2 times
year_labs <- rep(vec, each = 2) #2 species here

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

ggplot(weighted_indexes, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
  labs(y = "Index", title = "Minimum and Maximum food Potential of finfish")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
  ylim(0, 7)
# +
#       scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))

#SAME GRAPH
merge <- rbind(weighted_indexes, weighted_indexes_algae)

#graph
ggplot(merge, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
  labs(y = "Index", title = "Minimum and Maximum Food Security Potential of Finfish and Algae")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(values=c('#7D627F', '#7DBBC3', "pink", "magenta"))

```

#combine taxa on the same graph?
```{r}
#MERGE, put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexes_algae)

#graph
ggplot(merge, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
  labs(y = "Index", title = "Minimum and Maximum Food Security Potential of Finfish and Algae")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(values=c('#7D627F', '#7D627F', "pink", "pink"))
```


