---
title: "Temporal_analysis"
author: "Aleah Wong"
date: '2023-08-06'
output: html_document
---

```{r}
rm(list = ls())
.rs.restartR()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
```

############################################################
- Find 5 year average for last five years and from 1980-1985
  - Kite plot
  - with and without China incuded
- Find each continent's mean score for each of the three major periods
  - Plot Asia with and without China included
  - Dot/boxplot
############################################################

```{r}
#load in FAO data, calculate total production in each year

inland <- read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/inland_finfish_prod.csv")
marine <-read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/marine_finfish_prod.csv")
nonfish <- read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/nonfish_prod.csv")
```

```{r}
# #calculate annual production totals for each inland species
# str(inland)
# inland <- inland %>%mutate_at(c(6:77), as.numeric) #change yearly data to numeric
# 
# # unique(inland$ASFIS.species.Name.En) #find unique species in data
# #create list of each species to loop through
# unique <- c("Bighead carp", "Black carp", "Catla", "Common carp", "Crucian carp", "Grass carp(=White amur)", "Largemouth black bass", "Nile tilapia", "Rainbow trout", "Roho labeo", "Silver carp", "Striped catfish", "Wuchang bream") 
# 
# new_frame <- filter(inland, ASFIS.species.Name.En == unique[1]) #create new frame
# new_frame <- new_frame[FALSE, ]
# 
# for (i in 1:length(unique)){ #for each of the species
#   u_species <- filter(inland, ASFIS.species.Name.En == unique[i])
#   
# #calculate the annual mean each year
# 
#   for (b in 6:77){
#     a <- sum(u_species[, b], na.rm = TRUE)
#     print(a)
#     new_frame[i, 1] <- unique[i]
#     new_frame[i, b] <-a #change row number 
#   }
# }
# 
# new_frame <- new_frame[, c(1, 6:77)] #cut out unnecessary columns
# names(new_frame)[1] <- "Species" #rename first column
# 
# #save this data
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Fuzzy/inland_totals.csv")
# write.csv(new_frame,file=NAME)
```

```{r}
# #calculate annual production totals for each marine species
# # str(marine)
# marine <- marine %>%mutate_at(c(6:77), as.numeric) #change yearly data to numeric
# 
# unique(marine$ASFIS.species.Name.En) #find unique species in data
# 
# #create list of each species to loop through
# unique <- c("Atlantic salmon", "Barramundi(=Giant seaperch)", "Coho(=Silver) salmon", "European seabass", "Flathead grey mullet", "Gilthead seabream", "Japanese amberjack", "Japanese seabass", "Large yellow croaker", "Milkfish", "Nile tilapia", "Orange-spotted grouper", "Red drum") 
# 
# new_frame <- filter(marine, ASFIS.species.Name.En == unique[1]) #create new frame
# new_frame <- new_frame[FALSE, ]
# 
# for (i in 1:length(unique)){ #for each of the species
#   u_species <- filter(marine, ASFIS.species.Name.En == unique[i])
#   
# #calculate the annual mean each year
# 
#   for (b in 6:77){
#     a <- sum(u_species[, b], na.rm = TRUE)
#     print(a)
#     new_frame[i, 1] <- unique[i]
#     new_frame[i, b] <-a #change row number 
#   }
# }
# 
# new_frame <- new_frame[, c(1, 6:77)] #cut out unnecessary columns
# names(new_frame)[1] <- "Species" #rename first column
# 
# #save this data
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Fuzzy/marine_totals.csv")
# write.csv(new_frame,file=NAME)
```

```{r}
# #calculate annual production totals for each nonfish species
# # str(marine)
# nonfish <- nonfish %>%mutate_at(c(6:77), as.numeric) #change yearly data to numeric
# 
# unique(nonfish$ASFIS.species.Name.En) #find unique species in data
# 
# #create list of each species to loop through
# unique <- c("Blood cockle", "Blue mussel", "Chilean mussel", "Chinese mitten crab", "Constricted tagelus", "Elkhorn sea moss", "Eucheuma seaweeds nei", "Fusiform sargassum", "Giant river prawn", "Giant tiger prawn", "Gracilaria seaweeds", "Green mud crab", "Indo-Pacific swamp crab", "Japanese carpet shell", "Japanese kelp", "Laver (Nori)", "Mediterranean mussel", "New Zealand mussel", "Nori nei", "Oriental river prawn", "Pacific cupped oyster", "Red swamp crawfish", "Scallops nei", "Sea mussels nei", "Spiny eucheuma", "Wakame", "Whiteleg shrimp", "Yesso scallop") 
# 
# new_frame <- filter(nonfish, ASFIS.species.Name.En == unique[1]) #create new frame
# new_frame <- new_frame[FALSE, ]
# 
# for (i in 1:length(unique)){ #for each of the species
#   u_species <- filter(nonfish, ASFIS.species.Name.En == unique[i])
#   
# #calculate the annual mean each year
# 
#   for (b in 6:77){
#     a <- sum(u_species[, b], na.rm = TRUE)
#     print(a)
#     new_frame[i, 1] <- unique[i]
#     new_frame[i, b] <-a #change row number 
#   }
# }
# 
# new_frame <- new_frame[, c(1, 6:77)] #cut out unnecessary columns
# names(new_frame)[1] <- "Species" #rename first column
# 
# #save this data
# NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Fuzzy/nonfish_totals.csv")
# write.csv(new_frame,file=NAME)
```

START HERE (above code saved in csv files)
```{r}
#calculating weighted index scores

#load data
setwd("/Users/aleahw/Documents/Master's!!/Fuzzy/production_totals")
# category_totals <- read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/production_totals/category_totals.csv") #this needs to be updated bc it was calculated from other species not used in the analysis. but it is used to create an empty frame?


inland_totals <- read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/production_totals/inland_totals.csv")
marine_totals <- read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/production_totals/marine_totals.csv")
nonfish_totals <- read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/production_totals/nonfish_totals.csv")
nonfish_totals <- nonfish_totals[, c(-75)]

#rename (inland and marine) for nile tilapia and rainbow trout
inland_totals$Species[inland_totals$Species == "Rainbow trout"] <- "Oncorhynchus mykiss (inland)"
inland_totals$Species[inland_totals$Species == "Nile tilapia"] <- "Oreochromis niloticus (inland)"

#rename (inland and marine) for nile tilapia and rainbow trout
marine_totals$Species[marine_totals$Species == "Rainbow trout"] <- "Oncorhynchus mykiss (marine)"
marine_totals$Species[marine_totals$Species == "Nile tilapia"] <- "Oreochromis niloticus (marine)"

#remove random extra column in category_totals
# category_totals <- category_totals[, c(-1)]

#combine the datasets with rowbind
cat_totals <- rbind(inland_totals, marine_totals)
cat_totals <- rbind(cat_totals, nonfish_totals)
cat_totals <- cat_totals[, c(-1)]

#rename pompano, get rid of nori nei, scallops nei and sea mussels nei
cat_totals <- cat_totals[!cat_totals$Species == "Nori nei",]
cat_totals <- cat_totals[!cat_totals$Species == "Scallops nei",]
cat_totals <- cat_totals[!cat_totals$Species == "Sea mussels nei",]
```

```{r}
#put both datasets into the same order so the species line up
#rename the cat_totals species to be scientific names

cat_totals$Species[cat_totals$Species == "American cupped oyster"] <- "Crassostrea virginica"
cat_totals$Species[cat_totals$Species == "North African Catfish"] <- "Clarias gariepinus"
cat_totals$Species[cat_totals$Species == "Pompano"] <- "Trachinotus ovatus"

cat_totals$Species[cat_totals$Species == "Black carp"] <- "Mylopharyngodon piceus"
cat_totals$Species[cat_totals$Species == "Bighead carp"] <- "Hypophthalmichthys nobilis"
cat_totals$Species[cat_totals$Species == "Catla"] <- "Catla catla"
cat_totals$Species[cat_totals$Species == "Common carp"] <- "Cyprinus carpio"
cat_totals$Species[cat_totals$Species == "Crucian carp"] <- "Carassius carassius"
cat_totals$Species[cat_totals$Species == "Grass carp(=White amur)"] <- "Ctenopharyngodon idella"
cat_totals$Species[cat_totals$Species == "Largemouth black bass"] <- "Micropterus salmoides"
cat_totals$Species[cat_totals$Species == "Nile tilapia"] <- "Oreochromis niloticus"
cat_totals$Species[cat_totals$Species == "Rainbow trout"] <- "Oncorhynchus mykiss"
cat_totals$Species[cat_totals$Species == "Roho labeo"] <- "Labeo rohita"
cat_totals$Species[cat_totals$Species == "Silver carp"] <- "Hypophthalmichthys molitrix"
cat_totals$Species[cat_totals$Species == "Striped catfish"] <- "Pangasianodon hypophthalmus"
cat_totals$Species[cat_totals$Species == "Wuchang bream"] <- "Megalobrama amblycephala"
cat_totals$Species[cat_totals$Species == "Atlantic salmon"] <- "Salmo salar"
cat_totals$Species[cat_totals$Species == "Barramundi(=Giant seaperch)"] <- "Lates calcarifer"
cat_totals$Species[cat_totals$Species == "Coho(=Silver) salmon"] <- "Oncorhynchus kisutch"
cat_totals$Species[cat_totals$Species == "European seabass"] <- "Dicentrarchus labrax"
cat_totals$Species[cat_totals$Species == "Flathead grey mullet"] <- "Mugil cephalus"
cat_totals$Species[cat_totals$Species == "Gilthead seabream"] <- "Sparus aurata"
cat_totals$Species[cat_totals$Species == "Japanese amberjack"] <- "Seriola quinqueradiata"
cat_totals$Species[cat_totals$Species == "Japanese seabass"] <-  "Lateolabrax japonicus"
cat_totals$Species[cat_totals$Species == "Large yellow croaker"] <- "Larimichthys croceus"
cat_totals$Species[cat_totals$Species == "Milkfish"] <- "Chanos chanos"
cat_totals$Species[cat_totals$Species == "Nile tilapia"] <- "Oreochromis niloticus"
cat_totals$Species[cat_totals$Species == "Orange-spotted grouper"] <- "Epinephelus coioides"
cat_totals$Species[cat_totals$Species == "Red drum"] <- "Sciaenops ocellatus"
cat_totals$Species[cat_totals$Species == "Blood cockle"] <- "Tegillarca granosa"
cat_totals$Species[cat_totals$Species == "Blue mussel" ] <- "Mytilus edulis"
cat_totals$Species[cat_totals$Species == "Chilean mussel"] <- "Mytilus chilensis"
cat_totals$Species[cat_totals$Species == "Chinese mitten crab" ] <- "Eriocheir sinensis"
cat_totals$Species[cat_totals$Species == "Constricted tagelus"] <- "Sinonovacula constricta"
cat_totals$Species[cat_totals$Species == "Elkhorn sea moss"] <- "Kappaphycus alvarezii"

cat_totals$Species[cat_totals$Species == "Eucheuma seaweeds nei"] <- "Eucheuma seaweeds"
cat_totals$Species[cat_totals$Species == "Fusiform sargassum"] <- "Sargassum fusiforme"
cat_totals$Species[cat_totals$Species == "Giant river prawn"] <- "Macrobrachium rosenbergii"
cat_totals$Species[cat_totals$Species == "Giant tiger prawn"] <- "Penaeus monodon"

# cat_totals$Species[cat_totals$Species == "Gracilaria seaweeds"] <- "Gracilaria gracilis" 
cat_totals$Species[cat_totals$Species == "Green mud crab"] <- "Scylla paramamosain" 
cat_totals$Species[cat_totals$Species == "Indo-Pacific swamp crab"] <- "Scylla serrata"
cat_totals$Species[cat_totals$Species == "Japanese carpet shell"] <- "Ruditapes philippinarum"
cat_totals$Species[cat_totals$Species == "Japanese kelp"] <- "Laminaria japonica"

cat_totals$Species[cat_totals$Species == "Laver (Nori)"] <- "Porphyra"
cat_totals$Species[cat_totals$Species == "Mediterranean mussel"] <- "Mytilus galloprovincialis"
cat_totals$Species[cat_totals$Species == "New Zealand mussel"] <- "Perna canaliculus"

cat_totals$Species[cat_totals$Species == "Oriental river prawn" ] <- "Macrobrachium nipponense"
cat_totals$Species[cat_totals$Species == "Pacific cupped oyster"] <- "Magallana gigas"
cat_totals$Species[cat_totals$Species == "Red swamp crawfish"] <- "Procambarus clarkii"
cat_totals$Species[cat_totals$Species == "Spiny eucheuma"] <- "Eucheuma denticulatum"
cat_totals$Species[cat_totals$Species == "Wakame"] <- "Undaria pinnatifida"
cat_totals$Species[cat_totals$Species == "Whiteleg shrimp"] <- "Penaeus vannamei"
cat_totals$Species[cat_totals$Species == "Yesso scallop" ] <- "Mizuhopecten yessoensis"

#extra 2 porphyra in species data, create triplicate
cat_totals <- rbind(cat_totals, cat_totals[rep(45, 2), ])

#put rows in alphabetical order
cat_totals <- cat_totals[order(cat_totals$Species), ]

#renaming in cat_totals so cat_totals matches spp_list and they can be merged
cat_totals$Species[6] <- "Ctenopharyngodon idellus"

cat_totals$Species[42] <- "Porphrya tenera"
cat_totals$Species[43] <- "Porphyra haitanensis"
cat_totals$Species[44] <- "Porphyra yezoensis"

cat_totals$Species[13] <- "Gracilaria gracilis"
```

```{r}
#check
v <- unique(nonfish$ASFIS.species.Name.En) #29

v <- unique(cat_totals$Species) #57
```

```{r}
#load in the FCB scores data
speciesdata <- read.csv("/Users/aleahw/Documents/Master's!!/Fuzzy/FCB_scores_2024.csv") #updated

#alphabetical order:
speciesdata <- speciesdata[order(speciesdata$species), ]


#rename (inland and marine) for nile tilapia and rainbow trout
speciesdata$species[34] <- "Oncorhynchus mykiss (inland)"
speciesdata$species[35] <- "Oncorhynchus mykiss (marine)"
speciesdata$species[36] <- "Oreochromis niloticus (inland)"
speciesdata$species[37] <- "Oreochromis niloticus (marine)"

#rename Eucheuma cottonii to Eucheuma seaweeds (this entry represents the category)
speciesdata$species[11] <- "Eucheuma seaweeds"
```

MERGE
```{r}
#create species and type from speciesdata, bind with category totals by species

spplist <- speciesdata[, c("species", "category")]
names(cat_totals)[1] <- "species" #check this bc u use "Species" later on...

test <- merge(spplist, cat_totals, by="species")
```


```{r graph production of each species, split into graphs based on taxa}

#first up, the finfish:
finfish_prod <- test[(test$category == "finfish_marine" | test$category == "finfish_inland"),]

#first three species are inland species, next three are marine
# finfish_prod <- test[(test$species == "Ctenopharyngodon idellus" | test$species == "Hypophthalmichthys molitrix" | test$species == "Oreochromis niloticus (inland)" | test$species == "Salmo salar" | test$species == "Chanos chanos" | test$species == "Mugil cephalus"),]

#make long
finfish_prod <- pivot_longer(finfish_prod, cols = -c("species", "category"), 
                                names_to = "Year", values_to = "Production")

finfish_prod <- finfish_prod[order(finfish_prod$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 3 times
year_labs <- rep(vec, each = 6) #6 species here

#rename year column in weighted_indexes
for (i in 1:nrow(finfish_prod)) {
  #replace each year with the actual year number from year_labs
  
  finfish_prod$Year[i] <- year_labs[i]
}

finfish_prod$Year <- as.numeric(finfish_prod$Year) #make years numeric

#graph
plot <- ggplot(finfish_prod, aes(x=Year, y=Production, color = species)) +geom_line(size =1)+
  labs(y = "Tons (Live Weight)", title = "Production of Major Finfish Species Over Time")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))

# #SAVE
ggsave("prod_finfish.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set


#algae:
alg_prod <- test[(test$category == "algae"),]

# alg_prod <- test[(test$species == "Laminaria japonica" | test$species == "Eucheuma seaweeds" | test$species == "Gracilaria gracilis" | test$species == "Undaria pinnatifida"),]

#rename 

#make long
alg_prod <- pivot_longer(alg_prod, cols = -c("species", "category"), 
                                names_to = "Year", values_to = "Production")

alg_prod <- alg_prod[order(alg_prod$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 3 times
year_labs <- rep(vec, each = 4) #4 species here

#rename year column in weighted_indexes
for (i in 1:nrow(alg_prod)) {
  #replace each year with the actual year number from year_labs
  
  alg_prod$Year[i] <- year_labs[i]
}

alg_prod$Year <- as.numeric(alg_prod$Year) #make years numeric



#graph
plot <- ggplot(alg_prod, aes(x=Year, y=Production, color = species)) +geom_line(size =1)+
  labs(y = "Tons (Live Weight)", title = "Production of Major Algae Species Over Time")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))

# #SAVE
ggsave("prod_algae.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set

#molluscs:
moll_prod <- test[(test$category == "mollusc"),]
# moll_prod <- test[(test$species == "Crassostrea virginica" | test$species == "Ruditapes philippinarum" | test$species == "Magallana gigas" | test$species == "Mizuhopecten yessoensis"),]

#make long
moll_prod <- pivot_longer(moll_prod, cols = -c("species", "category"), 
                                names_to = "Year", values_to = "Production")

moll_prod <- moll_prod[order(moll_prod$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 3 times
year_labs <- rep(vec, each = 4) #4 species here

#rename year column in weighted_indexes
for (i in 1:nrow(moll_prod)) {
  #replace each year with the actual year number from year_labs
  
  moll_prod$Year[i] <- year_labs[i]
}

moll_prod$Year <- as.numeric(moll_prod$Year) #make years numeric

#graph
plot <- ggplot(moll_prod, aes(x=Year, y=Production, color = species)) +geom_line(size =1)+
  labs(y = "Tons (Live Weight)", title = "Production of Major Mollusc Species Over Time")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))

# #SAVE
ggsave("prod_mollusc.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set


#crustaceans:
crust_prod <- test[(test$category == "crustacean"),]
# 
# crust_prod <- test[(test$species == "Penaeus vannamei" | test$species == "Procambarus clarkii" | test$species == "Eriocheir sinensis"),]

#make long
crust_prod <- pivot_longer(crust_prod, cols = -c("species", "category"), 
                                names_to = "Year", values_to = "Production")

crust_prod <- crust_prod[order(crust_prod$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 3 times
year_labs <- rep(vec, each = 3) #3 species here

#rename year column in weighted_indexes
for (i in 1:nrow(crust_prod)) {
  #replace each year with the actual year number from year_labs
  
  crust_prod$Year[i] <- year_labs[i]
}

crust_prod$Year <- as.numeric(crust_prod$Year) #make years numeric

#graph
plot <- ggplot(crust_prod, aes(x=Year, y=Production, color = species)) +geom_line(size =1)+
  labs(y = "Tons (Live Weight)", title = "Production of Major Crustacean Species Over Time")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))

# #SAVE
ggsave("prod_crust.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

CALCULATING PRODUCTION BY ALL, then by INVERTS, then PLANTS, then FISH
```{r}

frame <- test[, c(1, 3:74)]

weighting_totals <- frame #create new frame
weighting_totals <- weighting_totals[FALSE, ]

#use "cat_totals" for total of all spp
for (i in 2:ncol(cat_totals)){ #for each year
  # print(test[1, i])
  #sum the columns into a new row
  a <- sum(cat_totals[, i], na.rm = TRUE)
  print(a)
  weighting_totals[1, i] <- a
  
}
#rename
weighting_totals[1, 1] <- "all species" 

#Calculate finfish totals
#filter out just the finfish from "test" dataframe
test2 <- filter(test, category =="finfish_inland"| category =="finfish_marine")

#remove the category column so that this dataframe and the weighting_totals dataframe line up
test2 <- test2[-c(2)]

for (i in 2:ncol(test2)){ #for each year
  #sum the columns into a new row
  a <- sum(test2[, i], na.rm = TRUE)
  print(a)
  weighting_totals[2, i] <- a #second row for this total
  
}

weighting_totals[2, 1] <- "finfish" #rename

#Calculate invert totals
#filter out just the inverts
test2 <- filter(test, category =="mollusc"| category =="crustacean")

#remove the category column so that this dataframe and the weighting_totals dataframe line up
test2 <- test2[-c(2)]

for (i in 2:ncol(test2)){ #for each year
  #sum the columns into a new row
  a <- sum(test2[, i], na.rm = TRUE)
  print(a)
  weighting_totals[3, i] <- a #third row for the inverts total
  
}

weighting_totals[3, 1] <- "inverts" #rename

#Calculate algae totals
#filter out just the algae
test2 <- filter(test, category =="algae")

#remove the category column so that this dataframe and the weighting_totals dataframe line up
test2 <- test2[-c(2)]

for (i in 2:ncol(test2)){ #for each year
  #sum the columns into a new row
  a <- sum(test2[, i], na.rm = TRUE)
  print(a)
  weighting_totals[4, i] <- a #fourth row for the algae total
  
}

weighting_totals[4, 1] <- "algae" #rename

#Calculate crustacean totals
#filter out just the crustaceans
test2 <- filter(test, category =="crustacean")

#remove the category column so that this dataframe and the weighting_totals dataframe line up
test2 <- test2[-c(2)]

for (i in 2:ncol(test2)){ #for each year
  #sum the columns into a new row
  a <- sum(test2[, i], na.rm = TRUE)
  print(a)
  weighting_totals[5, i] <- a #fifth row for the crustacean total
  
}

weighting_totals[5, 1] <- "crustacean" #rename



#Calculate mollusc totals
#filter out just the molluscs
test2 <- filter(test, category =="mollusc")

#remove the category column so that this dataframe and the weighting_totals dataframe line up
test2 <- test2[-c(2)]

for (i in 2:ncol(test2)){ #for each year
  #sum the columns into a new row
  a <- sum(test2[, i], na.rm = TRUE)
  print(a)
  weighting_totals[6, i] <- a #sixth row for the mollusc total
  
}

weighting_totals[6, 1] <- "mollusc" #rename
```

GRAPH production over time by taxa
```{r}

library(tidyr)
#make long
weighting_totals2 <- pivot_longer(weighting_totals, cols = -c(species), 
                                names_to = "Year", values_to = "Score")

year_labs <- rep(2021:1950, times=6) 

#rename year column in weighted_indexes
for (i in 1:nrow(weighting_totals2)) {
  print(weighting_totals2$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighting_totals2$Year[i] <- year_labs[i]
}

weighting_totals2$Year <- as.numeric(weighting_totals2$Year) #make years numeric
weighting_totals2 <- weighting_totals2[order(weighting_totals2$Year),] #put earlier years first

#remove inverts
weighting_totals2 <- weighting_totals2[!(weighting_totals2$species == "inverts"),]

#graph
plot<- ggplot(weighting_totals2, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
  labs(y = "Tons (Live Weight)", title = "Aquaculture Production over Time")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Taxa", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))

# #SAVE
# ggsave("prod_over_time_bytaxa.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set

model <- lm(Score ~ Year, data = weighting_totals2)
summary(model)

#calculate growth rates
```

##########################################
FOOD SECURITY WEIGHTED BY ALL PRODUCTION
```{r}
#calculate index scores for each species weighted by production of ALL 55 species

new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata)){ #for each species
  print(speciesdata$species[i])
  f_score <- speciesdata$f[i] #get the score, CHANGE FOR B AND C
  
  for (b in 2:73){ #for each year
  index <- f_score*cat_totals[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[1, b]) #weight by production of ALL species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric
```

FOOD SECURITY WEIGHTED BY FINFISH
```{r}
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each FINFISH species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine") #filter out just the finfish

test2 <- filter(test, category =="finfish_inland"| category =="finfish_marine")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the score #CHANGE THIS FOR B AND C
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category
  index <- index/(weighting_totals[2, b]) #weight by production of finfish
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year
for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes1 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes1 <- pivot_longer(weighted_indexes1, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes1 <- weighted_indexes1[order(weighted_indexes1$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes1)) {
  print(weighted_indexes1$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes1$Year[i] <- year_labs[i]
}

weighted_indexes1$Year <- as.numeric(weighted_indexes1$Year) #make years numeric
```

FOOD SECURITY WEIGHTED BY ALGAE
```{r}
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each ALGAE species and multiply f score by (production of that species/production of all species)
#test (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "algae") #filter out just the algae

test2 <- filter(test, category =="algae")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$f[i] #get the score #CHANGE THIS FOR B AND C
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[4, b]) #weight by production of algae, row 4
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes2 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes2 <- pivot_longer(weighted_indexes2, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes2 <- weighted_indexes2[order(weighted_indexes2$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes2)) {
  print(weighted_indexes2$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes2$Year[i] <- year_labs[i]
}

weighted_indexes2$Year <- as.numeric(weighted_indexes2$Year) #make years numeric
```

FOOD SECURITY WEIGHTED BY CRUSTACEANS and MOLLUSCS
```{r}
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "crustacean")

test2 <- filter(test, category == "crustacean")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[5, b]) #weight by production of crustaceans, row 5
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes3 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes3 <- pivot_longer(weighted_indexes3, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes3 <- weighted_indexes3[order(weighted_indexes3$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes3)) {
  print(weighted_indexes3$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes3$Year[i] <- year_labs[i]
}

weighted_indexes3$Year <- as.numeric(weighted_indexes3$Year) #make years numeric

#MOLLUSCS
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "mollusc")

test2 <- filter(test, category == "mollusc")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[6, b]) #weight by production of molluscs, row 6
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes4 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes4 <- pivot_longer(weighted_indexes4, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes4 <- weighted_indexes4[order(weighted_indexes4$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes4)) {
  print(weighted_indexes4$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes4$Year[i] <- year_labs[i]
}

weighted_indexes4$Year <- as.numeric(weighted_indexes4$Year) #make years numeric

```

Graph of food security index with all taxa weightings
```{r}
#weighted_indexes = weighted by all
#need to add identifier column
weighted_indexes$weighted_by <- "All"

#weighted_indexes1 = weighted by finfish
weighted_indexes1$weighted_by <- "Finfish"

#weighted_indexes2 = weighted by algae
weighted_indexes2$weighted_by <- "Algae"

#weighted_indexes3 = weighted by crustaceans
weighted_indexes3$weighted_by <- "Crustaceans"

#weighted_indexes4 = weighted by molluscs
weighted_indexes4$weighted_by <- "Molluscs"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexes1)
merge <- rbind(merge, weighted_indexes2)
merge <- rbind(merge, weighted_indexes3)
merge <- rbind(merge, weighted_indexes4)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = weighted_by)) +geom_line(size =1)+
  labs(y = "Mean Food Security Index", title = "Food Security Potential of Aquaculture")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))+
    ylim(30,65)

#SAVE
# ggsave("F_weighted_combined.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set

```

```{r}
merge %>%
  group_by(weighted_by) %>%
  summarise(min = min(Score), max = max(Score))
```

```{r}
#magnitude of F, C and B indices by taxonomic group in three time periods
setwd("/Users/aleahw/Documents/Master's!!/Fuzzy/Chapter_2")

#modify dataset: create three time periods
scores_3_periods <- merge

scores_3_periods$Year[scores_3_periods$Year >= 1950 & scores_3_periods$Year <= 1980] <- "1950-1980"
scores_3_periods$Year[scores_3_periods$Year >= 1981 & scores_3_periods$Year <= 2000] <- "1981-2000"
scores_3_periods$Year[scores_3_periods$Year >= 2001 & scores_3_periods$Year <= 2021] <- "2001-2021"

plot <- ggplot(scores_3_periods, aes(x = Year, y=Score, color = weighted_by))+ geom_boxplot()+
  stat_summary(fun.y=mean,geom='point', position = position_dodge(width = .75))+
  stat_summary(fun=mean, geom="line", aes(group=weighted_by), position = position_dodge(width = .75))+
  scale_x_discrete(limits = c("1950-1980", "1981-2000", "2001-2021"))+
  scale_y_continuous(limits = c(30, 67))+
  labs(y = "Mean Food Security Index")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))


ggsave("mean_F_boxplots.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

########################################################################
CLIMATE CHANGE WEIGHTED BY DIFFERENT TAXA
```{r CLIMATE CHANGE WEIGHTED BY DIFFERENT TAXA}
# CLIMATE CHANGE WEIGHTED BY ALL PRODUCTION
#calculate index scores for each species weighted by production of ALL 54 species

new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata)){ #for each species
  print(speciesdata$species[i])
  f_score <- speciesdata$c[i] #get the score
  
  for (b in 2:73){ #for each year
  index <- f_score*cat_totals[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[1, b]) #weight by production of ALL species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric


#CLIMATE CHANGE WEIGHTED BY FINFISH
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each FINFISH species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine") #filter out just the finfish

test2 <- filter(test, category =="finfish_inland"| category =="finfish_marine")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$c[i] #get the score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category
  index <- index/(weighting_totals[2, b]) #weight by production of finfish
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year
for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes1 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes1 <- pivot_longer(weighted_indexes1, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes1 <- weighted_indexes1[order(weighted_indexes1$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes1)) {
  print(weighted_indexes1$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes1$Year[i] <- year_labs[i]
}

weighted_indexes1$Year <- as.numeric(weighted_indexes1$Year) #make years numeric

#CLIMATE CHANGE WEIGHTED BY ALGAE
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each ALGAE species and multiply f score by (production of that species/production of all species)
#test (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "algae") #filter out just the algae

test2 <- filter(test, category =="algae")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$c[i] #get the score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[4, b]) #weight by production of algae, row 4
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes2 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes2 <- pivot_longer(weighted_indexes2, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes2 <- weighted_indexes2[order(weighted_indexes2$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes2)) {
  print(weighted_indexes2$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes2$Year[i] <- year_labs[i]
}

weighted_indexes2$Year <- as.numeric(weighted_indexes2$Year) #make years numeric

new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "crustacean")

test2 <- filter(test, category == "crustacean")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$c[i] #get the c score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[5, b]) #weight by production of crustaceans, row 5
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes3 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes3 <- pivot_longer(weighted_indexes3, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes3 <- weighted_indexes3[order(weighted_indexes3$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes3)) {
  print(weighted_indexes3$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes3$Year[i] <- year_labs[i]
}

weighted_indexes3$Year <- as.numeric(weighted_indexes3$Year) #make years numeric

#MOLLUSCS
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "mollusc")

test2 <- filter(test, category == "mollusc")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$c[i] #get the c score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[6, b]) #weight by production of molluscs, row 6
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes4 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes4 <- pivot_longer(weighted_indexes4, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes4 <- weighted_indexes4[order(weighted_indexes4$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes4)) {
  print(weighted_indexes4$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes4$Year[i] <- year_labs[i]
}

weighted_indexes4$Year <- as.numeric(weighted_indexes4$Year) #make years numeric


#GRAPHING
#need to add identifier column
weighted_indexes$weighted_by <- "All"

#weighted_indexes1 = weighted by finfish
weighted_indexes1$weighted_by <- "Finfish"

#weighted_indexes2 = weighted by algae
weighted_indexes2$weighted_by <- "Algae"

#weighted_indexes3 = weighted by crustaceans
weighted_indexes3$weighted_by <- "Crustaceans"

#weighted_indexes4 = weighted by molluscs
weighted_indexes4$weighted_by <- "Molluscs"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexes1)
merge <- rbind(merge, weighted_indexes2)
merge <- rbind(merge, weighted_indexes3)
merge <- rbind(merge, weighted_indexes4)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = weighted_by)) +geom_line(size =1)+
  labs(y = "Mean Climate Change Index", title = "Climate Change Potential of Aquaculture")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))+
  ylim(30,65)

#SAVE
# ggsave("C_weighted_combined.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set

```

```{r}
merge %>%
  group_by(weighted_by) %>%
  summarise(min = min(Score), max = max(Score))
```

```{r}
#magnitude of F, C and B indices by taxonomic group in three time periods
setwd("/Users/aleahw/Documents/Master's!!/Fuzzy/Chapter_2")

#modify dataset: create three time periods
scores_3_periods <- merge

scores_3_periods$Year[scores_3_periods$Year >= 1950 & scores_3_periods$Year <= 1980] <- "1950-1980"
scores_3_periods$Year[scores_3_periods$Year >= 1981 & scores_3_periods$Year <= 2000] <- "1981-2000"
scores_3_periods$Year[scores_3_periods$Year >= 2001 & scores_3_periods$Year <= 2021] <- "2001-2021"

plot <- ggplot(scores_3_periods, aes(x = Year, y=Score, color = weighted_by))+ geom_boxplot()+
  stat_summary(fun.y=mean,geom='point', position = position_dodge(width = .75))+
  stat_summary(fun=mean, geom="line", aes(group=weighted_by), position = position_dodge(width = .75))+
  scale_x_discrete(limits = c("1950-1980", "1981-2000", "2001-2021"))+
  scale_y_continuous(limits = c(30, 67))+
  labs(y = "Mean Climate Index")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))


ggsave("mean_C_boxplots.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```


#####################################################

BIODIV WEIGHTED BY DIFFERENT TAXA
```{r BIODIV WEIGHTED BY DIFFERENT TAXA}
#calculate index scores for each species weighted by production of ALL 54 species

new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each species and multiply b score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata)){ #for each species
  print(speciesdata$species[i])
  f_score <- speciesdata$b[i] #get the score
  
  for (b in 2:73){ #for each year
  index <- f_score*cat_totals[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[1, b]) #weight by production of ALL species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric


#WEIGHTED BY FINFISH
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each FINFISH species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine") #filter out just the finfish

test2 <- filter(test, category =="finfish_inland"| category =="finfish_marine")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$b[i] #get the score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category
  index <- index/(weighting_totals[2, b]) #weight by production of finfish
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year
for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes1 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes1 <- pivot_longer(weighted_indexes1, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes1 <- weighted_indexes1[order(weighted_indexes1$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes1)) {
  print(weighted_indexes1$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes1$Year[i] <- year_labs[i]
}

weighted_indexes1$Year <- as.numeric(weighted_indexes1$Year) #make years numeric

#WEIGHTED BY ALGAE
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each ALGAE species and multiply f score by (production of that species/production of all species)
#test (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "algae") #filter out just the algae

test2 <- filter(test, category =="algae")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$b[i] #get the score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[4, b]) #weight by production of algae, row 4
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes2 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes2 <- pivot_longer(weighted_indexes2, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes2 <- weighted_indexes2[order(weighted_indexes2$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes2)) {
  print(weighted_indexes2$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes2$Year[i] <- year_labs[i]
}

weighted_indexes2$Year <- as.numeric(weighted_indexes2$Year) #make years numeric

new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "crustacean")

test2 <- filter(test, category == "crustacean")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$b[i] #get the b score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[5, b]) #weight by production of crustaceans, row 5
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes3 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes3 <- pivot_longer(weighted_indexes3, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes3 <- weighted_indexes3[order(weighted_indexes3$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes3)) {
  print(weighted_indexes3$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes3$Year[i] <- year_labs[i]
}

weighted_indexes3$Year <- as.numeric(weighted_indexes3$Year) #make years numeric

#MOLLUSCS
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "mollusc")

test2 <- filter(test, category == "mollusc")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$b[i] #get the b score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[6, b]) #weight by production of molluscs, row 6
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes4 <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes4 <- pivot_longer(weighted_indexes4, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes4 <- weighted_indexes4[order(weighted_indexes4$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes4)) {
  print(weighted_indexes4$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes4$Year[i] <- year_labs[i]
}

weighted_indexes4$Year <- as.numeric(weighted_indexes4$Year) #make years numeric


#GRAPHING
#need to add identifier column
weighted_indexes$weighted_by <- "All"

#weighted_indexes1 = weighted by finfish
weighted_indexes1$weighted_by <- "Finfish"

#weighted_indexes2 = weighted by algae
weighted_indexes2$weighted_by <- "Algae"

#weighted_indexes3 = weighted by crustaceans
weighted_indexes3$weighted_by <- "Crustaceans"

#weighted_indexes4 = weighted by molluscs
weighted_indexes4$weighted_by <- "Molluscs"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexes1)
merge <- rbind(merge, weighted_indexes2)
merge <- rbind(merge, weighted_indexes3)
merge <- rbind(merge, weighted_indexes4)

#graph
ggplot(merge, aes(x=Year, y=Score, color = weighted_by)) +geom_line(size =1)+
  labs(y = "Mean Biodiversity Index", title = "Biodiversity Potential of Aquaculture")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))

#range 20-70?
#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = weighted_by)) +geom_line(size =1)+
  labs(y = "Mean Biodiversity Index", title = "Biodiversity Potential of Aquaculture")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))+
  ylim(30,70)

#SAVE
# ggsave("B_weighted_combined.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set

```

```{r}
#find minimum and maximum
merge %>%
  group_by(weighted_by) %>%
  summarise(min = min(Score), max = max(Score))

```

```{r}
#magnitude of F, C and B indices by taxonomic group in three time periods
setwd("/Users/aleahw/Documents/Master's!!/Fuzzy/Chapter_2")

#modify dataset: create three time periods
scores_3_periods <- merge

scores_3_periods$Year[scores_3_periods$Year >= 1950 & scores_3_periods$Year <= 1980] <- "1950-1980"
scores_3_periods$Year[scores_3_periods$Year >= 1981 & scores_3_periods$Year <= 2000] <- "1981-2000"
scores_3_periods$Year[scores_3_periods$Year >= 2001 & scores_3_periods$Year <= 2021] <- "2001-2021"

plot <- ggplot(scores_3_periods, aes(x = Year, y=Score, color = weighted_by))+ geom_boxplot()+
  stat_summary(fun.y=mean,geom='point', position = position_dodge(width = .75))+
  stat_summary(fun=mean, geom="line", aes(group=weighted_by), position = position_dodge(width = .75))+
  scale_x_discrete(limits = c("1950-1980", "1981-2000", "2001-2021"))+
  scale_y_continuous(limits = c(30, 67))+
  labs(y = "Mean Biodiversity Index")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Weighted By", values=c('#7D627F', '#7DBBC3', "pink", "orange", "lightgreen"))


ggsave("mean_B_boxplots.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

################################################################
--------------------------------------------------------------------------------------------------
#Combining F, C, B for each weighting: run this before graphing
```{r COMBINED FCB GRAPH WEIGHTED BY PRODUCTION OF ALL SPECIES}

#FOOD SECURITY
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each species and multiply f score by (production of that species/production of all species)

for (i in 1:nrow(speciesdata)){ #for each species
  print(speciesdata$species[i])
  f_score <- speciesdata$f[i] #get the FOOD SCORE
  
  for (b in 2:73){ #for each year
  index <- f_score*cat_totals[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[1, b]) #weight by production of ALL species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#CLIMATE CHANGE

new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata)){ #for each species
  print(speciesdata$species[i])
  f_score <- speciesdata$c[i] #get the CLIMATE SCORE
  
  for (b in 2:73){ #for each year
  index <- f_score*cat_totals[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[1, b]) #weight by production of ALL species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesc <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesc <- pivot_longer(weighted_indexesc, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesc <- weighted_indexesc[order(weighted_indexesc$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesc)) {
  print(weighted_indexesc$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesc$Year[i] <- year_labs[i]
}

weighted_indexesc$Year <- as.numeric(weighted_indexesc$Year) #make years numeric

#BIODIV

new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata)){ #for each species
  print(speciesdata$species[i])
  f_score <- speciesdata$b[i] #get the BIODIV SCORE
  
  for (b in 2:73){ #for each year
  index <- f_score*cat_totals[i, b] #multiply score by that species' total production in each year
  index <- index/(weighting_totals[1, b]) #weight by production of ALL species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesb <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesb <- pivot_longer(weighted_indexesb, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesb <- weighted_indexesb[order(weighted_indexesb$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesb)) {
  print(weighted_indexesb$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesb$Year[i] <- year_labs[i]
}

weighted_indexesb$Year <- as.numeric(weighted_indexesb$Year) #make years numeric

#GRAPHING
#weighted_indexes = F weighted by all
weighted_indexes$Index <- "Food Security"
#weighted_indexesc = C weighted by all 
weighted_indexesc$Index <- "Climate Change"
#weighted_indexesb = B weighted by all 
weighted_indexesb$Index <- "Biodiversity"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexesc)
merge <- rbind(merge, weighted_indexesb)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = Index)) +geom_line(size =1)+
  labs(y = "Mean Index", title = "FCB Potential of Aquaculture Weighted by 54 Major Species")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))+
  ylim(48, 70)


#SAVE
# setwd("/Users/aleahw/Documents/Master's!!/Fuzzy/production_totals")
# ggsave("FCB_weightedby_all2.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```


```{r correlations of FCB over time}

#reconfigure f
weighted_indexes_graph <- weighted_indexes[, -3]
colnames(weighted_indexes_graph)[2] = "Food_Security"

#reconfigure c
weighted_indexes_graphc <- weighted_indexesc[, -3]
colnames(weighted_indexes_graphc)[2] = "Climate_Change"

#reconfigure b
weighted_indexes_graphb <- weighted_indexesb[, -3]
colnames(weighted_indexes_graphb)[2] = "Biodiversity"

#merge by year
a <- merge(weighted_indexes_graph, weighted_indexes_graphc, by="Year")
a <- merge(a, weighted_indexes_graphb, by="Year")

#find correlations

# Do the data from each of the 2 variables (x, y) follow a normal distribution?
# Use Shapiro-Wilk normality test > R function: shapiro.test()
# and look at the normality plot > R function: ggpubr::ggqqplot()

# Shapiro-Wilk test can be performed as follow:
# Null hypothesis: the data are normally distributed
# Alternative hypothesis: the data are not normally distributed

# Shapiro-Wilk normality test for f
shapiro.test(a$Food_Security) # => p NOT NORMAL p-value = 0.0005529
# Shapiro-Wilk normality test for c
shapiro.test(a$Climate_Change) # => p NOT NORMAL p-value = 0.006915
# Shapiro-Wilk normality test for b
shapiro.test(a$Biodiversity) # => NORMAL: p = 0.01736

# Visual inspection of the data normality using Q-Q plots (quantile-quantile plots). Q-Q plot draws the correlation between a given sample and the normal distribution.
library(ggpubr)
ggqqplot(a$Food_Security, ylab = "Food Security")
ggqqplot(a$Climate_Change, ylab = "Climate Change")
ggqqplot(a$Biodiversity, ylab = "Biodiversity")

#histogram to see dist
ggdensity(a$Food_Security, 
          main = "Density plot of F index",
          xlab = "Food Security Index")

ggdensity(a$Climate_Change, 
          main = "Density plot of C index",
          xlab = "Climate Change Index")

ggdensity(a$Biodiversity, 
          main = "Density plot of B index",
          xlab = "Biodiversity Index")


#PEARSON, assuming normal distribution...
#FOOD SECURITY AND CLIMATE CHANGE
ggscatter(a, x = "Food_Security", y = "Climate_Change", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Food Security Index", ylab = "Climate Change Index") #SIGNIFICANT, R = 0.92; p = < 2.2 e-16

res <- cor.test(a$Food_Security, a$Climate_Change, 
                    method = "pearson")
res

#FOOD SECURITY AND BIODIVERSITY
ggscatter(a, x = "Food_Security", y = "Biodiversity", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Food Security Index", ylab = "Biodiversity Index") #SIGNIFICANT, R = 0.95; p = < 2.2 e-16

res <- cor.test(a$Food_Security, a$Biodiversity, 
                    method = "pearson")
res

#BIODIVERSITY AND CLIMATE CHANGE
ggscatter(a, x = "Climate_Change", y = "Biodiversity", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Climate Change Index", ylab = "Biodiversity Index") #SIGNIFICANT, R = 0.79; p = 3.086e-16

res <- cor.test(a$Biodiversity, a$Climate_Change, 
                    method = "pearson")
res

#Spearman rank correlation, assuming non-normal dist
# The Kendall rank correlation coefficient or Kendalls tau statistic is used to estimate a rank-based measure of association. This test may be used if the data do not necessarily come from a bivariate normal distribution.

res2 <- cor.test(a$Food_Security, a$Climate_Change,  method="kendall") #SIG; tau = 0.6697966, p-value < 2.2e-16
res2

res2 <- cor.test(a$Food_Security, a$Biodiversity,  method="kendall") #SIG; tau = 0.5547731; p-value = 5.453e-12
res2

res2 <- cor.test(a$Biodiversity, a$Climate_Change,  method="kendall") #SIG; tau = 0.7519562; p-value < 2.2e-16
res2
```


CALCULATING DIVERGENCE or CONVERGENCE
```{r}

#make five time slices, 13 to 14 years each

F1 <- filter(merge, Year < 1965, Index == "Food Security")
F2 <- filter(merge, Year > 1964 & Year < 1979, Index == "Food Security")
F3 <- filter(merge, Year > 1978 & Year < 1993, Index == "Food Security")
F4 <- filter(merge, Year > 1992 & Year < 2007, Index == "Food Security")
F5 <- filter(merge, Year > 2006, Index == "Food Security")

C1 <- filter(merge, Year < 1965, Index == "Climate Change")
C2 <- filter(merge, Year > 1964 & Year < 1979, Index == "Climate Change")
C3 <- filter(merge, Year > 1978 & Year < 1993, Index == "Climate Change")
C4 <- filter(merge, Year > 1992 & Year < 2007, Index == "Climate Change")
C5 <- filter(merge, Year > 2006, Index == "Climate Change")

B1 <- filter(merge, Year < 1965, Index == "Biodiversity")
B2 <- filter(merge, Year > 1964 & Year < 1979, Index == "Biodiversity")
B3 <- filter(merge, Year > 1978 & Year < 1993, Index == "Biodiversity")
B4 <- filter(merge, Year > 1992 & Year < 2007, Index == "Biodiversity")
B5 <- filter(merge, Year > 2006, Index == "Biodiversity")

#make new frame to store values
divergence = data.frame(matrix(
  vector(), 5, 3, dimnames=list(c(), c("B-F","B-C","F-C"))),
                stringsAsFactors=F)
rownames(divergence) <- c("1950-1964","1965-1978","1979-1992","1993-2006", "2007-2021")

#subtract and add values to dataframe
# divergence$B.F[1] <- B1-F1

BF1 <- data.frame(B1$Score - F1$Score)
BF2 <- data.frame(B2$Score - F2$Score)
BF3 <- data.frame(B3$Score - F3$Score)
BF4 <- data.frame(B4$Score - F4$Score)
BF5 <- data.frame(B5$Score - F5$Score)

BC1 <- data.frame(B1$Score - C1$Score)
BC2 <- data.frame(B2$Score - C2$Score)
BC3 <- data.frame(B3$Score - C3$Score)
BC4 <- data.frame(B4$Score - C4$Score)
BC5 <- data.frame(B5$Score - C5$Score)

FC1 <- data.frame(F1$Score - C1$Score)
FC2 <- data.frame(F2$Score - C2$Score)
FC3 <- data.frame(F3$Score - C3$Score)
FC4 <- data.frame(F4$Score - C4$Score)
FC5 <- data.frame(F5$Score - C5$Score)

slice1 <- cbind(BF1, BC1, FC1) #1950-1964
divergence$B.F[1] <- mean(slice1$B1.Score...F1.Score)
divergence$B.C[1] <- mean(slice1$B1.Score...C1.Score)
divergence$F.C[1] <- mean(slice1$F1.Score...C1.Score)

slice2 <- cbind(BF2, BC2, FC2)
divergence$B.F[2] <- mean(slice2$B2.Score...F2.Score)
divergence$B.C[2] <-mean(slice2$B2.Score...C2.Score)
divergence$F.C[2] <- mean(slice2$F2.Score...C2.Score)

slice3 <- cbind(BF3, BC3, FC3)
divergence$B.F[3] <- mean(slice3$B3.Score...F3.Score)
divergence$B.C[3] <- mean(slice3$B3.Score...C3.Score)
divergence$F.C[3] <- mean(slice3$F3.Score...C3.Score)

slice4 <- cbind(BF4, BC4, FC4)
divergence$B.F[4] <- mean(slice4$B4.Score...F4.Score)
divergence$B.C[4] <- mean(slice4$B4.Score...C4.Score)
divergence$F.C[4] <- mean(slice4$F4.Score...C4.Score)

slice5 <- cbind(BF5, BC5, FC5)
divergence$B.F[5] <- mean(slice5$B5.Score...F5.Score)
divergence$B.C[5] <- mean(slice5$B5.Score...C5.Score)
divergence$F.C[5] <- mean(slice5$F5.Score...C5.Score)

#SAVE

NAME<- sprintf("/Users/aleahw/Documents/Master's!!/Fuzzy/Figures/divergence2.csv")
write.csv(divergence,file=NAME)
```

```{r COMBINED FCB GRAPH WEIGHTED BY FINFISH}

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each FINFISH species and multiply f score by (production of that species/production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine") #filter out just the finfish

test2 <- filter(test, category =="finfish_inland"| category =="finfish_marine")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category
  index <- index/(weighting_totals[2, b]) #weight by production of finfish
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year
for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric


#CLIMATE CHANGE
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine") #filter out just the finfish

test2 <- filter(test, category =="finfish_inland"| category =="finfish_marine")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$c[i] #get the c score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category
  index <- index/(weighting_totals[2, b]) #weight by production of finfish
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year
for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesc <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesc <- pivot_longer(weighted_indexesc, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesc <- weighted_indexesc[order(weighted_indexesc$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesc)) {
  print(weighted_indexesc$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesc$Year[i] <- year_labs[i]
}

weighted_indexesc$Year <- as.numeric(weighted_indexesc$Year) #make years numeric


#BIODIVERSITY
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine") #filter out just the finfish

test2 <- filter(test, category =="finfish_inland"| category =="finfish_marine")

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$b[i] #get the b score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category
  index <- index/(weighting_totals[2, b]) #weight by production of finfish
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year
for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesb <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesb <- pivot_longer(weighted_indexesb, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesb <- weighted_indexesb[order(weighted_indexesb$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesb)) {
  print(weighted_indexesb$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesb$Year[i] <- year_labs[i]
}

weighted_indexesb$Year <- as.numeric(weighted_indexesb$Year) #make years numeric

#GRAPHING
#weighted_indexes = F weighted by all
weighted_indexes$Index <- "Food Security"
#weighted_indexesc = C weighted by all 
weighted_indexesc$Index <- "Climate Change"
#weighted_indexesb = B weighted by all 
weighted_indexesb$Index <- "Biodiversity"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexesc)
merge <- rbind(merge, weighted_indexesb)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = Index)) +geom_line(size =1)+
  labs(y = "Mean Index", title = "FCB Potential of Aquaculture Weighted by Finfish")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))+
        ylim(45, 54)


#SAVE
# ggsave("FCB_weightedby_finfish.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

```{r COMBINED FCB GRAPH WEIGHTED BY ALGAE}

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "algae") #filter out just the algae

test2 <- filter(test, category =="algae")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[4, b]) #weight by production of algae, row 4
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#CLIMATE CHANGE
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "algae") #filter out just the algae

test2 <- filter(test, category =="algae")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$c[i] #get the c score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[4, b]) #weight by production of algae, row 4
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesc <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesc <- pivot_longer(weighted_indexesc, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesc <- weighted_indexesc[order(weighted_indexesc$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesc)) {
  print(weighted_indexesc$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesc$Year[i] <- year_labs[i]
}

weighted_indexesc$Year <- as.numeric(weighted_indexesc$Year) #make years numeric

#BIODIVERSITY
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "algae") #filter out just the algae

test2 <- filter(test, category =="algae")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$b[i] #get the b score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[4, b]) #weight by production of algae, row 4
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesb <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesb <- pivot_longer(weighted_indexesb, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesb <- weighted_indexesb[order(weighted_indexesb$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesb)) {
  print(weighted_indexesb$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesb$Year[i] <- year_labs[i]
}

weighted_indexesb$Year <- as.numeric(weighted_indexesb$Year) #make years numeric

#GRAPHING
#weighted_indexes = F weighted by all
weighted_indexes$Index <- "Food Security"
#weighted_indexesc = C weighted by all 
weighted_indexesc$Index <- "Climate Change"
#weighted_indexesb = B weighted by all 
weighted_indexesb$Index <- "Biodiversity"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexesc)
merge <- rbind(merge, weighted_indexesb)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = Index)) +geom_line(size =1)+
  labs(y = "Mean Index", title = "FCB Potential of Aquaculture Weighted by Algae")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))+
  ylim(53, 67)


#SAVE
# ggsave("FCB_weightedby_algae.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set

```

```{r COMBINED FCB GRAPH WEIGHTED BY CRUSTACEANS}

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "crustacean")

test2 <- filter(test, category == "crustacean")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[5, b]) #weight by production of crustaceans
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#CLIMATE CHANGE
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "crustacean")

test2 <- filter(test, category == "crustacean")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$c[i] #get the c score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[5, b]) #weight by production of crustaceans
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesc <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesc <- pivot_longer(weighted_indexesc, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesc <- weighted_indexesc[order(weighted_indexesc$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesc)) {
  print(weighted_indexesc$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesc$Year[i] <- year_labs[i]
}

weighted_indexesc$Year <- as.numeric(weighted_indexesc$Year) #make years numeric

#BIODIVERSITY
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "crustacean")

test2 <- filter(test, category == "crustacean")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$b[i] #get the b score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[5, b]) #weight by crustaceans
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesb <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesb <- pivot_longer(weighted_indexesb, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesb <- weighted_indexesb[order(weighted_indexesb$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesb)) {
  print(weighted_indexesb$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesb$Year[i] <- year_labs[i]
}

weighted_indexesb$Year <- as.numeric(weighted_indexesb$Year) #make years numeric

#GRAPHING
#weighted_indexes = F weighted by all
weighted_indexes$Index <- "Food Security"
#weighted_indexesc = C weighted by all 
weighted_indexesc$Index <- "Climate Change"
#weighted_indexesb = B weighted by all 
weighted_indexesb$Index <- "Biodiversity"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexesc)
merge <- rbind(merge, weighted_indexesb)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = Index)) +geom_line(size =1)+
  labs(y = "Mean Index", title = "FCB Potential of Aquaculture Weighted by Crustaceans")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))+
  ylim(30, 55)


#SAVE
# ggsave("FCB_weightedby_crustaceans.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

```{r COMBINED FCB GRAPH WEIGHTED BY MOLLUSCS}

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "mollusc")

test2 <- filter(test, category == "mollusc")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[6, b]) #weight by production of molluscs
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexes <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexes <- pivot_longer(weighted_indexes, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  print(weighted_indexes$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#CLIMATE CHANGE
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "mollusc")

test2 <- filter(test, category == "mollusc")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$c[i] #get the c score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[6, b]) #weight by production of molluscs
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesc <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesc <- pivot_longer(weighted_indexesc, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesc <- weighted_indexesc[order(weighted_indexesc$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesc)) {
  print(weighted_indexesc$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesc$Year[i] <- year_labs[i]
}

weighted_indexesc$Year <- as.numeric(weighted_indexesc$Year) #make years numeric

#BIODIVERSITY
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

speciesdata2 <- filter(speciesdata, category == "mollusc")

test2 <- filter(test, category == "mollusc")

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$b[i] #get the b score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[6, b]) #weight by molluscs
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

#calculate the annual mean each year

for (i in 2:73){
    a <- sum(new_frame[, i], na.rm = TRUE) #sum all the columns
    print(a)
    
    new_frame[58, i] <- a
}

new_frame[58, 1] <- "Weighted Index" #name the newest row
weighted_indexesb <- new_frame[58, 2:73] #new frame with just the weighted scores each year

library(tidyr)
#make long
weighted_indexesb <- pivot_longer(weighted_indexesb, cols = everything(), 
                                names_to = "Year", values_to = "Score")

weighted_indexesb <- weighted_indexesb[order(weighted_indexesb$Year),] #put earlier years first

year_labs <- seq(1950, 2021, by = 1)

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexesb)) {
  print(weighted_indexesb$Year[i])
  #replace each year with the actual year number from year_labs
  
  weighted_indexesb$Year[i] <- year_labs[i]
}

weighted_indexesb$Year <- as.numeric(weighted_indexesb$Year) #make years numeric

#GRAPHING
#weighted_indexes = F weighted by all
weighted_indexes$Index <- "Food Security"
#weighted_indexesc = C weighted by all 
weighted_indexesc$Index <- "Climate Change"
#weighted_indexesb = B weighted by all 
weighted_indexesb$Index <- "Biodiversity"

#put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexesc)
merge <- rbind(merge, weighted_indexesb)

#graph
plot <- ggplot(merge, aes(x=Year, y=Score, color = Index)) +geom_line(size =1)+
  labs(y = "Mean Index", title = "FCB Potential of Aquaculture Weighted by Molluscs")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))+
  ylim(44, 66)



#SAVE
# ggsave("FCB_weightedby_molluscs.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```


CHECKING FOR CORRELATION
```{r}
library("ggpubr")
library(ggplot2)

# Do the data from each of the 2 variables (x, y) follow a normal distribution?
# Use Shapiro-Wilk normality test > R function: shapiro.test()
# and look at the normality plot > R function: ggpubr::ggqqplot()

# Shapiro-Wilk test can be performed as follow:
# Null hypothesis: the data are normally distributed
# Alternative hypothesis: the data are not normally distributed

# Shapiro-Wilk normality test for f
shapiro.test(speciesdata$f) # => p-value = 0.00916 NOT NORMAL
# Shapiro-Wilk normality test for c
shapiro.test(speciesdata$c) # => p-value = 0.109 NORMAL
# Shapiro-Wilk normality test for b
shapiro.test(speciesdata$b) # => p-value = 0.01076 NORMAL 

# Visual inspection of the data normality using Q-Q plots (quantile-quantile plots). Q-Q plot draws the correlation between a given sample and the normal distribution.

ggqqplot(speciesdata$f, ylab = "Food Security")
ggqqplot(speciesdata$c, ylab = "Climate Change")
ggqqplot(speciesdata$b, ylab = "Biodiversity")

#histogram to see dist
ggdensity(speciesdata$f, 
          main = "Density plot of F index",
          xlab = "Food Security Index")

ggdensity(speciesdata$c, 
          main = "Density plot of C index",
          xlab = "Climate Change Index")

ggdensity(speciesdata$b, 
          main = "Density plot of B index",
          xlab = "Biodiversity Index")


#PEARSON, assuming normal distribution...
#FOOD SECURITY AND CLIMATE CHANGE
ggscatter(speciesdata, x = "f", y = "c", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Food Security Index", ylab = "Climate Change Index") #SIGNIFICANT, R = 0.48; p = 0.0001504

res <- cor.test(speciesdata$f, speciesdata$c, 
                    method = "pearson")
res

#FOOD SECURITY AND BIODIVERSITY
ggscatter(speciesdata, x = "f", y = "b", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Food Security Index", ylab = "Biodiversity Index") #SIGNIFICANT, R = 0.44; p = 0.0005818

res <- cor.test(speciesdata$f, speciesdata$b, 
                    method = "pearson")
res

#BIODIVERSITY AND CLIMATE CHANGE
ggscatter(speciesdata, x = "c", y = "b", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Climate Change Index", ylab = "Biodiversity Index") #SIGNIFICANT, R = 0.48; p = 0.000145

res <- cor.test(speciesdata$c, speciesdata$b, 
                    method = "pearson")
res

#Spearman rank correlation, assuming non-normal dist
# The Kendall rank correlation coefficient or Kendalls tau statistic is used to estimate a rank-based measure of association. This test may be used if the data do not necessarily come from a bivariate normal distribution.

res2 <- cor.test(speciesdata$f, speciesdata$c,  method="kendall") #SIG; tau = 0.4261061, p-value = 2.944e-06
res2

res2 <- cor.test(speciesdata$f, speciesdata$b,  method="kendall") #SIG; tau = 0.3795163; p-value = 3.199e-05; z = 4.9297
res2

res2 <- cor.test(speciesdata$c, speciesdata$b,  method="kendall") #SIG; tau = 0.359071; p-value = 8.212e-05
res2

```

#################################################
Percent contribution to index
#make sure to run from line 565 (for food security) so you can use "weighted_indexes_1 etc" or from 878 for climate change, or 1183 for biodiv

#new approach: % contribution of taxonomic groups relative to F, C, B relative to the proportion of total production
#Make three graphs, one for each FCB. For each taxonomic group, divide mean F/total production etc, plot all taxonomic groups on the same graph. Y-axis will be 0 to 1; don't plot production line, just the ratio.

```{r}
#first, calculate the % contribution of each taxa to annual production. We will eventually divide FCB indices by this proportion to find a ratio.

weighting_totals4 <- weighting_totals[-c(3), ]

new_frame <- weighting_totals4 #create new frame
new_frame <- new_frame[FALSE, ]

#divide taxa production in each year by the total prod in that year

for (i in 2:nrow(weighting_totals4)){ #for each taxonomic group
    for (b in 2:73){ #for each year
      total_prod <- weighting_totals4[1, b] #get the total prod in that year
      index <- weighting_totals4[i, b]/total_prod*100 #divide score by that species' total prod
      new_frame[i, b] <- index   #add these numbers to the frame
      
  } 
}

#delete first row
new_frame <- new_frame[-c(1), ]
#rename
new_frame[1, 1] <-"Finfish"
new_frame[2, 1] <-"Algae"
new_frame[3, 1] <-"Crustacean"
new_frame[4, 1] <-"Mollusc"

#make long?
new_frame <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Perc_prod")



#fix years
year_labs <- rep(2021:1950, times=6) 


#FIX from here
#rename year column in weighted_indexes
for (i in 1:nrow(new_frame)) {
  print(new_frame$Year[i])
  #replace each year with the actual year number from year_labs
  
  new_frame$Year[i] <- year_labs[i]
}


#save this
totalproduction <- new_frame
#goes one taxa at a time, from 2021 to 1950

#NEED TO PUT IN year order??
totalproduction <- totalproduction %>% arrange(Year)
```

```{r}
#now, calculate the percentage contribution of each taxa to F, C and B in each year.

#FOOD00 from line 570
#CLIMATE--run from 920
#CLIMATE--run from 1258

#weighting totals has an extra row: get rid of it here
weighting_totals3 <- weighting_totals[-3, ]
#also put in reverse order
weighting_totals3 <- rev(weighting_totals3)

#merge data
all_taxa <- merge(weighted_indexes1, weighted_indexes2, by="Year")
all_taxa <- merge(all_taxa, weighted_indexes3, by="Year")
all_taxa <- merge(all_taxa, weighted_indexes4, by="Year")
all_taxa <- merge(all_taxa, weighted_indexes, by="Year")

all_taxa <- all_taxa[ -c(3, 5, 7, 9, 11) ]

#rename
colnames(all_taxa)[2] <- "Finfish"
colnames(all_taxa)[3] <- "Algae"
colnames(all_taxa)[4] <- "Crustaceans"
colnames(all_taxa)[5] <- "Molluscs"
colnames(all_taxa)[6] <- "All"

new_frame <- all_taxa[, -c(6)] #create new frame
# new_frame <- new_frame[FALSE, ]

#for each taxa
for (i in 2:5) {
  # print(weighting_totals3$species[i])
 for (y in 1:72){ #for each year
   taxa_perc <- weighting_totals3[i, y]/weighting_totals3[1,y]
   # print(weighting_totals3[i, y])
   # print(taxa_perc)
   product <- taxa_perc*all_taxa[y, i] #multiply the index of the taxa by taxa_perc
   # print(all_taxa[y, i])
   # print(product)
   product <- product/all_taxa[y, 6] #fix!!!!!!!?
   print(product) #put this number in a dataframe?
   
   new_frame[y, i] <- product*100
   
 }
}

#make long?
new_frame <- pivot_longer(new_frame, cols = -c("Year"),
                                names_to = "species", values_to = "index_perc")

#save this
totalindex <- new_frame
```

```{r}
#combine totalindex and totalprod
prod_index <- cbind(totalindex, totalproduction)
prod_index <- prod_index[, -c(4, 5)]

#new column divides perc contribution to index, by perc contribution to production
prod_index[, 5] <- "new"
colnames(prod_index)[5] <- "ratio"

#calculate ratio for every row
for (i in 1:nrow(prod_index)) {
  ratio <- prod_index$index_perc[i]/prod_index$Perc_prod[i]
  prod_index$ratio[i] <- ratio
}

#graph this:
plot <- ggplot(prod_index, aes(x=Year, y=as.numeric(ratio) , color=species)) + geom_point()+
      geom_hline(yintercept = 1)+
      labs(y = "Contribution to B index:contribution to production")
      # ylim(0, 1.5)

#SAVE
# setwd("/Users/aleahw/Documents/Master's!!/Chapter_2")
# ggsave("Ratio_B_prod.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set

```


```{r}
#graph
#make long
new_frame <- pivot_longer(new_frame, cols = -c("Year"),
                                names_to = "Taxa", values_to = "Score")

# weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

#specify order of bars (from top to bottom)
new_frame$Taxa <- factor(new_frame$Taxa, levels=c('Algae','Crustaceans', 'Molluscs','Finfish'))

#create stacked bar chart
ggplot(new_frame, aes(x=Year, y=Score, fill=Taxa)) + 
    geom_bar(position='stack', stat='identity')+
     labs(y = "Percent contribution to F index")

# weighting_totals4 <- pivot_wider(merge, names_from = "Year", values_from = "Score")
# 
# new_frame <- weighting_totals4 #create new frame
# new_frame <- new_frame[FALSE, ]
```


#OLD
```{r OLD}

#weighting totals has an extra row: get rid of it here
weighting_totals3 <- weighting_totals[-3, ]
#also put in reverse order
weighting_totals3 <- rev(weighting_totals3)

#merge data
all_taxa <- merge(weighted_indexes1, weighted_indexes2, by="Year")
all_taxa <- merge(all_taxa, weighted_indexes3, by="Year")
all_taxa <- merge(all_taxa, weighted_indexes4, by="Year")
all_taxa <- merge(all_taxa, weighted_indexes, by="Year")

all_taxa <- all_taxa[ -c(3, 5, 7, 9, 11) ]

#rename
colnames(all_taxa)[2] <- "Finfish"
colnames(all_taxa)[3] <- "Algae"
colnames(all_taxa)[4] <- "Crustaceans"
colnames(all_taxa)[5] <- "Molluscs"
colnames(all_taxa)[6] <- "All"

new_frame <- all_taxa[, -c(6)] #create new frame
# new_frame <- new_frame[FALSE, ]

#for each taxa
for (i in 2:5) {
  # print(weighting_totals3$species[i])
 for (y in 1:72){ #for each year
   taxa_perc <- weighting_totals3[i, y]/weighting_totals3[1,y]
   # print(weighting_totals3[i, y])
   # print(taxa_perc)
   product <- taxa_perc*all_taxa[y, i] #multiply the index of the taxa by taxa_perc
   # print(all_taxa[y, i])
   # print(product)
   product <- product/all_taxa[y, 6] #fix!!!!!!!?
   print(product) #put this number in a dataframe?
   
   new_frame[y, i] <- product
 }
}

#graph
#make long
new_frame <- pivot_longer(new_frame, cols = -c("Year"),
                                names_to = "Taxa", values_to = "Score")

# weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

#specify order of bars (from top to bottom)
new_frame$Taxa <- factor(new_frame$Taxa, levels=c('Algae','Crustaceans', 'Molluscs','Finfish'))

#create stacked bar chart
ggplot(new_frame, aes(x=Year, y=Score, fill=Taxa)) + 
    geom_bar(position='stack', stat='identity')+
     labs(y = "Percent contribution to F index")


# 
# ggplot(new_frame, aes(fill=Taxa, y=Score, x=Year)) + 
#     geom_bar(position="fill", stat="identity")+
#     labs(y = "Percent contribution to B index")+
#       scale_y_continuous(labels = scales::percent) +
#         geom_text(aes(label = paste0(Score*100,"%")),
#                   position = position_stack(vjust = 0.5), size = 0.5)
              
#change colors?

#SAVE
# setwd("/Users/aleahw/Documents/Master's!!/Fuzzy/production_totals")
# ggsave("percent_contribution_B2.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

```{r}
#compared to a stacked % barchart of just production

#delete "all" from weighting totals 2
weighting_totals2  <- weighting_totals2 %>%
  filter(species != "all species")

#specify order of bars (from top to bottom)
weighting_totals2$species <- factor(weighting_totals2$species, levels=c('algae','crustacean', 'mollusc','finfish'))

ggplot(weighting_totals2, aes(fill=species, y=Score, x=Year)) + 
    geom_bar(position="fill", stat="identity")+
    labs(y = "Percent contribution to total production")

# ggsave("percent_contribution_prod2.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set
```

#########################################
#Minimum and Maximum FCB index of each taxa weighted by total production of taxa
```{r}
#find minimum and maximum FCB index of each species

#finfish:
speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine")

test2 <- filter(test, category =="finfish_inland"| category =="finfish_marine")

fmax <- subset(speciesdata2, f == max(f))
fmin <- subset(speciesdata2, f == min(f))

cmax <- subset(speciesdata2, c == max(c))
cmin <- subset(speciesdata2, c == min(c))

bmax <- subset(speciesdata2, b == max(b))
bmin <- subset(speciesdata2, b == min(b))

#combine in a dataframe
speciesdata2 <- rbind(fmax, fmin) #change for c and b

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each FINFISH species and multiply f score by (production of finfish /production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*weighting_totals[2, b] #multiply score by the taxa's total production in each year, b is 2 which works because the year starts in the second column
  index <- index/(weighting_totals[1, b]) #weight by production of all species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

library(tidyr)
#make long
weighted_indexes <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 2 times
year_labs <- rep(vec, each = 2) #2 species here

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

#graph
# ggplot(weighted_indexes, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
#   labs(y = "Index", title = "Minimum and Maximum FCB Potential of Finfish")+
#     theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
#       scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))

#SAVE
# ggsave("FCB_weightedby_molluscs.png", plot, height = 4, width = 8, dpi = 600) #make sure wd is set


#ALGAE
speciesdata2 <- filter(speciesdata, category == "algae")

test2 <- filter(test, category == "algae")

fmax <- subset(speciesdata2, f == max(f))
fmin <- subset(speciesdata2, f == min(f))

cmax <- subset(speciesdata2, c == max(c))
cmin <- subset(speciesdata2, c == min(c))

bmax <- subset(speciesdata2, b == max(b))
bmin <- subset(speciesdata2, b == min(b))

#combine in a dataframe
speciesdata2 <- rbind(fmax, fmin)

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each ALGAE species and multiply f score by (production of algae /production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*weighting_totals[4, b] #multiply score by the taxa's total production in each year, b is 2 which works because the year starts in the second column
  index <- index/(weighting_totals[1, b]) #weight by production of all species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

library(tidyr)
#make long
weighted_indexes_algae <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Score")

weighted_indexes_algae <- weighted_indexes_algae[order(weighted_indexes_algae$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 2 times
year_labs <- rep(vec, each = 2) #2 species here

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes_algae)) {
  #replace each year with the actual year number from year_labs
  
  weighted_indexes_algae$Year[i] <- year_labs[i]
}

weighted_indexes_algae$Year <- as.numeric(weighted_indexes_algae$Year) #make years numeric


#MOLLUSCS
speciesdata2 <- filter(speciesdata, category == "mollusc")

test2 <- filter(test, category == "mollusc")

fmax <- subset(speciesdata2, f == max(f))
fmin <- subset(speciesdata2, f == min(f))

cmax <- subset(speciesdata2, c == max(c))
cmin <- subset(speciesdata2, c == min(c))

bmax <- subset(speciesdata2, b == max(b))
bmin <- subset(speciesdata2, b == min(b))

#combine in a dataframe
speciesdata2 <- rbind(fmax, fmin)

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each mollusc species and multiply f score by (production of mollusc /production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*weighting_totals[6, b] #multiply score by the taxa's total production in each year, b is 2 which works because the year starts in the second column
  index <- index/(weighting_totals[1, b]) #weight by production of all species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

library(tidyr)
#make long
weighted_indexes_moll <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Score")

weighted_indexes_moll <- weighted_indexes_moll[order(weighted_indexes_moll$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 2 times
year_labs <- rep(vec, each = 2) #2 species here

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes_moll)) {
  #replace each year with the actual year number from year_labs
  
  weighted_indexes_moll$Year[i] <- year_labs[i]
}

weighted_indexes_moll$Year <- as.numeric(weighted_indexes_moll$Year) #make years numeric

#CRUSTACEANS
speciesdata2 <- filter(speciesdata, category == "crustacean")

test2 <- filter(test, category == "crustacean")

fmax <- subset(speciesdata2, f == max(f))
fmin <- subset(speciesdata2, f == min(f))

cmax <- subset(speciesdata2, c == max(c))
cmin <- subset(speciesdata2, c == min(c))

bmax <- subset(speciesdata2, b == max(b))
bmin <- subset(speciesdata2, b == min(b))

#combine in a dataframe
speciesdata2 <- rbind(fmax, fmin)

#Food security
new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

#loop through each mollusc species and multiply f score by (production of mollusc /production of all species)
#cat_totals (production in THOUSANDS of tons) and speciesdata (f scores) align

for (i in 1:nrow(speciesdata2)){ #for each species
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*weighting_totals[5, b] #multiply score by the taxa's total production in each year, b is 2 which works because the year starts in the second column
  index <- index/(weighting_totals[1, b]) #weight by production of all species
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

library(tidyr)
#make long
weighted_indexes_crust <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Score")

weighted_indexes_crust <- weighted_indexes_crust[order(weighted_indexes_crust$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 2 times
year_labs <- rep(vec, each = 2) #2 species here

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes_crust)) {
  #replace each year with the actual year number from year_labs
  
  weighted_indexes_crust$Year[i] <- year_labs[i]
}

weighted_indexes_crust$Year <- as.numeric(weighted_indexes_crust$Year) #make years numeric
```

#combine taxa on the same graph?
```{r}
#MERGE, put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexes_algae, weighted_indexes_moll, weighted_indexes_crust)

#graph
ggplot(merge, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
  labs(y = "Index", title = "Minimum and Maximum Food Potential by Taxa")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+ scale_color_manual(values=c('#7D627F',"orange",'#7D627F',"orange", "pink", "skyblue", "skyblue", "pink"), name = "Taxa")

# 
#   scale_color_manual(name = "Taxa", values=c('pink', 'skyblue', '#7D627F', "orange"), labels = c("Algae", "Crustaceans", "Finfish", "Molluscs"))


```

#OLD
#use species instead of taxa
```{r}
#ALGAE
speciesdata2 <- filter(speciesdata, category == "algae")
test2 <- filter(test, category == "algae")

fmax <- subset(speciesdata2, f == max(f))
fmin <- subset(speciesdata2, f == min(f))

cmax <- subset(speciesdata2, c == max(c))
cmin <- subset(speciesdata2, c == min(c))

bmax <- subset(speciesdata2, b == max(b))
bmin <- subset(speciesdata2, b == min(b))

#combine in a dataframe
speciesdata2 <- rbind(fmax, fmin)

new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[1, b]) #weight by production of all species, row 1
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

library(tidyr)
#make long
weighted_indexes_algae <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Score")

weighted_indexes_algae <- weighted_indexes_algae[order(weighted_indexes_algae$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 2 times
year_labs <- rep(vec, each = 2) #2 species here

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes_algae)) {
  #replace each year with the actual year number from year_labs
  
  weighted_indexes_algae$Year[i] <- year_labs[i]
}

weighted_indexes_algae$Year <- as.numeric(weighted_indexes_algae$Year) #make years numeric

ggplot(weighted_indexes_algae, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
  labs(y = "Index", title = "Minimum and Maximum FCB Potential of algae")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))
# +
#       scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))

###########################################################
#FINFISH
speciesdata2 <- filter(speciesdata, category == "finfish_inland" | category =="finfish_marine")

test2 <- filter(test, category =="finfish_inland"| category =="finfish_marine")

fmax <- subset(speciesdata2, f == max(f))
fmin <- subset(speciesdata2, f == min(f))

cmax <- subset(speciesdata2, c == max(c))
cmin <- subset(speciesdata2, c == min(c))

bmax <- subset(speciesdata2, b == max(b))
bmin <- subset(speciesdata2, b == min(b))

#combine in a dataframe (F)
speciesdata2 <- rbind(fmax, fmin)

new_frame <- frame #create new frame
new_frame <- new_frame[FALSE, ]

for (i in 1:nrow(speciesdata2)){ #for each species
  # print(speciesdata$species[i])
  f_score <- speciesdata2$f[i] #get the f score
  
  for (b in 2:73){ #for each year
  index <- f_score*test2[i, b+1] #multiply score by that species' total production in each year, b+1 because column two is category, need to start at column 3
  index <- index/(weighting_totals[1, b]) #weight by production of all species, row 1
  
  new_frame[i, b] <- index   #add these numbers to the frame
  new_frame[i, 1] <- speciesdata2[i, 2]#include species
  } 
  
}

library(tidyr)
#make long
weighted_indexes <- pivot_longer(new_frame, cols = -c("species"),
                                names_to = "Year", values_to = "Score")

weighted_indexes <- weighted_indexes[order(weighted_indexes$Year),] #put earlier years first

vec <- c(1950 : 2021)
# replicate each integer in vec 2 times
year_labs <- rep(vec, each = 2) #2 species here

#rename year column in weighted_indexes
for (i in 1:nrow(weighted_indexes)) {
  #replace each year with the actual year number from year_labs
  
  weighted_indexes$Year[i] <- year_labs[i]
}

weighted_indexes$Year <- as.numeric(weighted_indexes$Year) #make years numeric

ggplot(weighted_indexes, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
  labs(y = "Index", title = "Minimum and Maximum food Potential of finfish")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
  ylim(0, 7)
# +
#       scale_color_manual(name = "Index", values=c('#7D627F', '#7DBBC3', "pink", "orange"))

#SAME GRAPH
merge <- rbind(weighted_indexes, weighted_indexes_algae)

#graph
ggplot(merge, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
  labs(y = "Index", title = "Minimum and Maximum Food Security Potential of Finfish and Algae")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(values=c('#7D627F', '#7DBBC3', "pink", "magenta"))

```


#combine taxa on the same graph?
```{r}
#MERGE, put into one dataframe, all stacked up
merge <- rbind(weighted_indexes, weighted_indexes_algae)

#graph
ggplot(merge, aes(x=Year, y=Score, color = species)) +geom_line(size =1)+
  labs(y = "Index", title = "Minimum and Maximum Food Security Potential of Finfish and Algae")+
    theme(axis.text.x = element_text(size = 10), axis.title.y = element_text(size=10))+
      scale_color_manual(values=c('#7D627F', '#7D627F', "pink", "pink"))
```

